### 2023-02-14 bibat/{{cookiecutter.repo_name}}/docs/report.qmd
Switch markdown docs format to quarto

---
title: "{{ cookiecutter.project_name }}"
subtitle: "A report created using bibat version {{ cookiecutter.bibat_version }}"
author: "{{ cookiecutter.author_name }}"
bibliography: bibliography.bib
reference-location: margin
format:
  html: default
  pdf: default
---

Write your report here in [pandoc
markdown](https://pandoc.org/MANUAL.html#pandocs-markdown).

Turn it into a pdf by running `quarto render report.qmd` from the project root.

Refer to entries in `bibliography.bib` using their bibtex keys like
this: @gelmanBayesianWorkflow2020.

Include images like this:

![This is an example image](img/example.png)

Include margin notes like this^[I'm a margin note!]

Write inline code like this:

```{python}
def print_hello():
    print("Hello!")
```

Check out the [quarto guide](https://quarto.org/docs/guide/) for more
information about how to document your work using quarto.

# References

A list of references should appear here:

### 2023-02-14 bibat/hooks/post_gen_project.py

@@ -12,10 +12,11 @@ SPHINX_ONLY_PATHS = [
         "conf.py",
         "index.rst",
         "make.bat",
+        "Makefile",
     )
 ]
-MARKDOWN_ONLY_PATHS = [
-    os.path.join("docs", f) for f in ("report.md", "bibliography.bib")
+QUARTO_ONLY_PATHS = [
+    os.path.join("docs", f) for f in ("report.qmd", "bibliography.bib")
 ]
 
 REMOVE_PATHS = [
@@ -26,8 +27,8 @@ REMOVE_PATHS = [
 if "{{ cookiecutter.docs_format }}" not in ["Sphinx", "No docs"]:
     REMOVE_PATHS += SPHINX_ONLY_PATHS
 
-if "{{ cookiecutter.docs_format }}" not in ["Markdown", "No docs"]:
-    REMOVE_PATHS += MARKDOWN_ONLY_PATHS
+if "{{ cookiecutter.docs_format }}" not in ["Quarto", "No docs"]:
+    REMOVE_PATHS += QUARTO_ONLY_PATHS
 
 for path in REMOVE_PATHS:
     path = path.strip()

### 2023-02-16 bibat/cli.py
Change names of wizard classes

@@ -7,11 +7,11 @@ import argparse
 import os
 from pathlib import Path
 
-import click
 from cookiecutter.main import cookiecutter
 
 from bibat import __version__ as bibat_version
-from bibat.wizarding import WizardFieldChoice, WizardFieldStr, prompt_user
+from bibat.wizarding import WizardChoice, WizardStr, prompt_user
+
 
 THIS_DIR = os.path.dirname(__file__)
 TOOLING_PACKAGES = [
@@ -28,44 +28,42 @@ TOOLING_PACKAGES = [
 ]
 
 WIZARD_FIELDS = [
-    WizardFieldStr(
-        "project_name", "What is your project called?", "Project Name"
-    ),
-    WizardFieldStr("author_name", "What is your name?", "Author name"),
-    WizardFieldStr("author_email", "What is your email?", "Author email"),
-    WizardFieldStr(
+    WizardStr("project_name", "What is your project called?", "Project Name"),
+    WizardStr("author_name", "What is your name?", "Author name"),
+    WizardStr("author_email", "What is your email?", "Author email"),
+    WizardStr(
         "coc_contact",
         "Who should be the code of conduct contact?",
         "Code of conduct contact",
     ),
-    WizardFieldStr(
+    WizardStr(
         "description",
         "Please briefly describe your project",
         "A short description of the project.",
     ),
-    WizardFieldChoice(
+    WizardChoice(
         "open_source_license",
         "Choose an open source license from these options:",
         ["MIT", "BSD-3-Clause", "No license file"],
         "MIT",
     ),
-    WizardFieldChoice(
+    WizardChoice(
         "docs_format",
         "How would you like to document your project?",
-        ["Markdown", "Sphinx", "No docs"],
-        "Sphinx",
+        ["Quarto", "Sphinx", "No docs"],
+        "Quarto",
     ),
-    WizardFieldStr(
+    WizardStr(
         "create_tests_directory",
         "Would you like to create a tests directory?",
         "y",
     ),
-    WizardFieldStr(
+    WizardStr(
         "create_dotgithub_directory",
         "Would you like to create a .github directory?",
         "y",
     ),
-    WizardFieldStr(
+    WizardStr(
         "install_python_tooling",
         "Would you like to install these handy Python tools?\n\t"
         + "\n\t".join(TOOLING_PACKAGES),

### 2023-02-16 bibat/wizarding.py

@@ -7,7 +7,7 @@ from typing import List, Union
 
 
 @dataclass
-class WizardFieldStr:
+class WizardStr:
     """A string field."""
 
     name: str
@@ -16,7 +16,7 @@ class WizardFieldStr:
 
 
 @dataclass
-class WizardFieldChoice:
+class WizardChoice:
     """A choice field."""
 
     name: str
@@ -32,11 +32,11 @@ class WizardFieldChoice:
         return values
 
 
-def prompt_user(wf: Union[WizardFieldStr, WizardFieldChoice]) -> str:
+def prompt_user(wf: Union[WizardStr, WizardChoice]) -> str:
     """Prompt the user for an input and parse it with click."""
-    if isinstance(wf, WizardFieldStr):
+    if isinstance(wf, WizardStr):
         return click.prompt(wf.prompt, default=wf.default, type=str)
-    elif isinstance(wf, WizardFieldChoice):
+    elif isinstance(wf, WizardChoice):
         return click.prompt(
             wf.prompt,
             default=wf.default,
@@ -45,5 +45,5 @@ def prompt_user(wf: Union[WizardFieldStr, WizardFieldChoice]) -> str:
         )
     else:
         raise ValueError(
-            f"input {wf} is not a WizardFieldStr or a WizardFieldChoice"
+            f"input {wf} is not a WizardStr or a WizardChoice"
         )

### 2023-02-16 bibat/examples/baseball/docs/report.qmd
Baseball example

---
title: "Modelling at-bats in baseball using the generalised Pareto distribution"
subtitle: "A report created using bibat version 0.0.4"
author: "Teddy Groves"
bibliography: bibliography.bib
reference-location: margin
format:
  html: default
  pdf: default
filters:
  - include-code-files
---

I used to do a lot of statistical analyses of sports data where there was a
latent parameter for the player's ability. You can see an example
[here](https://github.com/teddygroves/cricket).

It was a natural choice to use a Bayesian hierarchical model where the abilities
have a location/scale type distribution with support on the whole real line, and
in fact this worked pretty well! You would see the kind of players at the top
and bottom of the ability scale that you would expect.

Still, there were a few problems. In particular the data were typically
unbalanced because better players tended to fproduce more data than worse
players. The result of this was that my models would often inappropriately think
the bad players were like the good players: they would not only tend to be too
certain about the abilities of low-data players, but also be biased, thinking
that these players are probably a bit better than they actually are. I never
came up with a good way to solve this problem, despite trying a lot of things!

Even though I don't work with sports data very much any more, the problem still
haunted me, so when I read [this great case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis)
about geomagnetic storms it gave me an idea for yet another potential solution.

The idea was this: just as data about intense storms that cause electricity
problems tell us about a tail of the bigger solar magnetism distribution, maybe
data about professional sportspeople is best thought about as coming from a tail
of the general sports ability distribution. If so, maybe something like the
[generalised pareto
distribution](https://en.wikipedia.org/wiki/Generalized_Pareto_distribution)
might be better than the bog standard normal distribution for describing the
pros' abilities.

I thought I'd test this out with some sports data, and luckily there is a really
nice baseball example on [the Stan case studies
website](https://mc-stan.org/users/documentation/case-studies/pool-binary-trials.html),
complete with [data from the 2006 Major league
season])https://github.com/stan-dev/example-models/blob/master/knitr/pool-binary-trials/baseball-hits-2006.csv).
After I posted some early results [on the Stan discourse
forum](https://discourse.mc-stan.org/t/baseball-analysis-using-latent-generalised-pareto-distribution/29222),
user jd_c kindly gathered and posted [even more
data](https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData),
covering 5 full major league baseball seasons.

With a few datasets and at least two different statistical models to consider,
the full analysis looked like it would be a little too big to fit in a trivial
file structure, so it seemed like a good opportunity to showcase my
batteries-included Bayesian analysis template package
[bibat](https://github.com/teddygroves/bibat).

The rest of this vignette describes how I used bibat to (relatively) painlessly
see if my generalised Pareto distribution idea would work.

Check out the analysis's [github
repository](https://github.com/teddygroves/baseball) for all the details.

# Setup

First I installed bibat in my global Python 3.11 environment with this command:

```{sh}
$ pip install bibat
```

Next I started bibat's wizard like this:

```{sh}
bibat
```

After I answered the wizard's questions bibat creted a new folder called
`baseball` that looked like this:

```{sh}
baseball
├── CODE_OF_CONDUCT.md
├── LICENSE
├── Makefile
├── README.md
├── bibat_version.txt
├── data
│   └── raw
│       ├── raw_measurements.csv
│       └── readme.md
├── docs
│   ├── bibliography.bib
│   ├── img
│   │   ├── example.png
│   │   └── readme.md
│   └── report.qmd
├── inferences
│   ├── fake_interaction
│   │   └── config.toml
│   ├── interaction
│   │   └── config.toml
│   └── no_interaction
│       └── config.toml
├── investigate.ipynb
├── plots
│   ├── posterior_ll_comparison.png
│   └── posterior_predictive_comparison.png
├── prepare_data.py
├── pyproject.toml
├── requirements-tooling.txt
├── requirements.txt
├── sample.py
├── src
│   ├── __init__.py
│   ├── data_preparation_functions.py
│   ├── inference_configuration.py
│   ├── prepared_data.py
│   ├── readme.md
│   ├── stan
│   │   ├── custom_functions.stan
│   │   ├── multilevel-linear-regression.stan
│   │   └── readme.md
│   ├── stan_input_functions.py
│   └── util.py
└── tests
    ├── test_integration
    │   └── test_data_preparation.py
    └── test_unit
        ├── test_inference_configuration.py
        └── test_util.py
```

This folder implements bibat's example analysis - a comparison of linear
regression with two different design matrices. To check that everything was
working correctly I ran the analysis like this:

```{sh}
$ cd baseball
$ make analysis
```

Some cogs turned, some new lines appeared in my terminal and some new files were
created - nice, the setup worked!

# Getting raw data

I added a line with the text `pyreadr` to the file `requirements.txt`, then
added the following code to a new script called `fetch_data.py`:

```{.python include="../fetch_data.py"}
```

To get the files I ran the script:

```{sh}
$ source .venv/bin/activate
$ python fetch_data.py
```

Finally I removed the example analysis's raw data:

```{sh}
$ rm data/raw/raw_measurements.csv
```

# Preparing the data
The first step in preparing data is to decide what prepared data looks like for
the purposes of our analysis. Bibat provides dataclasses called `PreparedData`
and `MeasurementsDF ` in the file `src/prepared_data.py` which can help get us
started with this.

As it happens, prepared data looks very similar in our analysis and the example.
All we need to do is change the `MeasurementsDF` definition a little^[note that
this class uses [pandera](https://pandera.readthedocs.io), a handy library for
defining what a pandas dataframe should look like]:

```{python}
from typing import Optional
import pandera as pa

# ...

class MeasurementsDF(pa.SchemaModel):
    """A PreparedData should have a measurements dataframe like this.

    Other columns are also allowed!
    """

    player: pa.typing.Series[str]
    season: pa.typing.Series[str]
    n_attempt: pa.typing.Series[int] = pa.Field(ge = 1)
    n_success: pa.typing.Series[int] = pa.Field(ge = 0)
```

Next we can write some functions that create `PreparedData` objects. These
should live in the file `data_preparation_functions.py`. In this case we write a
couple of data preparation functions: `prepare_data_2006` and
`prepare_data_recent`:

```{.python include="../src/data_preparation_functions.py"}
```

Finally we need to update `prepare_data.py`, the script that runs the data
preparation functions. Again there isn't much to change from the example
analysis.

```{.python include="../prepare_data.py"}
```

To check that all this works, we can run the script `prepare_data.py` manually
or using `make analysis`. Now if we look in the file
`data/prepared/recent/measurements.csv` we should see some lines that look like
this:

```{csv}
,player,season,n_attempt,n_success
0,abreujo02,2018,553,180
1,acunaro01,2018,487,178
3,adamewi01,2018,322,112
6,adamsla01,2018,29,10
7,adamsma01,2018,337,104
8,adamsma01,2018,277,92
9,adamsma01,2018,60,12
```

# Specifying statistical models

I wanted to test two statistical models: one with the modelling the distribution
of per-player logit-scale at-bat success rates as a normal distbution with
unknown mean and standard deviation, and another where the same logit-scale
rates have a generalised pareto distribution.

So, genve a table of $N$ player profiles, with each player has $y$ successes out
of $K$ at-bats and an unknown latent success rate $\alpha$, I wanted to use this
measurement model:

$$
y \sim \text{binomial logit}(K, \alpha)
$$

In the generalised pareto model I would give the $\alpha$s this prior model,
with the hyperparameter $\text{min }\alpha$ assumed to be known exactly and $k$
and $\sigma$ given prior distributions that put the $\alpha$s in the generally
plausible range of between roughly 0.1 and 0.4.

\begin{align*}
\alpha &\sim GPareto(\text{min }\alpha, k, \sigma)
\end{align*}

In the normal model I would use a standard hierarchical regression model with an
effect for the log-scale number of at-bats to attempt to explicitly model the
tendency of players with more appearances to be better:

\begin{align*}
\alpha &\sim Normal(\mu + b_{K} \cdot \ln{K}, \tau) \\
\end{align*}

Again I would choose priors for the hyperparameters that put most of the alphas
between 0.1 and 0.4.

To implement these models using Stan I first added the following function to the
file `custom_functions.stan`. This was simply copied from [the relevant part of
the geomagnetic storms case
study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis).

```{stan}
real gpareto_lpdf(vector y, real ymin, real k, real sigma) {
  // generalised Pareto log pdf 
  int N = rows(y);
  real inv_k = inv(k);
  if (k<0 && max(y-ymin)/sigma > -inv_k)
    reject("k<0 and max(y-ymin)/sigma > -1/k; found k, sigma =", k, ", ", sigma);
  if (sigma<=0)
    reject("sigma<=0; found sigma =", sigma);
  if (fabs(k) > 1e-15)
    return -(1+inv_k)*sum(log1p((y-ymin) * (k/sigma))) -N*log(sigma);
  else
    return -sum(y-ymin)/sigma -N*log(sigma); // limit k->0
}
```

Next I wrote a file `gpareto-reg.stan`:

```{.stan include="../src/stan/gpareto-reg.stan"}
```

Finally I wrote a file `normal.stan`:

```{.stan include="../src/stan/normal.stan"}
```

# Generating Stan inputs
Next we need to tell our analysis how to turn some prepared data into a
dictionary that can be used as input for Stan. Bibat assumes that this task is
handled by functions that live in the file `src/stan_input_functions.py`, each
of which takes in a `PreparedData` and returns a Python dictionary. You can
write as many Stan input functions as you like and choose which one to run for
any given inference.

We can start by defining some Stan input functions that pass arbitary prepared data
on to each of the models:

```{.python include="../src/stan_input_functions.py" start-line=12 end-line=34}
```

But why stop there? It can also be useful to generate Stan inputs consistently
with a model, based on some hardcoded hyperparameter values. Here are some
functions that do this for both of our models:

```{.python include="../src/stan_input_functions.py" start-line=37}
```

# Specifying inferences
Now all the building blocks for making statistical inferences - raw data, data
preparation rules, statistical models and recipes for turning prepared data into
model inputs - are in place. The last step before actually running Stan is to
write down how put these blocks together. Bibat has another concept for this,
called 'inferences'.

An inference in bibat is a folder containing a special file called
`config.toml`. This file sets out what inferences you want to make: which
statistical model, which prepared data function, which Stan input function,
which parameters have which dimensions, which sampling modes to use and how to
configure the sampler. The folder will later be filled up with the results of
performing the specified inferences.

I started by deleting the example inferences and creating two fresh folders,
leaving me with an `inferences` folder looking like this:

```{sh}
.
├── gpareto2006
│   └── config.toml
└── normal2006
    └── config.toml
```

Here is the file `inferences/gpareto2006/config.toml`:

```{.toml include="../inferences/gpareto2006/config.toml"}
```

Here is the file `inferences/normal2006/config.toml`:

```{.toml include="../inferences/normal2006/config.toml"}
```

Note that:
* The Stan file, prepared data folder and stan input function are referred to by
  strings. The analysis should raise an error if you enter a non-existing value.
* Both inferences are set to run in "prior" and "posterior" modes - the other
  pre-defined mode is "kfold", but you can also write your own!
* You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
  in the `[sample_kwargs]` table.
* You can enter mode-specific overrides in `[sample_kwargs.<MODE>]`. This can be
  handy if you want to run more or fewer iterations for a certain mode.

Now when I ran `make analysis` again, I saw messages indicating that Stan had
run, and found that the `inferences` subfolders had been populated:

```{sh}
inferences
├── gpareto2006
│   ├── config.toml
│   └── idata.json
└── normal2006
    ├── config.toml
    └── idata.json
```

# Investigating the inferences
Now that the inferences are ready it's time to check them out. Bibat provides a
jupyter notebook called `investigate.ipynb` for exactly this purpose.

A lot of code from the example analysis's notebook was reusable, so I largely
followed its structure, with a few tweaks.

# Choosing priors using push-forward calibration

The trickiest thing about my analysis was setting prior distributions for the
parameters $k$ and $\sigma$ in the generalised Pareto models. To choose some
more or less plausible values I did a few prior-mode model runs and checked the
distributions of the resulting `alpha` variables. I wanted to make sure that
they all lay in the range corresponding to batting averages between about 0.15
and a little over 0.4. Here is a graph that shows the 1% to 99% prior quantiles
for each player's latent success percentage in the 2006 dataset alongside their
actually realised success rate.

![](img/prior_quantiles.png)

# Extending the analysis to the recent data

To model the more recent data, all I had to do was create some new inference
folders with appropriate `prepared_data_dir` fields in their `config.toml`
files. For example, here is the `config.toml` file for the `gparetorecent`
inference:

```{.toml include="../inferences/gparetorecent/config.toml"}
```

After running `make analysis` one more time, I went back to the notebook
`investigate.ipynb` and made plots of both models' posterior 1%-99% success rate
intervals for both datasets:

![](img/posterior_quantiles.png)

I think this is very interesting. Both models' prior distributions had similar
regularisation levels, and they more or less agree about the abilities of the
players with the most at-bats, both in terms of locations and the widths of the
plausible intervals for the true success rate. However, the models ended up with
dramatically different certainty levels about the abilities of players with few
at-bats. This pattern was true both for the small 2006 dataset and the much
larger 'recent' dataset.

### 2023-02-16 bibat/examples/baseball/fetch_data.py

@@ -1,0 +1,25 @@
+import os
+import requests
+import pandas as pd
+import pyreadr
+
+URLS = {
+    "2006": "https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
+    "recent": "https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData"
+}
+OUT_FILES = {
+    "2006": os.path.join("data", "raw", "baseball-hits-2006.csv"),
+    "recent": os.path.join("data", "raw", "recent.RData")
+}
+
+if __name__ == "__main__":
+    print(f"Fetching 2006 data from {URLS['2006']}")
+    data_2006 = pd.read_csv(URLS['2006'], comment="#")
+    print(f"Writing 2006 data to {OUT_FILES['2006']}")
+    data_2006.to_csv(OUT_FILES['2006'])
+    print(f"Fetching recent data from {URLS['recent']}")
+    with requests.get(URLS["recent"]) as f:
+        recent = f.text
+    print(f"Writing to {OUT_FILES['recent']}")
+    with open(OUT_FILES["recent"], "wb") as f:
+        f.write(recent.encode("utf8", errors="replace"))

### 2023-02-16 bibat/examples/baseball/prepare_data.py

@@ -1,0 +1,41 @@
+"""Read the data in RAW_DIR and save prepared data to PREPARED_DIR."""
+
+import os
+
+import pandas as pd
+from src import data_preparation_functions
+from src.prepared_data import write_prepared_data
+
+DATA_PREPARATION_FUNCTIONS_TO_RUN = {
+    "2006": data_preparation_functions.prepare_data_2006,
+    "recent": data_preparation_functions.prepare_data_recent,
+}
+
+DATA_DIR = os.path.join(os.path.dirname(__file__), "data")
+RAW_DIR = os.path.join(DATA_DIR, "raw")
+PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
+RAW_DATA_FILES = {
+    "2006": os.path.join(RAW_DIR, "baseball-hits-2006.csv"),
+    "recent": os.path.join(RAW_DIR, "recent.csv"),
+}
+
+
+def main():
+    """Save prepared data in the PREPARED_DIR."""
+    print("Reading raw data...")
+    raw_data = {
+        k: pd.read_csv(v, index_col=None) for k, v in RAW_DATA_FILES.items()
+    }
+    print("Preparing data...")
+    for name, dpf in DATA_PREPARATION_FUNCTIONS_TO_RUN.items():
+        print(f"Running data preparation function {dpf.__name__}...")
+        prepared_data = dpf(raw_data[name])
+        output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
+        print(f"\twriting files to {output_dir}")
+        if not os.path.exists(PREPARED_DIR):
+            os.mkdir(PREPARED_DIR)
+        write_prepared_data(prepared_data, output_dir)
+
+
+if __name__ == "__main__":
+    main()

### 2023-02-16 bibat/examples/baseball/sample.py

@@ -1,0 +1,121 @@
+"""Run all the configs in the model_configurations folder."""
+
+import os
+from typing import Dict
+
+import arviz as az
+import cmdstanpy
+import numpy as np
+import xarray as xr
+from sklearn.model_selection import KFold
+from src.inference_configuration import (
+    AVAILABLE_MODES,
+    InferenceConfiguration,
+    load_inference_configuration,
+)
+from src.prepared_data import load_prepared_data
+from src.util import CoordDict
+
+RUNS_DIR = "inferences"
+STAN_DIR = os.path.join("src", "stan")
+
+
+def get_sample_kwargs(raw: Dict, mode: str) -> Dict:
+    """Get the sampel kwargs for a mode from raw toml."""
+    out = {k: v for k, v in raw.items() if k not in AVAILABLE_MODES}
+    if mode in raw.keys():
+        out |= raw[mode]
+    return out
+
+
+def get_kfold_llik(
+    ic: InferenceConfiguration,
+    stan_input_base: Dict,
+    model: cmdstanpy.CmdStanModel,
+    coords: CoordDict,
+) -> xr.DataArray:
+    """Run k-fold cross-validation and return out-of-sample log likelihood.
+
+    :param ic: InferenceConfiguration, must have non-null field kfold_folds
+
+    :param stan_input_base: stan input dictionary
+
+    :param model: CmdStanModel
+    """
+    n_folds = ic.kfold_folds
+    sample_kwargs = get_sample_kwargs(ic.sample_kwargs, "kfold")
+    kf = KFold(n_folds, shuffle=True, random_state=1234)
+    lliks_by_fold = []
+    full_ix = np.array(stan_input_base["ix_train"])
+    for fold, (ix_train, ix_test) in enumerate(kf.split(full_ix)):
+        input_dict = stan_input_base | {
+            "likelihood": 1,
+            "N_train": len(ix_train),
+            "N_test": len(ix_test),
+            "ix_train": full_ix[ix_train].tolist(),
+            "ix_test": full_ix[ix_test].tolist(),
+        }
+        mcmc = model.sample(data=input_dict, **sample_kwargs)
+        llik_fold = mcmc.draws_xr(vars=["llik"])
+        # remember the fold
+        llik_fold["fold"] = fold
+        llik_fold = llik_fold.set_coords("fold")
+        # set value of index "chain" to zero to match arviz convention
+        llik_fold = llik_fold.assign_coords(
+            {"new_chain": ("chain", [0])}
[…]

### 2023-02-16 bibat/examples/baseball/src/__init__.py

@@ -1,0 +1,1 @@
+"""Source code for baseball."""

### 2023-02-16 bibat/examples/baseball/src/data_preparation_functions.py

@@ -1,0 +1,75 @@
+"""Provides functions prepare_data_x.
+
+These functions should take in a dataframe of measurements and return a
+PreparedData object.
+
+"""
+
+import pandas as pd
+from src.prepared_data import PreparedData
+
+
+def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:
+    """Prepare the 2006 data."""
+    measurements = (
+        measurements_raw
+        .rename(columns={"K": "n_attempt", "y": "n_success"})
+        .assign(
+            season="2006",
+            player=lambda df: [f"2006-player-{i+1}" for i in range(len(df))]
+        )
+    )
+    return PreparedData(
+        name="2006",
+        coords={
+            "player": measurements["player"].tolist(),
+            "season": measurements["season"].tolist()
+        },
+        measurements=measurements,
+    )
+
+
+def prepare_data_recent(measurements_raw: pd.DataFrame) -> PreparedData:
+    """Prepare the recent data.
+
+    There are a few substantive data choices here.
+
+    First, the function excludes players who have a '1' in their position as
+    these are likely pitchers, as well as players with fewer than 20 at bats.
+
+    Second, the function defines a successes and attempts according to the
+    'on-base percentage' metric, so a success is a time when a player got a hit,
+    a base on ball/walk or a hit-by-pitch and an attempt is an at-bat or a
+    base-on-ball/walk or a hit-by-pitch or a sacrifice fly. This could have
+    alternatively been calculated as just hits divided by at-bats, but my
+    understanding is that this method underrates players who are good at getting
+    walks.
+
+    """
+    def filter_batters(df: pd.DataFrame):
+        return (
+            (df["AB"] >= 20)
+            & ~df["Pos.Summary"].astype(str).str.contains("1")
+        )
+    measurements = (
+        measurements_raw
+        .assign(
+            season=lambda df: df["year"].astype(str),
+            player=lambda df: df["Name.additional"].str.cat(df["Tm"]).str.cat(df["Lg"]).str.cat(df["season"]),
+            n_attempt=lambda df: df[["AB", "BB", "HBP", "SF"]].fillna(0).sum(
+                axis=1).astype(int),
+            n_success=lambda df: (
+                df[["H", "BB", "HBP"]].fillna(0).sum(axis=1).astype(int)
+            )
+        )
+        .loc[filter_batters, ["player", "season", "n_attempt", "n_success"]]
+        .copy()
+    )
+    return PreparedData(
[…]

### 2023-02-16 bibat/examples/baseball/src/prepared_data.py

@@ -1,0 +1,58 @@
+"""Class defining what prepared data looks like."""
+
+import json
+import os
+
+import pandas as pd
+import pandera as pa
+from pydantic.dataclasses import dataclass
+from src.util import CoordDict
+
+NAME_FILE = "name.txt"
+COORDS_FILE = "coords.json"
+MEASUREMENTS_FILE = "measurements.csv"
+
+class MeasurementsDF(pa.SchemaModel):
+    """A PreparedData should have a measurements dataframe like this.
+
+    Other columns are also allowed!
+    """
+
+    player: pa.typing.Series[str]
+    season: pa.typing.Series[str] = pa.Field(coerce=True)
+    n_attempt: pa.typing.Series[int] = pa.Field(ge = 1)
+    n_success: pa.typing.Series[int] = pa.Field(ge = 0)
+
+
+@dataclass
+class PreparedData:
+    """What prepared data looks like in this analysis."""
+
+    name: str
+    coords: CoordDict
+    measurements: pa.typing.DataFrame[MeasurementsDF]
+
+
+def load_prepared_data(directory: str) -> PreparedData:
+    """Load prepared data from files in directory."""
+    with open(os.path.join(directory, COORDS_FILE), "r") as f:
+        coords = json.load(f)
+    with open(os.path.join(directory, NAME_FILE), "r") as f:
+        name = f.read()
+    measurements = pd.read_csv(os.path.join(directory, MEASUREMENTS_FILE))
+    return PreparedData(
+        name=name,
+        coords=coords,
+        measurements=measurements,
+    )
+
+
+def write_prepared_data(prepped: PreparedData, directory):
+    """Write prepared data files to a directory."""
+    if not os.path.exists(directory):
+        os.mkdir(directory)
+        prepped.measurements.to_csv(os.path.join(directory, MEASUREMENTS_FILE))
+    with open(os.path.join(directory, COORDS_FILE), "w") as f:
+        json.dump(prepped.coords, f)
+    with open(os.path.join(directory, NAME_FILE), "w") as f:
+        f.write(prepped.name)

### 2023-02-16 bibat/examples/baseball/src/stan_input_functions.py

@@ -1,0 +1,90 @@
+"""Functions for generating input to Stan from prepared data."""
+
+
+from typing import Dict, List
+
+import numpy as np
+import pandas as pd
+from scipy.special import expit, logit
+from src.prepared_data import PreparedData
+
+
+def get_stan_input_normal(ppd: PreparedData) -> Dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"].values,
+        "y": ppd.measurements["n_success"].values,
+        "prior_mu": [logit(0.25), 0.2],
+        "prior_tau": [0.2, 0.1],
+        "prior_b_K": [0, 0.03],
+    }
+
+
+def get_stan_input_gpareto(ppd: PreparedData) -> Dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"].values,
+        "y": ppd.measurements["n_success"].values,
+        "min_alpha": logit(0.1),
+        "max_alpha": logit(0.5),
+        "prior_sigma": [1.5, 0.4],
+        "prior_k": [-0.5, 1],
+    }
+
+
+def get_stan_input_normal_fake(ppd: PreparedData) -> Dict:
+    """Generate fake Stan input consistent with the normal model."""
+    N = len(ppd.measurements)
+    rng = np.random.default_rng(seed=1234)
+    true_param_values = {
+        "mu": logit(0.25),
+        "tau": 0.18,   # 2sds is 0.19 to 0.32 batting average
+        "b_K": 0.04,    # slight effect of more attempts
+        "alpha_std": rng.random.normal(loc=0, scale=1, size=N)
+    }
+    K = ppd.measurements["n_attempt"].values
+    log_K_std = (np.log(K) - np.log(K).mean()) / np.log(K).std()
+    alpha = (
+        true_param_values["mu"]
+        + true_param_values["b_K"] * log_K_std
+        + true_param_values["tau"] * true_param_values["alpha_std"]
+    )
+    y = rng.random.binomial(K, expit(alpha))
+    return {"N": N, "K": K, "y": y} | true_param_values
+
+
+def get_stan_input_gpareto_fake(ppd: PreparedData) -> Dict:
+    """Generate fake Stan input consistent with the gpareto model."""
+    N = len(ppd.measurements)
[…]

### 2023-02-16 bibat/examples/baseball/src/util.py

@@ -1,0 +1,83 @@
+"""Some handy python functions."""
+
+from typing import Dict, List, NewType, Optional, Union
+
+import numpy as np
+import pandas as pd
+
+CoordDict = NewType("CoordDict", Dict[str, List[str]])
+StanInputDict = Dict["str", Union[int, float, List]]
+
+
+def one_encode(s: pd.Series) -> pd.Series:
+    """Replace a series's values with 1-indexed integer factors.
+
+    :param s: a pandas Series that you want to factorise.
+
+    """
+    return pd.Series(pd.factorize(s)[0] + 1, index=s.index)
+
+
+def make_columns_lower_case(df: pd.DataFrame) -> pd.DataFrame:
+    """Make a DataFrame's columns lower case.
+
+    :param df: a pandas DataFrame
+    """
+    new = df.copy()
+    if isinstance(new.columns, pd.MultiIndex):
+        new.columns = pd.MultiIndex.from_arrays(
+            [
+                [c.lower() for c in new.columns.get_level_values(i)]
+                for i in range(len(new.columns.levels))
+            ]
+        )
+    else:
+        new.columns = pd.Index([c.lower() for c in new.columns])
+    return new
+
+
+def check_is_df(maybe_df) -> pd.DataFrame:
+    """Shut up the type checker."""
+    assert isinstance(maybe_df, pd.DataFrame)
+    return maybe_df
+
+
+def stanify_dict(d: Dict) -> StanInputDict:
+    """Make sure a dictionary is a valid Stan input.
+
+    :param d: input dictionary, possibly with wrong types
+    """
+    out: StanInputDict = {}
+    for k, v in d.items():
+        if not isinstance(k, str):
+            raise ValueError(f"key {str(k)} is not a string!")
+        elif isinstance(v, pd.Series):
+            out[k] = v.to_list()
+        elif isinstance(v, pd.DataFrame):
+            out[k] = v.values.tolist()
+        elif isinstance(v, np.ndarray):
+            out[k] = v.tolist()
+        else:
+            out[k] = v
+    return out
+
+
+def standardise(
+    s: pd.Series, mu: Optional[float] = None, std: Optional[float] = None
+) -> pd.Series:
+    """Standardise a series by subtracting mu and dividing by sd."""
+    if mu is None:
+        mu = s.mean()
+    if std is None:
+        std = s.std()
+    return s.subtract(mu).divide(std)
+
+
+def center(
[…]

### 2023-02-16 bibat/examples/baseball/docs/report.qmd
bump version

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.0.4"
+subtitle: "A report created using bibat version 0.0.5"
 author: "Teddy Groves"

### 2023-02-16 bibat/cli.py
Run black and isort

@@ -12,7 +12,6 @@ from cookiecutter.main import cookiecutter
 from bibat import __version__ as bibat_version
 from bibat.wizarding import WizardChoice, WizardStr, prompt_user
 
-
 THIS_DIR = os.path.dirname(__file__)
 TOOLING_PACKAGES = [
     "pandas-stubs",

### 2023-02-16 bibat/examples/baseball/fetch_data.py

@@ -1,22 +1,23 @@
 import os
-import requests
+
 import pandas as pd
 import pyreadr
+import requests
 
 URLS = {
     "2006": "https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
-    "recent": "https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData"
+    "recent": "https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData",
 }
 OUT_FILES = {
     "2006": os.path.join("data", "raw", "baseball-hits-2006.csv"),
-    "recent": os.path.join("data", "raw", "recent.RData")
+    "recent": os.path.join("data", "raw", "recent.RData"),
 }
 
 if __name__ == "__main__":
     print(f"Fetching 2006 data from {URLS['2006']}")
-    data_2006 = pd.read_csv(URLS['2006'], comment="#")
+    data_2006 = pd.read_csv(URLS["2006"], comment="#")
     print(f"Writing 2006 data to {OUT_FILES['2006']}")
-    data_2006.to_csv(OUT_FILES['2006'])
+    data_2006.to_csv(OUT_FILES["2006"])
     print(f"Fetching recent data from {URLS['recent']}")
     with requests.get(URLS["recent"]) as f:
         recent = f.text

### 2023-02-16 bibat/examples/baseball/src/data_preparation_functions.py

@@ -11,19 +11,17 @@ from src.prepared_data import PreparedData
 
 def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:
     """Prepare the 2006 data."""
-    measurements = (
-        measurements_raw
-        .rename(columns={"K": "n_attempt", "y": "n_success"})
-        .assign(
-            season="2006",
-            player=lambda df: [f"2006-player-{i+1}" for i in range(len(df))]
-        )
+    measurements = measurements_raw.rename(
+        columns={"K": "n_attempt", "y": "n_success"}
+    ).assign(
+        season="2006",
+        player=lambda df: [f"2006-player-{i+1}" for i in range(len(df))],
     )
     return PreparedData(
         name="2006",
         coords={
             "player": measurements["player"].tolist(),
-            "season": measurements["season"].tolist()
+            "season": measurements["season"].tolist(),
         },
         measurements=measurements,
     )
@@ -46,21 +44,26 @@ def prepare_data_recent(measurements_raw: pd.DataFrame) -> PreparedData:
     walks.
 
     """
+
     def filter_batters(df: pd.DataFrame):
-        return (
-            (df["AB"] >= 20)
-            & ~df["Pos.Summary"].astype(str).str.contains("1")
+        return (df["AB"] >= 20) & ~df["Pos.Summary"].astype(str).str.contains(
+            "1"
         )
+
     measurements = (
-        measurements_raw
-        .assign(
+        measurements_raw.assign(
             season=lambda df: df["year"].astype(str),
-            player=lambda df: df["Name.additional"].str.cat(df["Tm"]).str.cat(df["Lg"]).str.cat(df["season"]),
-            n_attempt=lambda df: df[["AB", "BB", "HBP", "SF"]].fillna(0).sum(
-                axis=1).astype(int),
+            player=lambda df: df["Name.additional"]
+            .str.cat(df["Tm"])
+            .str.cat(df["Lg"])
+            .str.cat(df["season"]),
+            n_attempt=lambda df: df[["AB", "BB", "HBP", "SF"]]
+            .fillna(0)
+            .sum(axis=1)
+            .astype(int),
             n_success=lambda df: (
                 df[["H", "BB", "HBP"]].fillna(0).sum(axis=1).astype(int)
-            )
+            ),
         )
         .loc[filter_batters, ["player", "season", "n_attempt", "n_success"]]
         .copy()
[…]

### 2023-02-16 bibat/examples/baseball/src/prepared_data.py

@@ -12,6 +12,7 @@ NAME_FILE = "name.txt"
 COORDS_FILE = "coords.json"
 MEASUREMENTS_FILE = "measurements.csv"
 
+
 class MeasurementsDF(pa.SchemaModel):
     """A PreparedData should have a measurements dataframe like this.
 
@@ -20,8 +21,8 @@ class MeasurementsDF(pa.SchemaModel):
 
     player: pa.typing.Series[str]
     season: pa.typing.Series[str] = pa.Field(coerce=True)
-    n_attempt: pa.typing.Series[int] = pa.Field(ge = 1)
-    n_success: pa.typing.Series[int] = pa.Field(ge = 0)
+    n_attempt: pa.typing.Series[int] = pa.Field(ge=1)
+    n_success: pa.typing.Series[int] = pa.Field(ge=0)
 
 
 @dataclass

### 2023-02-16 bibat/examples/baseball/src/stan_input_functions.py

@@ -40,9 +40,9 @@ def get_stan_input_normal_fake(ppd: PreparedData) -> Dict:
     rng = np.random.default_rng(seed=1234)
     true_param_values = {
         "mu": logit(0.25),
-        "tau": 0.18,   # 2sds is 0.19 to 0.32 batting average
-        "b_K": 0.04,    # slight effect of more attempts
-        "alpha_std": rng.random.normal(loc=0, scale=1, size=N)
+        "tau": 0.18,  # 2sds is 0.19 to 0.32 batting average
+        "b_K": 0.04,  # slight effect of more attempts
+        "alpha_std": rng.random.normal(loc=0, scale=1, size=N),
     }
     K = ppd.measurements["n_attempt"].values
     log_K_std = (np.log(K) - np.log(K).mean()) / np.log(K).std()
@@ -64,27 +64,23 @@ def get_stan_input_gpareto_fake(ppd: PreparedData) -> Dict:
     true_param_values = {
         "sigma": -1.098,
         "k": 0.18,
-        "alpha":gpareto_rvs(rng, N, min_alpha, k, sigma)
+        "alpha": gpareto_rvs(rng, N, min_alpha, k, sigma),
     }
     y = rng.binomial(k, expit(alpha))
     return {"N": N, "K": k, "y": y, "min_alpha": min_alpha} | true_param_values
 
 
 def gpareto_rvs(
-        rng: np.random.Generator,
-        size: int,
-        mu: float,
-        k: float,
-        sigma: float
+    rng: np.random.Generator, size: int, mu: float, k: float, sigma: float
 ):
     """Generate random numbers from a generalised pareto distribution.
-    
+
     See https://en.wikipedia.org/wiki/Generalized_Pareto_distribution for
     source.
-    
+
     """
     U = rng.uniform(size)
     if k == 0:
         return mu - sigma * np.log(U)
     else:
-        return mu + (sigma * (U ** -k) - 1) / sigma
+        return mu + (sigma * (U**-k) - 1) / sigma

### 2023-02-16 bibat/wizarding.py

@@ -1,9 +1,10 @@
 """Dataclasses and functions for running bibat's command line wizard."""
 
-import click
-from pydantic.dataclasses import dataclass
-from pydantic import root_validator
 from typing import List, Union
+
+import click
+from pydantic import root_validator
+from pydantic.dataclasses import dataclass
 
 
 @dataclass
@@ -44,6 +45,4 @@ def prompt_user(wf: Union[WizardStr, WizardChoice]) -> str:
             show_choices=True,
         )
     else:
-        raise ValueError(
-            f"input {wf} is not a WizardStr or a WizardChoice"
-        )
+        raise ValueError(f"input {wf} is not a WizardStr or a WizardChoice")

### 2023-02-16 bibat/examples/baseball/fetch_data.py
flake8

@@ -1,12 +1,15 @@
+"""Fetch raw data from the internet."""
+
 import os
 
 import pandas as pd
-import pyreadr
 import requests
 
 URLS = {
-    "2006": "https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
-    "recent": "https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData",
+    "2006": "https://raw.githubusercontent.com/stan-dev/"
+    "example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
+    "recent": "https://discourse.mc-stan.org/uploads/short-url/"
+    "cRJtia2jKxiX01ZEraHUVK7jbHI.RData",
 }
 OUT_FILES = {
     "2006": os.path.join("data", "raw", "baseball-hits-2006.csv"),

### 2023-02-16 bibat/examples/baseball/src/stan_input_functions.py

@@ -1,10 +1,9 @@
 """Functions for generating input to Stan from prepared data."""
 
 
-from typing import Dict, List
+from typing import Dict
 
 import numpy as np
-import pandas as pd
 from scipy.special import expit, logit
 from src.prepared_data import PreparedData
 
@@ -61,13 +60,16 @@ def get_stan_input_gpareto_fake(ppd: PreparedData) -> Dict:
     K = ppd.measurements["n_attempt"].values
     min_alpha = 0.1
     rng = np.random.default_rng(seed=1234)
-    true_param_values = {
-        "sigma": -1.098,
-        "k": 0.18,
-        "alpha": gpareto_rvs(rng, N, min_alpha, k, sigma),
-    }
-    y = rng.binomial(k, expit(alpha))
-    return {"N": N, "K": k, "y": y, "min_alpha": min_alpha} | true_param_values
+    true_param_values = {"sigma": -1.098, "k": 0.18}
+    true_param_values["alpha"] = gpareto_rvs(
+        rng,
+        N,
+        min_alpha,
+        true_param_values["k"],
+        true_param_values["sigma"],
+    )
+    y = rng.binomial(K, expit(true_param_values["alpha"]))
+    return {"N": N, "K": K, "y": y, "min_alpha": min_alpha} | true_param_values
 
 
 def gpareto_rvs(

### 2023-02-16 bibat/examples/baseball/docs/report.qmd
Correct wrong path in baseball report
bump version

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.0.5"
+subtitle: "A report created using bibat version 0.0.6"
 author: "Teddy Groves"
@@
 
-Next I wrote a file `gpareto-reg.stan`:
+Next I wrote a file `gpareto.stan`:
 
-```{.stan include="../src/stan/gpareto-reg.stan"}
+```{.stan include="../src/stan/gpareto.stan"}
 ```

### 2023-02-17 bibat/examples/baseball/docs/report.qmd
Bump version

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.0.6"
+subtitle: "A report created using bibat version 0.0.7"
 author: "Teddy Groves"

### 2023-02-20 bibat/examples/baseball/docs/report.qmd
Update case study

@@
 forum](https://discourse.mc-stan.org/t/baseball-analysis-using-latent-generalised-pareto-distribution/29222),
-user jd_c kindly gathered and posted [even more
-data](https://discourse.mc-stan.org/uploads/short-url/cRJtia2jKxiX01ZEraHUVK7jbHI.RData),
-covering 5 full major league baseball seasons.
+other users suggested that it might be interesting to model similar data from
+different seasons. This data can now be found quite easily on the [baseball
+databank](https://baseballdb.lawlesst.net/).
 
@@
 
-I added a line with the text `pyreadr` to the file `requirements.txt`, then
-added the following code to a new script called `fetch_data.py`:
+To fetch raw data from the internet, I wrote a new script called `fetch_data.py`:
 
@@
 $ python fetch_data.py
+```
+
+Since this worked, I added a new makefile target for the raw data files:
+
+```{Makefile}
+RAW_DATA = data/raw/2006.csv data/raw/bdb-main.csv data/raw/bdb-post.csv data/raw/bdb-apps.csv
+...
+$(RAW_DATA): $(ACTIVATE_VENV) && python fetch_data.py
 ```
@@
 
-    player: pa.typing.Series[str]
+    player_season: pa.typing.Series[str]
     season: pa.typing.Series[str]
@@
 couple of data preparation functions: `prepare_data_2006` and
-`prepare_data_recent`:
+`prepare_data_bdb`:
 
@@
 or using `make analysis`. Now if we look in the file
-`data/prepared/recent/measurements.csv` we should see some lines that look like
+`data/prepared/bdb/measurements.csv` we should see some lines that look like
 this:
@@
 ```{csv}
-,player,season,n_attempt,n_success
+,player_season,season,n_attempt,n_success
 0,abreujo02,2018,553,180
@@
 
-# Extending the analysis to the recent data
+# Extending the analysis to the baseballdatabank data
 
@@
 folders with appropriate `prepared_data_dir` fields in their `config.toml`
-files. For example, here is the `config.toml` file for the `gparetorecent`
+files. For example, here is the `config.toml` file for the `gparetobdb`
 inference:
 
-```{.toml include="../inferences/gparetorecent/config.toml"}
+```{.toml include="../inferences/gparetobdb/config.toml"}
 ```
@@
 at-bats. This pattern was true both for the small 2006 dataset and the much
-larger 'recent' dataset.
+larger baseballdatabank dataset.

### 2023-02-20 bibat/examples/baseball/fetch_data.py

@@ -3,27 +3,28 @@
 import os
 
 import pandas as pd
-import requests
+
 
 URLS = {
     "2006": "https://raw.githubusercontent.com/stan-dev/"
     "example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
-    "recent": "https://discourse.mc-stan.org/uploads/short-url/"
-    "cRJtia2jKxiX01ZEraHUVK7jbHI.RData",
+    "bdb-main": "https://raw.githubusercontent.com/chadwickbureau/"
+    "baseballdatabank/master/core/Batting.csv",
+    "bdb-post": "https://raw.githubusercontent.com/chadwickbureau/"
+    "baseballdatabank/master/core/BattingPost.csv",
+    "bdb-apps": "https://raw.githubusercontent.com/chadwickbureau/"
+    "baseballdatabank/master/core/Appearances.csv"
 }
 OUT_FILES = {
-    "2006": os.path.join("data", "raw", "baseball-hits-2006.csv"),
-    "recent": os.path.join("data", "raw", "recent.RData"),
+    "2006": os.path.join("data", "raw", "2006.csv"),
+    "bdb-main": os.path.join("data", "raw", "bdb-main.csv"),
+    "bdb-post": os.path.join("data", "raw", "bdb-post.csv"),
+    "bdb-apps": os.path.join("data", "raw", "bdb-apps.csv"),
 }
 
 if __name__ == "__main__":
-    print(f"Fetching 2006 data from {URLS['2006']}")
-    data_2006 = pd.read_csv(URLS["2006"], comment="#")
-    print(f"Writing 2006 data to {OUT_FILES['2006']}")
-    data_2006.to_csv(OUT_FILES["2006"])
-    print(f"Fetching recent data from {URLS['recent']}")
-    with requests.get(URLS["recent"]) as f:
-        recent = f.text
-    print(f"Writing to {OUT_FILES['recent']}")
-    with open(OUT_FILES["recent"], "wb") as f:
-        f.write(recent.encode("utf8", errors="replace"))
+    for name, url in URLS.items():
+        print(f"Fetching {name} data from {url}")
+        data = pd.read_csv(url, comment="#")
+        print(f"Writing {name} data to {OUT_FILES[name]}")
+        data.to_csv(OUT_FILES[name])

### 2023-02-20 bibat/examples/baseball/prepare_data.py

@@ -8,15 +8,19 @@ from src.prepared_data import write_prepared_data
 
 DATA_PREPARATION_FUNCTIONS_TO_RUN = {
     "2006": data_preparation_functions.prepare_data_2006,
-    "recent": data_preparation_functions.prepare_data_recent,
+    "bdb": data_preparation_functions.prepare_data_bdb,
 }
 
 DATA_DIR = os.path.join(os.path.dirname(__file__), "data")
 RAW_DIR = os.path.join(DATA_DIR, "raw")
 PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
 RAW_DATA_FILES = {
-    "2006": os.path.join(RAW_DIR, "baseball-hits-2006.csv"),
-    "recent": os.path.join(RAW_DIR, "recent.csv"),
+    "2006": [os.path.join(RAW_DIR, "baseball-hits-2006.csv")],
+    "bdb": [
+        os.path.join(RAW_DIR, "bdb-main.csv"),
+        os.path.join(RAW_DIR, "bdb-post.csv"),
+        os.path.join(RAW_DIR, "bdb-apps.csv"),
+    ]
 }
 
 
@@ -24,12 +28,13 @@ def main():
     """Save prepared data in the PREPARED_DIR."""
     print("Reading raw data...")
     raw_data = {
-        k: pd.read_csv(v, index_col=None) for k, v in RAW_DATA_FILES.items()
+        k: [pd.read_csv(f, index_col=None) for f in v]
+        for k, v in RAW_DATA_FILES.items()
     }
     print("Preparing data...")
     for name, dpf in DATA_PREPARATION_FUNCTIONS_TO_RUN.items():
         print(f"Running data preparation function {dpf.__name__}...")
-        prepared_data = dpf(raw_data[name])
+        prepared_data = dpf(*raw_data[name])
         output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
         print(f"\twriting files to {output_dir}")
         if not os.path.exists(PREPARED_DIR):

### 2023-02-20 bibat/examples/baseball/src/data_preparation_functions.py

@@ -15,20 +15,24 @@ def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:
         columns={"K": "n_attempt", "y": "n_success"}
     ).assign(
         season="2006",
-        player=lambda df: [f"2006-player-{i+1}" for i in range(len(df))],
+        player_season=lambda df: [f"2006-player-{i+1}" for i in range(len(df))],
     )
     return PreparedData(
         name="2006",
         coords={
-            "player": measurements["player"].tolist(),
+            "player_season": measurements["player_season"].tolist(),
             "season": measurements["season"].tolist(),
         },
         measurements=measurements,
     )
 
 
-def prepare_data_recent(measurements_raw: pd.DataFrame) -> PreparedData:
-    """Prepare the recent data.
+def prepare_data_bdb(
+    measurements_main: pd.DataFrame,
+    measurements_post: pd.DataFrame,
+    appearances: pd.DataFrame,
+) -> PreparedData:
+    """Prepare the baseballdatabank data.
 
     There are a few substantive data choices here.
 
@@ -44,19 +48,23 @@ def prepare_data_recent(measurements_raw: pd.DataFrame) -> PreparedData:
     walks.
 
     """
+    pitchers = appearances.loc[
+        lambda df: df["G_p"] == df["G_all"], "playerID"
+    ].unique()
 
     def filter_batters(df: pd.DataFrame):
-        return (df["AB"] >= 20) & ~df["Pos.Summary"].astype(str).str.contains(
-            "1"
+        return (
+            (df["AB"] >= 20)
+            & (df["season"].ge(2017))
+            & (~df["player"].isin(pitchers))
         )
 
-    measurements = (
-        measurements_raw.assign(
-            season=lambda df: df["year"].astype(str),
-            player=lambda df: df["Name.additional"]
-            .str.cat(df["Tm"])
-            .str.cat(df["Lg"])
-            .str.cat(df["season"]),
+    measurements_main, measurements_post = (
+        m.rename(columns={"yearID": "season", "playerID": "player"})
+        .assign(
+            player_season=lambda df: df["player"].str.cat(
+                df["season"].astype(str)
+            ),
             n_attempt=lambda df: df[["AB", "BB", "HBP", "SF"]]
             .fillna(0)
             .sum(axis=1)
@@ -65,13 +73,23 @@ def prepare_data_recent(measurements_raw: pd.DataFrame) -> PreparedData:
[…]

### 2023-02-20 bibat/examples/baseball/src/prepared_data.py

@@ -19,7 +19,7 @@ class MeasurementsDF(pa.SchemaModel):
     Other columns are also allowed!
     """
 
-    player: pa.typing.Series[str]
+    player_season: pa.typing.Series[str]
     season: pa.typing.Series[str] = pa.Field(coerce=True)
     n_attempt: pa.typing.Series[int] = pa.Field(ge=1)
     n_success: pa.typing.Series[int] = pa.Field(ge=0)

### 2023-02-20 bibat/examples/baseball/src/stan_input_functions.py

@@ -26,7 +26,7 @@ def get_stan_input_gpareto(ppd: PreparedData) -> Dict:
         "N": len(ppd.measurements),
         "K": ppd.measurements["n_attempt"].values,
         "y": ppd.measurements["n_success"].values,
-        "min_alpha": logit(0.1),
+        "min_alpha": logit(0.07),
         "max_alpha": logit(0.5),
         "prior_sigma": [1.5, 0.4],
         "prior_k": [-0.5, 1],

### 2023-02-21 bibat/examples/baseball/docs/report.qmd
Update docs

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.0.7"
+subtitle: "A report created using bibat version 0.0.12"
 author: "Teddy Groves"
@@
 reference-location: margin
+toc: true
 format:
-  html: default
-  pdf: default
+  html:
+    embed-resources: true
 filters:
@@
 
-```{sh}
+```zsh
 $ pip install bibat
@@
 
-```{sh}
+```zsh
 bibat
@@
 
-```{sh}
+```zsh
 baseball
@@
 
-```{sh}
+```zsh
 $ cd baseball
@@
 
-```{sh}
+```zsh
 $ source .venv/bin/activate
@@
 
-```{Makefile}
-RAW_DATA = data/raw/2006.csv data/raw/bdb-main.csv data/raw/bdb-post.csv data/raw/bdb-apps.csv
+```makefile
+RAW_DATA = data/raw/2006.csv \
+	data/raw/bdb-main.csv \
+	data/raw/bdb-post.csv \
+	data/raw/bdb-apps.csv
+
 ...
-$(RAW_DATA): $(ACTIVATE_VENV) && python fetch_data.py
+$(RAW_DATA):
+	. $(ACTIVATE_VENV) && python fetch_data.py
 ```
@@
 
-```{sh}
+```zsh
 $ rm data/raw/raw_measurements.csv
@@
 
-```{python}
+```python
 from typing import Optional
@@
 
-```{sh}
+```zsh
 .
@@
 
-```{sh}
+```zsh
 inferences
@@
 distributions of the resulting `alpha` variables. I wanted to make sure that
-they all lay in the range corresponding to batting averages between about 0.15
+they all lay in the range corresponding to batting averages between about 0.1
 and a little over 0.4. Here is a graph that shows the 1% to 99% prior quantiles
-for each player's latent success percentage in the 2006 dataset alongside their
-actually realised success rate.
+for each player's latent success percentage in both datasets alongside their
+actually realised success rates.
 
-![](img/prior_quantiles.png)
+![](../plots/prior_quantiles.png)
 
@@
 
-![](img/posterior_quantiles.png)
+![](../plots/posterior_quantiles.png)
 

### 2023-02-21 bibat/examples/baseball/fetch_data.py

@@ -4,7 +4,6 @@ import os
 
 import pandas as pd
 
-
 URLS = {
     "2006": "https://raw.githubusercontent.com/stan-dev/"
     "example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv",
@@ -13,7 +12,7 @@ URLS = {
     "bdb-post": "https://raw.githubusercontent.com/chadwickbureau/"
     "baseballdatabank/master/core/BattingPost.csv",
     "bdb-apps": "https://raw.githubusercontent.com/chadwickbureau/"
-    "baseballdatabank/master/core/Appearances.csv"
+    "baseballdatabank/master/core/Appearances.csv",
 }
 OUT_FILES = {
     "2006": os.path.join("data", "raw", "2006.csv"),

### 2023-02-21 bibat/examples/baseball/prepare_data.py
Fix prepare data bug

@@ -15,12 +15,12 @@ DATA_DIR = os.path.join(os.path.dirname(__file__), "data")
 RAW_DIR = os.path.join(DATA_DIR, "raw")
 PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
 RAW_DATA_FILES = {
-    "2006": [os.path.join(RAW_DIR, "baseball-hits-2006.csv")],
+    "2006": [os.path.join(RAW_DIR, "2006.csv")],
     "bdb": [
         os.path.join(RAW_DIR, "bdb-main.csv"),
         os.path.join(RAW_DIR, "bdb-post.csv"),
         os.path.join(RAW_DIR, "bdb-apps.csv"),
-    ]
+    ],
 }
 
 

### 2023-04-27 bibat/examples/baseball/docs/report.qmd
Update vignette

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.0.12"
+subtitle: "A report created using bibat version 0.1.3"
 author: "Teddy Groves"
@@
 ```zsh
-$ pip install bibat
+> pip install bibat
 ```
 
-Next I started bibat's wizard like this:
+Next I ran bibat's wizard like this:
 
-```zsh
-bibat
+```{.zsh}
+> bibat
+Welcome to the Batteries-Included Bayesian Analysis Template!
+What is your project called? [Project Name]: Baseball
+What should the project repository be called [baseball]: baseball
+What is your name? [Author name]: Teddy Groves
+What is your email? [author@email.com]:
+Who should be the code of conduct contact? [author@email.com]:
+Please briefly describe your project [A short description of the project.]: Comparison of distributions for modelling baseball hitting
+Choose an open source license from these options: (MIT, BSD-3-Clause, No license file) [MIT]:
+How would you like to document your project? (Quarto, Sphinx, No docs) [Quarto]:
+Would you like to create a tests directory? [y]: n
+Would you like to create a .github directory? [y]: n
+> 
 ```
@@
 
-```zsh
+```{.zsh}
 baseball
@@
 ├── README.md
+├── baseball
+│   ├── __init__.py
+│   ├── data_preparation.py
+│   ├── inference_configuration.py
+│   ├── investigate.ipynb
+│   ├── notebooks
+│   ├── prepare_data.py
+│   ├── readme.md
+│   ├── sample.py
+│   ├── stan
+│   │   ├── custom_functions.stan
+│   │   ├── multilevel-linear-regression.stan
+│   │   └── readme.md
+│   ├── stan_input_functions.py
+│   └── util.py
 ├── bibat_version.txt
@@
 │       └── config.toml
-├── investigate.ipynb
 ├── plots
@@
 │   └── posterior_predictive_comparison.png
-├── prepare_data.py
 ├── pyproject.toml
-├── requirements-tooling.txt
-├── requirements.txt
-├── sample.py
-├── src
-│   ├── __init__.py
-│   ├── data_preparation_functions.py
-│   ├── inference_configuration.py
-│   ├── prepared_data.py
-│   ├── readme.md
-│   ├── stan
-│   │   ├── custom_functions.stan
-│   │   ├── multilevel-linear-regression.stan
-│   │   └── readme.md
-│   ├── stan_input_functions.py
-│   └── util.py
-└── tests
-    ├── test_integration
-    │   └── test_data_preparation.py
-    └── test_unit
-        ├── test_inference_configuration.py
-        └── test_util.py
+└── requirements.txt
 ```
@@
 
-Some cogs turned, some new lines appeared in my terminal and some new files were
-created - nice, the setup worked!
+This ran without errors, running some data preparation functions and then analysing the data with Stan in a few configurations.
 
@@
 
-To fetch raw data from the internet, I wrote a new script called `fetch_data.py`:
+To fetch raw data from the internet, I wrote a new script in the file
+`baseball/fetch_data.py`:
 
-```{.python include="../fetch_data.py"}
+```{.python include="../baseball/fetch_data.py"}
 ```
@@
 ```zsh
-$ source .venv/bin/activate
-$ python fetch_data.py
-```
-
-Since this worked, I added a new makefile target for the raw data files:
-
-```makefile
-RAW_DATA = data/raw/2006.csv \
-	data/raw/bdb-main.csv \
-	data/raw/bdb-post.csv \
-	data/raw/bdb-apps.csv
-
-...
-$(RAW_DATA):
-	. $(ACTIVATE_VENV) && python fetch_data.py
+> source .venv/bin/activate
+(baseball) > python baseball/fetch_data.py
+Fetching 2006 data from https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv
+Writing 2006 data to data/raw/2006.csv
+Fetching bdb-main data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Batting.csv
+Writing bdb-main data to data/raw/bdb-main.csv
+Fetching bdb-post data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/BattingPost.csv
+Writing bdb-post data to data/raw/bdb-post.csv
+Fetching bdb-apps data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Appearances.csv
+Writing bdb-apps data to data/raw/bdb-apps.csv
 ```
@@
 ```zsh
-$ rm data/raw/raw_measurements.csv
+(baseball) > rm data/raw/raw_measurements.csv
 ```
@@
 # Preparing the data
+
 The first step in preparing data is to decide what prepared data looks like for
 the purposes of our analysis. Bibat provides dataclasses called `PreparedData`
-and `MeasurementsDF ` in the file `src/prepared_data.py` which can help get us
-started with this.
+and `MeasurementsDF` to help get started with this, which I found in the file
+`baseball/prepared_data.py`.
 
-As it happens, prepared data looks very similar in our analysis and the example.
-All we need to do is change the `MeasurementsDF` definition a little^[note that
-this class uses [pandera](https://pandera.readthedocs.io), a handy library for
-defining what a pandas dataframe should look like]:
+As it happens, prepared data looks very similar in this analysis and the
+example. All I had to do was change the `MeasurementsDF` definition a
+little^[note that this class uses [pandera](https://pandera.readthedocs.io), a
+handy library for defining what a pandas dataframe should look like]:
 
 ```python
-from typing import Optional
-import pandera as pa
-
-# ...
-
 class MeasurementsDF(pa.SchemaModel):
@@
 
-    player_season: pa.typing.Series[str]
-    season: pa.typing.Series[str]
-    n_attempt: pa.typing.Series[int] = pa.Field(ge = 1)
-    n_success: pa.typing.Series[int] = pa.Field(ge = 0)
+    player_season: Series[str]
+    season: Series[str] = pa.Field(coerce=True)
+    n_attempt: Series[int] = pa.Field(ge = 1)
+    n_success: Series[int] = pa.Field(ge = 0)
 ```
 
-Next we can write some functions that create `PreparedData` objects. These
-should live in the file `data_preparation_functions.py`. In this case we write a
-couple of data preparation functions: `prepare_data_2006` and
+The next step is to write functions that return `PreparedData` objects. In this
+case I wrote a couple of data preparation functions: `prepare_data_2006` and
 `prepare_data_bdb`:
 
-```{.python include="../src/data_preparation_functions.py"}
+```{.python include="../baseball/data_preparation.py" start-line=125}
 ```
 
-Finally we need to update `prepare_data.py`, the script that runs the data
-preparation functions. Again there isn't much to change from the example
-analysis.
+To check that all this worked, I ran the data preparation script manually:^[I
+could also have just run `make analysis` again]
 
-```{.python include="../prepare_data.py"}
+```{.zsh}
+> source .venv/bin/activate
+(baseball) > python baseball/prepare_data.py
 ```
 
-To check that all this works, we can run the script `prepare_data.py` manually
-or using `make analysis`. Now if we look in the file
-`data/prepared/bdb/measurements.csv` we should see some lines that look like
-this:
+Now the folder `data/prepared/bdb` contained a file
+`data/prepared/bdb/measurements.csv` that looked like this:
 
@@
 
-```{stan}
+```{.stan}
 real gpareto_lpdf(vector y, real ymin, real k, real sigma) {
@@
 
-```{.stan include="../src/stan/gpareto.stan"}
+```{.stan include="../baseball/stan/gpareto.stan"}
 ```
@@
 
-```{.stan include="../src/stan/normal.stan"}
+```{.stan include="../baseball/stan/normal.stan"}
 ```
@@
 # Generating Stan inputs
-Next we need to tell our analysis how to turn some prepared data into a
+
+Next I needed to tell the analysis how to turn some prepared data into a
 dictionary that can be used as input for Stan. Bibat assumes that this task is
-handled by functions that live in the file `src/stan_input_functions.py`, each
-of which takes in a `PreparedData` and returns a Python dictionary. You can
-write as many Stan input functions as you like and choose which one to run for
-any given inference.
+handled by functions that live in the file `baseball/stan_input_functions.py`,
+each of which takes in a `PreparedData` and returns a Python dictionary. You
+can write as many Stan input functions as you like and choose which one to run
+for any given inference.
 
-We can start by defining some Stan input functions that pass arbitary prepared data
-on to each of the models:
+I started by defining some Stan input functions that pass arbitary prepared
+data on to each of the models:
 
-```{.python include="../src/stan_input_functions.py" start-line=12 end-line=34}
+```{.python include="../baseball/stan_input_functions.py" start-line=11 end-line=34}
 ```
 
-But why stop there? It can also be useful to generate Stan inputs consistently
-with a model, based on some hardcoded hyperparameter values. Here are some
-functions that do this for both of our models:
+But why stop there? It can also be useful to generate Stan input data from a
+model with a model, by running it in simulation model with hardcoded parameter
+values. Here are some functions that do this for both of our models:
 
-```{.python include="../src/stan_input_functions.py" start-line=37}
+```{.python include="../baseball/stan_input_functions.py" start-line=36}
 ```
@@
 # Specifying inferences
+
 Now all the building blocks for making statistical inferences - raw data, data
-preparation rules, statistical models and recipes for turning prepared data into
-model inputs - are in place. The last step before actually running Stan is to
-write down how put these blocks together. Bibat has another concept for this,
-called 'inferences'.
+preparation rules, statistical models and recipes for turning prepared data
+into model inputs - were in place. The last step before actually running Stan
+was to write down how put the blocks together. Bibat has another concept for
+this, called 'inferences'.
 
@@
 ```zsh
-.
+inferences
 ├── gpareto2006
@@
 # Investigating the inferences
+
 Now that the inferences are ready it's time to check them out. Bibat provides a
-jupyter notebook called `investigate.ipynb` for exactly this purpose.
+jupyter notebook at `baseball/investigate.ipynb` for exactly this purpose.
 
@@
 After running `make analysis` one more time, I went back to the notebook
-`investigate.ipynb` and made plots of both models' posterior 1%-99% success rate
-intervals for both datasets:
+`baseball/investigate.ipynb` and made plots of both models' posterior 1%-99%
+success rate intervals for both datasets:
 

### 2023-04-27 bibat/examples/baseball/baseball/data_preparation.py

@@ -1,0 +1,210 @@
+"""Provides functions prepare_data_x.
+
+These functions should take in a dataframe of measurements and return a
+PreparedData object.
+
+"""
+import json
+import os
+
+import numpy as np
+import pandas as pd
+import pandera as pa
+from pandera.typing import DataFrame, Series
+from pydantic.dataclasses import dataclass
+from baseball.util import CoordDict, make_columns_lower_case
+
+NAME_FILE = "name.txt"
+COORDS_FILE = "coords.json"
+MEASUREMENTS_FILE = "measurements.csv"
+NEW_COLNAMES = {"yButIThoughtIdAddSomeLetters": "y"}
+DROPNA_COLS = ["y"]
+N_CV_FOLDS = 10
+DIMS = {
+    "b": ["covariate"],
+    "y": ["observation"],
+    "yrep": ["observation"],
+    "llik": ["observation"],
+}
+
+HERE = os.path.dirname(__file__)
+DATA_DIR = os.path.join(HERE, "..", "data")
+RAW_DIR = os.path.join(DATA_DIR, "raw")
+PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
+RAW_DATA_FILES = {
+    "2006": [os.path.join(RAW_DIR, "2006.csv")],
+    "bdb": [
+        os.path.join(RAW_DIR, "bdb-main.csv"),
+        os.path.join(RAW_DIR, "bdb-post.csv"),
+        os.path.join(RAW_DIR, "bdb-apps.csv"),
+    ],
+}
+
+
+def prepare_data():
+    """Main function."""
+    print("Reading raw data...")
+    raw_data = {
+        k: [pd.read_csv(file, index_col=None) for file in v] 
+        for k, v in RAW_DATA_FILES.items()
+    }
+    data_preparation_functions_to_run = {
+        "2006": prepare_data_2006, "bdb": prepare_data_bdb
+    }
+    print("Preparing data...")
+    for name, dpf in data_preparation_functions_to_run.items():
+        print(f"Running data preparation function {dpf.__name__}...")
+        prepared_data = dpf(*raw_data[name])
+        output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
+        print(f"\twriting files to {output_dir}")
+        if not os.path.exists(PREPARED_DIR):
+            os.mkdir(PREPARED_DIR)
+        write_prepared_data(prepared_data, output_dir)
+
+
+class MeasurementsDF(pa.SchemaModel):
[…]

### 2023-04-27 bibat/examples/baseball/baseball/inference_configuration.py

@@ -5,7 +5,7 @@ from typing import Callable, Dict, List, Optional
 
 import toml
 from pydantic import BaseModel, Field, root_validator, validator
-from src import stan_input_functions
+from baseball import stan_input_functions
 
 AVAILABLE_MODES = ["prior", "posterior", "kfold"]
 HERE = os.path.dirname(os.path.abspath(__file__))
@@ -78,8 +78,8 @@ class InferenceConfiguration(BaseModel):
     @validator("stan_file")
     def check_stan_file_exists(cls, v):
         """Check that the stan file exists."""
-        if not os.path.exists(os.path.join("src", "stan", v)):
-            raise ValueError(f"{v} is not a file in src/stan.")
+        if not os.path.exists(os.path.join(STAN_DIR, v)):
+            raise ValueError(f"{v} is not a file in {STAN_DIR}.")
         return v
 
     @validator("modes")

### 2023-04-27 bibat/examples/baseball/baseball/sample.py

@@ -8,16 +8,17 @@ import cmdstanpy
 import numpy as np
 import xarray as xr
 from sklearn.model_selection import KFold
-from src.inference_configuration import (
+from baseball.data_preparation import load_prepared_data
+from baseball.inference_configuration import (
     AVAILABLE_MODES,
     InferenceConfiguration,
     load_inference_configuration,
 )
-from src.prepared_data import load_prepared_data
-from src.util import CoordDict
+from baseball.util import CoordDict
 
-RUNS_DIR = "inferences"
-STAN_DIR = os.path.join("src", "stan")
+HERE = os.path.dirname(__file__)
+RUNS_DIR = os.path.join(HERE, "..", "inferences")
+STAN_DIR = os.path.join(HERE, "stan")
 
 
 def get_sample_kwargs(raw: Dict, mode: str) -> Dict:

### 2023-04-27 bibat/examples/baseball/baseball/stan_input_functions.py

@@ -5,7 +5,7 @@ from typing import Dict
 
 import numpy as np
 from scipy.special import expit, logit
-from src.prepared_data import PreparedData
+from baseball.data_preparation import PreparedData
 
 
 def get_stan_input_normal(ppd: PreparedData) -> Dict:

### 2023-04-27 bibat/examples/baseball/prepare_data.py

@@ -1,46 +1,0 @@
-"""Read the data in RAW_DIR and save prepared data to PREPARED_DIR."""
-
-import os
-
-import pandas as pd
-from src import data_preparation_functions
-from src.prepared_data import write_prepared_data
-
-DATA_PREPARATION_FUNCTIONS_TO_RUN = {
-    "2006": data_preparation_functions.prepare_data_2006,
-    "bdb": data_preparation_functions.prepare_data_bdb,
-}
-
-DATA_DIR = os.path.join(os.path.dirname(__file__), "data")
-RAW_DIR = os.path.join(DATA_DIR, "raw")
-PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
-RAW_DATA_FILES = {
-    "2006": [os.path.join(RAW_DIR, "2006.csv")],
-    "bdb": [
-        os.path.join(RAW_DIR, "bdb-main.csv"),
-        os.path.join(RAW_DIR, "bdb-post.csv"),
-        os.path.join(RAW_DIR, "bdb-apps.csv"),
-    ],
-}
-
-
-def main():
-    """Save prepared data in the PREPARED_DIR."""
-    print("Reading raw data...")
-    raw_data = {
-        k: [pd.read_csv(f, index_col=None) for f in v]
-        for k, v in RAW_DATA_FILES.items()
-    }
-    print("Preparing data...")
-    for name, dpf in DATA_PREPARATION_FUNCTIONS_TO_RUN.items():
-        print(f"Running data preparation function {dpf.__name__}...")
-        prepared_data = dpf(*raw_data[name])
-        output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
-        print(f"\twriting files to {output_dir}")
-        if not os.path.exists(PREPARED_DIR):
-            os.mkdir(PREPARED_DIR)
-        write_prepared_data(prepared_data, output_dir)
-
-
-if __name__ == "__main__":
-    main()

### 2023-04-27 bibat/examples/baseball/src/data_preparation_functions.py

@@ -1,96 +1,0 @@
-"""Provides functions prepare_data_x.
-
-These functions should take in a dataframe of measurements and return a
-PreparedData object.
-
-"""
-
-import pandas as pd
-from src.prepared_data import PreparedData
-
-
-def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:
-    """Prepare the 2006 data."""
-    measurements = measurements_raw.rename(
-        columns={"K": "n_attempt", "y": "n_success"}
-    ).assign(
-        season="2006",
-        player_season=lambda df: [f"2006-player-{i+1}" for i in range(len(df))],
-    )
-    return PreparedData(
-        name="2006",
-        coords={
-            "player_season": measurements["player_season"].tolist(),
-            "season": measurements["season"].tolist(),
-        },
-        measurements=measurements,
-    )
-
-
-def prepare_data_bdb(
-    measurements_main: pd.DataFrame,
-    measurements_post: pd.DataFrame,
-    appearances: pd.DataFrame,
-) -> PreparedData:
-    """Prepare the baseballdatabank data.
-
-    There are a few substantive data choices here.
-
-    First, the function excludes players who have a '1' in their position as
-    these are likely pitchers, as well as players with fewer than 20 at bats.
-
-    Second, the function defines a successes and attempts according to the
-    'on-base percentage' metric, so a success is a time when a player got a hit,
-    a base on ball/walk or a hit-by-pitch and an attempt is an at-bat or a
-    base-on-ball/walk or a hit-by-pitch or a sacrifice fly. This could have
-    alternatively been calculated as just hits divided by at-bats, but my
-    understanding is that this method underrates players who are good at getting
-    walks.
-
-    """
-    pitchers = appearances.loc[
-        lambda df: df["G_p"] == df["G_all"], "playerID"
-    ].unique()
-
-    def filter_batters(df: pd.DataFrame):
-        return (
-            (df["AB"] >= 20)
-            & (df["season"].ge(2017))
-            & (~df["player"].isin(pitchers))
-        )
-
-    measurements_main, measurements_post = (
-        m.rename(columns={"yearID": "season", "playerID": "player"})
-        .assign(
-            player_season=lambda df: df["player"].str.cat(
-                df["season"].astype(str)
-            ),
-            n_attempt=lambda df: df[["AB", "BB", "HBP", "SF"]]
-            .fillna(0)
-            .sum(axis=1)
-            .astype(int),
-            n_success=lambda df: (
[…]

### 2023-04-27 bibat/examples/baseball/baseball/data_preparation.py
qa + update readme

@@ -10,9 +10,9 @@ import os
 import numpy as np
 import pandas as pd
 import pandera as pa
+from baseball.util import CoordDict, make_columns_lower_case
 from pandera.typing import DataFrame, Series
 from pydantic.dataclasses import dataclass
-from baseball.util import CoordDict, make_columns_lower_case
 
 NAME_FILE = "name.txt"
 COORDS_FILE = "coords.json"
@@ -45,11 +45,12 @@ def prepare_data():
     """Main function."""
     print("Reading raw data...")
     raw_data = {
-        k: [pd.read_csv(file, index_col=None) for file in v] 
+        k: [pd.read_csv(file, index_col=None) for file in v]
         for k, v in RAW_DATA_FILES.items()
     }
     data_preparation_functions_to_run = {
-        "2006": prepare_data_2006, "bdb": prepare_data_bdb
+        "2006": prepare_data_2006,
+        "bdb": prepare_data_bdb,
     }
     print("Preparing data...")
     for name, dpf in data_preparation_functions_to_run.items():
@@ -106,6 +107,7 @@ def write_prepared_data(prepped: PreparedData, directory):
         json.dump(prepped.coords, f)
     with open(os.path.join(directory, NAME_FILE), "w") as f:
         f.write(prepped.name)
+
 
 def prepare_data_interaction(measurements_raw: pd.DataFrame) -> PreparedData:
     """Prepare data with an interaction column."""
@@ -207,4 +209,3 @@ def prepare_data_bdb(
         },
         measurements=measurements,
     )
-

### 2023-04-27 bibat/examples/baseball/baseball/inference_configuration.py

@@ -4,8 +4,8 @@ import os
 from typing import Callable, Dict, List, Optional
 
 import toml
-from pydantic import BaseModel, Field, root_validator, validator
 from baseball import stan_input_functions
+from pydantic import BaseModel, Field, root_validator, validator
 
 AVAILABLE_MODES = ["prior", "posterior", "kfold"]
 HERE = os.path.dirname(os.path.abspath(__file__))

### 2023-04-27 bibat/examples/baseball/baseball/sample.py

@@ -7,7 +7,6 @@ import arviz as az
 import cmdstanpy
 import numpy as np
 import xarray as xr
-from sklearn.model_selection import KFold
 from baseball.data_preparation import load_prepared_data
 from baseball.inference_configuration import (
     AVAILABLE_MODES,
@@ -15,6 +14,7 @@ from baseball.inference_configuration import (
     load_inference_configuration,
 )
 from baseball.util import CoordDict
+from sklearn.model_selection import KFold
 
 HERE = os.path.dirname(__file__)
 RUNS_DIR = os.path.join(HERE, "..", "inferences")

### 2023-04-27 bibat/examples/baseball/baseball/stan_input_functions.py

@@ -4,8 +4,8 @@
 from typing import Dict
 
 import numpy as np
-from scipy.special import expit, logit
 from baseball.data_preparation import PreparedData
+from scipy.special import expit, logit
 
 
 def get_stan_input_normal(ppd: PreparedData) -> Dict:

### 2023-04-27 bibat/examples/baseball/baseball/data_preparation.py
fix flake8

@@ -1,4 +1,4 @@
-"""Provides functions prepare_data_x.
+"""Provide functions prepare_data_x.
 
 These functions should take in a dataframe of measurements and return a
 PreparedData object.
@@ -7,10 +7,9 @@ PreparedData object.
 import json
 import os
 
-import numpy as np
 import pandas as pd
 import pandera as pa
-from baseball.util import CoordDict, make_columns_lower_case
+from baseball.util import CoordDict
 from pandera.typing import DataFrame, Series
 from pydantic.dataclasses import dataclass
 
@@ -42,7 +41,7 @@ RAW_DATA_FILES = {
 
 
 def prepare_data():
-    """Main function."""
+    """Run main function."""
     print("Reading raw data...")
     raw_data = {
         k: [pd.read_csv(file, index_col=None) for file in v]
@@ -107,21 +106,6 @@ def write_prepared_data(prepped: PreparedData, directory):
         json.dump(prepped.coords, f)
     with open(os.path.join(directory, NAME_FILE), "w") as f:
         f.write(prepped.name)
-
-
-def prepare_data_interaction(measurements_raw: pd.DataFrame) -> PreparedData:
-    """Prepare data with an interaction column."""
-    measurements = process_measurements(measurements_raw)
-    return PreparedData(
-        name="interaction",
-        coords=CoordDict(
-            {
-                "covariate": ["x1", "x2", "x1:x2"],
-                "observation": measurements.index.tolist(),
-            }
-        ),
-        measurements=measurements,
-    )
 
 
 def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:

### 2023-04-27 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/data_preparation.py

@@ -37,7 +37,7 @@ RAW_DATA_FILES = {
 
 
 def prepare_data():
-    """Main function."""
+    """Run main function."""
     print("Reading raw data...")
     raw_data = {
         k: pd.read_csv(v, index_col=None) for k, v in RAW_DATA_FILES.items()

### 2023-04-27 docs/conf.py
bump version

@@ -24,7 +24,7 @@ copyright = "2022, Teddy Groves"
 author = "Teddy Groves"
 
 # The full version, including alpha/beta/rc tags
-release = "0.1.3"
+release = "0.1.4"
 
 root_doc = "index"
 

### 2023-05-04 bibat/{{cookiecutter.repo_name}}/docs/report.qmd

@@
 title: "{{ cookiecutter.project_name }}"
+resource_path: ["../"]
 subtitle: "A report created using bibat version {{ cookiecutter.bibat_version }}"
@@
 
-Turn it into a pdf by running `quarto render report.qmd` from the project root.
+Create Pdf and Html files based on this document by running `quarto render
+report.qmd` from the project root. If you want to remove some of these formats
+or add a new one, you can do so by editing the `format` field of the
+document's yaml front matter. For example if you want to create a version that
+your collaborators can open in Microsoft Word:
+
+```{.yaml}
+---
+format:
+  html: default
+  pdf: default
+  docx: default
+---
+```
 
@@
 
+Note that you can also include links to images in the main project's `plots`
+folder:
+
+![This image comes from the main analysis!](../plots/posterior_predictive_comparison.png)
+
 Include margin notes like this^[I'm a margin note!]
 
-Write inline code like this:
+Write code blocks like this:
 
@@
     print("Hello!")
+print_hello()
+```
+
+Note that when quarto rendered the document it evaluated the code above and
+printed the results! Quarto supports execution of Python, R and Julia code, and
+can print figures that or even interactive widgets. If you want to include a
+block of code from one of these languages, but don't want quarto to execute it,
+or want to [customise what quarto does with the code in another
+way](https://quarto.org/docs/computations/execution-options.html) make sure the
+first line is `#| `, e.g.:
+
+```{python}
+#| eval: false
+def print_hello():
+    print("Hello!")
+print_hello()
 ```

### 2023-05-23 bibat/examples/baseball/baseball/data_preparation.py
delete unused global variables in data preparation file

@@ -9,23 +9,14 @@ import os
 
 import pandas as pd
 import pandera as pa
-from baseball.util import CoordDict
 from pandera.typing import DataFrame, Series
 from pydantic.dataclasses import dataclass
 
+from baseball.util import CoordDict
+
 NAME_FILE = "name.txt"
 COORDS_FILE = "coords.json"
-MEASUREMENTS_FILE = "measurements.csv"
-NEW_COLNAMES = {"yButIThoughtIdAddSomeLetters": "y"}
-DROPNA_COLS = ["y"]
 N_CV_FOLDS = 10
-DIMS = {
-    "b": ["covariate"],
-    "y": ["observation"],
-    "yrep": ["observation"],
-    "llik": ["observation"],
-}
-
 HERE = os.path.dirname(__file__)
 DATA_DIR = os.path.join(HERE, "..", "data")
 RAW_DIR = os.path.join(DATA_DIR, "raw")

### 2023-05-23 bibat/examples/baseball/docs/report.qmd
Update vignette

@@
 and `MeasurementsDF` to help get started with this, which I found in the file
-`baseball/prepared_data.py`.
+`baseball/data_preparation.py`.
 
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=125}
+```{.python include="../baseball/data_preparation.py" start-line=101}
 ```
+
+I had to update the `prepare_data` function to take into account the
+two raw data sources:
+
+```{.python include="../baseball/data_preparation.py" start-line=33 end-line=52}
+```
+
+To finish off I deleted the unused global variables `MEASUREMENTS_FILE`,
+`NEW_COLNAMES`, `DROPNA_COLS` and `DIMS`, then checked if the function
+`load_prepared_data` needed any changes: I was pretty sure it didn't.
 
@@
 I started by defining some Stan input functions that pass arbitary prepared
-data on to each of the models:
+data on to each of the models:^[Note that this code uses the scipy function
+`logit`, which it imported like this: `from scipy.special import logit`]
 
@@
 Note that:
-* The Stan file, prepared data folder and stan input function are referred to by
+
+- The Stan file, prepared data folder and stan input function are referred to by
   strings. The analysis should raise an error if you enter a non-existing value.
-* Both inferences are set to run in "prior" and "posterior" modes - the other
+- Both inferences are set to run in "prior" and "posterior" modes - the other
   pre-defined mode is "kfold", but you can also write your own!
-* You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
+- You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
   in the `[sample_kwargs]` table.
-* You can enter mode-specific overrides in `[sample_kwargs.<MODE>]`. This can be
+- You can enter mode-specific overrides in `[sample_kwargs.<MODE>]`. This can be
   handy if you want to run more or fewer iterations for a certain mode.
@@
 Now that the inferences are ready it's time to check them out. Bibat provides a
-jupyter notebook at `baseball/investigate.ipynb` for exactly this purpose.
+Jupyter notebook at `baseball/investigate.ipynb` for exactly this purpose. The
+notebook's main job is to create plots and save them in the `plots` directory
+when it is executed with the command `jupyter execute investigate.ipynb`, which
+is the final step in the chain of commands that is triggered by `make
+analysis`.
 
-A lot of code from the example analysis's notebook was reusable, so I largely
-followed its structure, with a few tweaks.
+A notebook is arguably a nicer home for code that creates plots than a plain
+python script because it allows for literate documentation and an iterative
+workflow. A notebook makes it easy to, for example, add some code to change the
+scale of a plot, execute the code and see the new results, then update the
+relevant documentation all in the same place.
+
+The code from the example analysis's notebook for loading `InferenceData` was
+reusable with a few tweaks to avoid missing file errors, so I kept it. On the
+other hand, I wanted to make some different plots from the ones in the example
+analysis, including some that required loading prepared data. To check out
+everything I did, see
+[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/baseball/investigate.ipynb).
 
@@
 larger baseballdatabank dataset.
+
+# Documenting the analysis
+
+The final step was to document my analysis. To do this I edited the file
+`docs/report.qmd`, then ran `quarto render docs/report.qmd`, which produced the
+very html document that you are probably reading now! You can find the complete
+`report.qmd` file
+[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/docs/report.qmd).

### 2023-05-23 bibat/examples/baseball/baseball/data_preparation.py
Fix isort first party package identification

@@ -16,6 +16,7 @@ from baseball.util import CoordDict
 
 NAME_FILE = "name.txt"
 COORDS_FILE = "coords.json"
+MEASUREMENTS_FILE = "measurements.csv"
 N_CV_FOLDS = 10
 HERE = os.path.dirname(__file__)
 DATA_DIR = os.path.join(HERE, "..", "data")

### 2023-05-23 bibat/examples/baseball/baseball/sample.py

@@ -7,6 +7,8 @@ import arviz as az
 import cmdstanpy
 import numpy as np
 import xarray as xr
+from sklearn.model_selection import KFold
+
 from baseball.data_preparation import load_prepared_data
 from baseball.inference_configuration import (
     AVAILABLE_MODES,
@@ -14,7 +16,6 @@ from baseball.inference_configuration import (
     load_inference_configuration,
 )
 from baseball.util import CoordDict
-from sklearn.model_selection import KFold
 
 HERE = os.path.dirname(__file__)
 RUNS_DIR = os.path.join(HERE, "..", "inferences")

### 2023-05-23 bibat/examples/baseball/docs/report.qmd
Fix mistakes in baseball vignette

@@
 complete with [data from the 2006 Major league
-season])https://github.com/stan-dev/example-models/blob/master/knitr/pool-binary-trials/baseball-hits-2006.csv).
+season](https://github.com/stan-dev/example-models/blob/master/knitr/pool-binary-trials/baseball-hits-2006.csv).
 After I posted some early results [on the Stan discourse
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=101}
+```{.python include="../baseball/data_preparation.py" start-line=103}
 ```
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=33 end-line=52}
+```{.python include="../baseball/data_preparation.py" start-line=35 end-line=53}
 ```
 
-To finish off I deleted the unused global variables `MEASUREMENTS_FILE`,
-`NEW_COLNAMES`, `DROPNA_COLS` and `DIMS`, then checked if the function
-`load_prepared_data` needed any changes: I was pretty sure it didn't.
+To finish off I deleted the unused global variables `NEW_COLNAMES`,
+`DROPNA_COLS` and `DIMS`, then checked if the function `load_prepared_data`
+needed any changes: I was pretty sure it didn't.
 

### 2023-06-02 bibat/examples/baseball/docs/report.qmd
Fix include in baseball report
update baseball example

@@
 title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.1.3"
+subtitle: "A report created using bibat version 0.1.6"
 author: "Teddy Groves"
@@
 Still, there were a few problems. In particular the data were typically
-unbalanced because better players tended to fproduce more data than worse
+unbalanced because better players tended to produce more data than worse
 players. The result of this was that my models would often inappropriately think
@@
 ├── README.md
+├── bibat_version.txt
+├── data
+│   └── raw
+│       ├── raw_measurements.csv
+│       └── readme.md
+├── docs
+│   ├── bibliography.bib
+│   ├── img
+│   │   ├── example.png
+│   │   └── readme.md
+│   ├── report.html
+│   ├── report.pdf
+│   └── report.qmd
 ├── baseball
@@
 │   ├── data_preparation.py
+│   ├── fitting_mode.py
 │   ├── inference_configuration.py
@@
 │   └── util.py
-├── bibat_version.txt
-├── data
-│   └── raw
-│       ├── raw_measurements.csv
-│       └── readme.md
-├── docs
-│   ├── bibliography.bib
-│   ├── img
-│   │   ├── example.png
-│   │   └── readme.md
-│   └── report.qmd
 ├── inferences
@@
 │   └── posterior_predictive_comparison.png
-├── pyproject.toml
-└── requirements.txt
+└── pyproject.toml
 ```
 
-This folder implements bibat's example analysis - a comparison of linear
+This folder implements bibat's example analysis – a comparison of linear
 regression with two different design matrices. To check that everything was
@@
 
-Finally I removed the example analysis's raw data:
+Finally, I removed the example analysis's raw data:
 
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=35 end-line=53}
+```{.python include="../baseball/data_preparation.py" start-line=35 end-line=54}
 ```
@@
 To check that all this worked, I ran the data preparation script manually:^[I
-could also have just run `make analysis` again]
+could also have just run `make analysis` again. This would have caused an error
+on the step after prepare_data.py, which is fine!]
 
@@
 I wanted to test two statistical models: one with the modelling the distribution
-of per-player logit-scale at-bat success rates as a normal distbution with
+of per-player logit-scale at-bat success rates as a normal distribution with
 unknown mean and standard deviation, and another where the same logit-scale
-rates have a generalised pareto distribution.
+rates have a generalised Pareto distribution.
 
-So, genve a table of $N$ player profiles, with each player has $y$ successes out
+So, given a table of $N$ player profiles, with each player has $y$ successes out
 of $K$ at-bats and an unknown latent success rate $\alpha$, I wanted to use this
@@
 
-In the generalised pareto model I would give the $\alpha$s this prior model,
+In the generalised Pareto model I would give the $\alpha$s this prior model,
 with the hyperparameter $\text{min }\alpha$ assumed to be known exactly and $k$
@@
 
-I started by defining some Stan input functions that pass arbitary prepared
+I started by defining some Stan input functions that pass arbitrary prepared
 data on to each of the models:^[Note that this code uses the scipy function
@@
 
-But why stop there? It can also be useful to generate Stan input data from a
-model with a model, by running it in simulation model with hardcoded parameter
-values. Here are some functions that do this for both of our models:
+But why stop there? It can also be useful to generate Stan input data with a
+model, by running it in simulation mode with hardcoded parameter values. Here
+are some functions that do this for both of our models:
 
@@
 
-Now all the building blocks for making statistical inferences - raw data, data
+Now all the building blocks for making statistical inferences – raw data, data
 preparation rules, statistical models and recipes for turning prepared data
-into model inputs - were in place. The last step before actually running Stan
+into model inputs – were in place. The last step before actually running Stan
 was to write down how put the blocks together. Bibat has another concept for
@@
 
-- The Stan file, prepared data folder and stan input function are referred to by
+- The Stan file, prepared data folder and Stan input function are referred to by
   strings. The analysis should raise an error if you enter a non-existing value.
-- Both inferences are set to run in "prior" and "posterior" modes - the other
+- Both inferences are set to run in "prior" and "posterior" modes – the other
   pre-defined mode is "kfold", but you can also write your own!
@@
   in the `[sample_kwargs]` table.
-- You can enter mode-specific overrides in `[sample_kwargs.<MODE>]`. This can be
+- You can enter mode-specific overrides in `[mode_options.<MODE>]`. This can be
   handy if you want to run more or fewer iterations for a certain mode.
 
-Now when I ran `make analysis` again, I saw messages indicating that Stan had
-run, and found that the `inferences` subfolders had been populated:
+With all these changes made I ran `make analysis` again. There were no errors
+until after sampling, which was expected as I hadn't yet customised the
+investigation code, and I saw messages indicating that Stan had run. I also
+found that the `inferences` subfolders had been populated:
 
@@
 `docs/report.qmd`, then ran `quarto render docs/report.qmd`, which produced the
-very html document that you are probably reading now! You can find the complete
+very HTML document that you are probably reading now! You can find the complete
 `report.qmd` file

### 2023-06-02 bibat/examples/baseball/baseball/fitting_mode.py

@@ -1,0 +1,143 @@
+"""A general definition of a fitting mode, plus some mode instances."""
+
+import textwrap
+from enum import Enum
+from typing import Callable, Dict, Union
+
+import numpy as np
+import xarray as xr
+from cmdstanpy import CmdStanMCMC, CmdStanModel
+from pydantic import BaseModel
+from sklearn.model_selection import KFold
+
+
+class IdataTarget(str, Enum):
+    prior = "prior"
+    posterior = "posterior"
+    log_likelihood = "log_likelihood"
+
+
+class FittingMode(BaseModel):
+    """A way of fitting a statistical model.
+
+    Each fitting mode has a name, an 'idata_target', i.e. the group that this
+    mode will create in the output InferenceData, and a function that performs
+    the fitting.
+
+    Inferences can be configured to use any of the fitting modes defined here
+    by including them by name in the top-level list 'modes'. For example:
+
+      ...
+      modes = ['prior', 'posterior', 'kfold']
+      ...
+
+    The idata_target must be one out of 'prior', 'posterior' and
+    'log_likelihood', and specifies which out of these InferenceData groups the
+    output will be written to.
+
+    Each fitting mode's fit function must match the signature specified in the
+    FittingMode class. This can also be changed, but must agree with the
+    'sample' module.
+
+
+    """
+
+    name: str
+    idata_target: IdataTarget
+    fit: Callable[
+        [CmdStanModel, Dict, Dict[str, str]], Union[CmdStanMCMC, xr.Dataset]
+    ]
+
+
+def fit_prior(model: CmdStanModel, input_dict: dict, kwargs) -> CmdStanMCMC:
+    """Create a CmdStanMCMC from a model, data and config, in prior mode.
+
+    :param model: a CmdStanModel. It must have a data variable called
+    'likelihood': setting this variable to 0 should trigger prior mode,
+    typically by causing the code that performs likelihood evaluations not to
+    run.
+
+    :param input_dict: a Stan input dictionary. It doesn't need to have a
+    'likelihood' entry: if it does, it will be overwritten.
+
+    :param kwargs: keyword arguments for CmdStanModel.sample
+
+    """
+    input_dict_final = input_dict | {"likelihood": 0}
+    return model.sample(input_dict_final, **kwargs)
+
+
+def fit_posterior(model: CmdStanModel, input_dict: dict, kwargs) -> CmdStanMCMC:
[…]

### 2023-06-02 bibat/examples/baseball/baseball/sample.py

@@ -1,81 +1,20 @@
-"""Run all the configs in the model_configurations folder."""
+"""Run all the inferences in the inferences folder."""
 
 import os
-from typing import Dict
 
 import arviz as az
 import cmdstanpy
-import numpy as np
-import xarray as xr
-from sklearn.model_selection import KFold
 
 from baseball.data_preparation import load_prepared_data
-from baseball.inference_configuration import (
-    AVAILABLE_MODES,
-    InferenceConfiguration,
-    load_inference_configuration,
-)
-from baseball.util import CoordDict
+from baseball.inference_configuration import load_inference_configuration
 
 HERE = os.path.dirname(__file__)
 RUNS_DIR = os.path.join(HERE, "..", "inferences")
 STAN_DIR = os.path.join(HERE, "stan")
 
 
-def get_sample_kwargs(raw: Dict, mode: str) -> Dict:
-    """Get the sampel kwargs for a mode from raw toml."""
-    out = {k: v for k, v in raw.items() if k not in AVAILABLE_MODES}
-    if mode in raw.keys():
-        out |= raw[mode]
-    return out
-
-
-def get_kfold_llik(
-    ic: InferenceConfiguration,
-    stan_input_base: Dict,
-    model: cmdstanpy.CmdStanModel,
-    coords: CoordDict,
-) -> xr.DataArray:
-    """Run k-fold cross-validation and return out-of-sample log likelihood.
-
-    :param ic: InferenceConfiguration, must have non-null field kfold_folds
-
-    :param stan_input_base: stan input dictionary
-
-    :param model: CmdStanModel
-    """
-    n_folds = ic.kfold_folds
-    sample_kwargs = get_sample_kwargs(ic.sample_kwargs, "kfold")
-    kf = KFold(n_folds, shuffle=True, random_state=1234)
-    lliks_by_fold = []
-    full_ix = np.array(stan_input_base["ix_train"])
-    for fold, (ix_train, ix_test) in enumerate(kf.split(full_ix)):
-        input_dict = stan_input_base | {
-            "likelihood": 1,
-            "N_train": len(ix_train),
-            "N_test": len(ix_test),
-            "ix_train": full_ix[ix_train].tolist(),
-            "ix_test": full_ix[ix_test].tolist(),
-        }
-        mcmc = model.sample(data=input_dict, **sample_kwargs)
-        llik_fold = mcmc.draws_xr(vars=["llik"])
-        # remember the fold
[…]

### 2023-06-02 bibat/examples/baseball/baseball/fitting_mode.py
fix flake8

@@ -12,6 +12,8 @@ from sklearn.model_selection import KFold
 
 
 class IdataTarget(str, Enum):
+    """An enum for choosing the group that a fitting mode writes to."""
+
     prior = "prior"
     posterior = "posterior"
     log_likelihood = "log_likelihood"

### 2023-06-02 bibat/examples/baseball/docs/report.qmd

@@
 The final step was to document my analysis. To do this I edited the file
-`docs/report.qmd`, then ran `quarto render docs/report.qmd`, which produced the
+`docs/report.qmd`, then ran `make docs`, which produced the
 very HTML document that you are probably reading now! You can find the complete

### 2023-07-07 bibat/wizarding.py
update target project for pydantic v2

@@ -3,7 +3,7 @@
 from typing import Callable, Dict, List, Optional, Union
 
 import click
-from pydantic import root_validator
+from pydantic import model_validator
 from pydantic.dataclasses import dataclass
 
 
@@ -23,13 +23,13 @@ class WizardStr:
     default: Optional[str] = None
     default_function: Optional[Callable] = None
 
-    @root_validator
-    def default_or_default_function_exists(cls, values):
+    @model_validator(mode="after")
+    def default_or_default_function_exists(cls, m: "WizardStr"):
         """Either the default or the default function must not be None."""
-        assert (values["default"] is not None) or (
-            values["default_function"] is not None
+        assert (m.default is not None) or (
+            m.default_function is not None
         ), "Either the default or the default function must not be None."
-        return values
+        return m
 
 
 @dataclass
@@ -49,24 +49,21 @@ class WizardChoice:
     default: Optional[str] = None
     default_function: Optional[Callable] = None
 
-    @root_validator
-    def default_is_an_option(cls, values):
+    @model_validator(mode="after")
+    def default_is_an_option(cls, m: "WizardChoice"):
         """Check that the default is one of the options."""
-        if isinstance(values["default"], str):
-            msg = (
-                f"default {values['default']} "
-                f"not in options {values['options']}"
-            )
-            assert values["default"] in values["options"], msg
-            return values
+        if isinstance(m.default, str):
+            msg = f"default {m.default} not in options {m.options}"
+            assert m.default in m.options, msg
+            return m
 
-    @root_validator
-    def default_or_default_function_exists(cls, values):
+    @model_validator(mode="after")
+    def default_or_default_function_exists(cls, m: "WizardChoice"):
         """Either the default or the default function must not be None."""
-        assert (values["default"] is not None) or (
-            values["default_function"] is not None
+        assert (m.default is not None) or (
+            m.default_function is not None
         ), "Either the default or the default function must not be None."
-        return values
+        return m
 
 
 def prompt_user(

### 2023-07-07 bibat/examples/baseball/docs/report.qmd
Update baseball example

@@
 
-```{.python include="../baseball/data_preparation.py" start-line=103}
+```{.python include="../baseball/data_preparation.py" start-line=77 end-line=161}
 ```
 
-I had to update the `prepare_data` function to take into account the
-two raw data sources:
+To take into account the inconsistency between the two raw data sources, I
+first had to change the variable `RAW_DATA_FILES`:
+
+```{.python include="../baseball/data_preparation.py" start-line=26 end-line=33}
+```
+
+Next I changed the `prepare_data` function to handle the two different data
+sources.
 

### 2023-07-07 bibat/examples/baseball/baseball/data_preparation.py

@@ -1,4 +1,4 @@
-"""Provide functions prepare_data_x.
+"""Provides functions prepare_data_x.
 
 These functions should take in a dataframe of measurements and return a
 PreparedData object.
@@ -7,17 +7,19 @@ PreparedData object.
 import json
 import os
 
+import numpy as np
 import pandas as pd
 import pandera as pa
 from pandera.typing import DataFrame, Series
-from pydantic.dataclasses import dataclass
+from pydantic import BaseModel
 
-from baseball.util import CoordDict
+from baseball import util
 
 NAME_FILE = "name.txt"
 COORDS_FILE = "coords.json"
 MEASUREMENTS_FILE = "measurements.csv"
 N_CV_FOLDS = 10
+
 HERE = os.path.dirname(__file__)
 DATA_DIR = os.path.join(HERE, "..", "data")
 RAW_DIR = os.path.join(DATA_DIR, "raw")
@@ -66,38 +68,12 @@ class MeasurementsDF(pa.SchemaModel):
     n_success: Series[int] = pa.Field(ge=0)
 
 
-@dataclass
-class PreparedData:
+class PreparedData(BaseModel, arbitrary_types_allowed=True):
     """What prepared data looks like in this analysis."""
 
     name: str
-    coords: CoordDict
+    coords: util.CoordDict
     measurements: DataFrame[MeasurementsDF]
-
-
-def load_prepared_data(directory: str) -> PreparedData:
-    """Load prepared data from files in directory."""
-    with open(os.path.join(directory, COORDS_FILE), "r") as f:
-        coords = json.load(f)
-    with open(os.path.join(directory, NAME_FILE), "r") as f:
-        name = f.read()
-    measurements = pd.read_csv(os.path.join(directory, MEASUREMENTS_FILE))
-    return PreparedData(
-        name=name,
-        coords=coords,
-        measurements=measurements,
-    )
-
-
-def write_prepared_data(prepped: PreparedData, directory):
-    """Write prepared data files to a directory."""
-    if not os.path.exists(directory):
-        os.mkdir(directory)
-        prepped.measurements.to_csv(os.path.join(directory, MEASUREMENTS_FILE))
-    with open(os.path.join(directory, COORDS_FILE), "w") as f:
-        json.dump(prepped.coords, f)
-    with open(os.path.join(directory, NAME_FILE), "w") as f:
-        f.write(prepped.name)
 
 
 def prepare_data_2006(measurements_raw: pd.DataFrame) -> PreparedData:
[…]

### 2023-07-07 docs/conf.py
bump version

@@ -24,7 +24,7 @@ copyright = "2022, Teddy Groves"
 author = "Teddy Groves"
 
 # The full version, including alpha/beta/rc tags
-release = "0.1.8"
+release = "0.1.9"
 
 root_doc = "index"
 

### 2023-07-07 bibat/examples/baseball/baseball/data_preparation.py
Fix flake8 error

@@ -7,7 +7,6 @@ PreparedData object.
 import json
 import os
 
-import numpy as np
 import pandas as pd
 import pandera as pa
 from pandera.typing import DataFrame, Series

### 2023-07-07 docs/conf.py
bump version

@@ -24,7 +24,7 @@ copyright = "2022, Teddy Groves"
 author = "Teddy Groves"
 
 # The full version, including alpha/beta/rc tags
-release = "0.1.8"
+release = "0.1.10"
 
 root_doc = "index"
 

### 2023-07-07 bibat/examples/baseball/docs/report.qmd

@@
 
-```{.python include="../baseball/data_preparation.py" start-line=77 end-line=161}
+```{.python include="../baseball/data_preparation.py" start-line=78 end-line=162}
 ```
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=35 end-line=54}
+```{.python include="../baseball/data_preparation.py" start-line=36 end-line=56}
 ```

### 2023-12-15 paper/jmlr_cover_letter.qmd
start writing jlmr paper

---
title: Cover letter - Bibat, Batteries-included Bayesian analysis template
---

Dear editors

This paper is intended for the machine learning open source software section
of JLMR.

All the authors are aware of this submission and consent to its review by JMLR.

Please find below all the information listed on the relevant JMLR page here
<https://www.jmlr.org/mloss/mloss-info.html>.

Best wishes,
Teddy Groves

## Open source licence

MIT licence: <https://github.com/teddygroves/bibat/LICENSE>

## Web address of the project

<https://github.com/teddygroves/bibat/>

## Software version to be reviewed

XXX VERSION XXX

## Evidence of active user community

Bibat has XXX GitHub stars and XXX active devlopers. In addition, bibat is a
PyOpenSci affiliated package: the PyOpenSci community provides rigorous peer
review (though not publication), publicity, community engagement and help with
software maintenance.

## Suggested reviewers

The following people have already reviewed bibat for PyOpenSci and therefore
might find it easy to provide another review:

- Meenal Jhajharia <meenal@mjhajharia.com>
- Oriol Abril Plaza <https://github.com/OriolAbril>
- Alexey Izmailov <alizma@brown.edu>

Editor: @mjhajharia
Reviewer 1: @OriolAbril
Reviewer 2: @alizma

### 2023-12-15 paper/main.qmd
start writing jlmr paper

---
title: "Bibat: Batteries-include Bayesian Analysis Template"
keywords: Bayesian inference,
abstract: >
  Despite their suitability for many scientific problems and the existence of
  sound theoretical and computational frameworks, adoption of Bayesian workflow
  in science remains a challenge. One reason for the difficulty is that it is
  difficult to write software that implements non-trivial Bayesian workflows.
  We aimed to address this difficulty by developing Bibat, a Python package
  providing  an interactive template that targets Bayesian statistical analysis
  projects. Bibat trivialises the otherwise painful process of structuring such
  a project and integrating together tools that help with individual parts of a
  Bayesian workflow, without compromising flexibility, scalability,
  interoperability or reproducibility. Bibat is available on the Python
  Package index,  documented at <https://bibat.readthedocs.io/> and
  developed at <https://github.com/teddygroves/bibat/>.
number-sections: true
author:
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations:
      - id: biosustain
        name: DTU
        department: The Novo Nordisk Foundation Center for Biosustainability
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
bibliography: bibliography.bib
format:
  html:
    embed-resources: true
  pdf:
    keep-tex: true
---

Many areas of science would benefit from replacing traditional statistical
methodologies with Bayesian workflow, as described in, for example
[@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
@gabryVisualizationBayesianWorkflow2019].

There are now software tools that address most individual aspects of a Bayesian
workflow, from fetching and manipulating data to specifying, computing, storing,
inspecting and documenting statistical inferences. However, writing a software
project implementing a Bayesian workflow remains challenging due partly to the
difficulty of orchestrating and configuring these elements.

Bibat is a Python package that addresses this difficulty by providing an
interactive template for Bayesian workflow projects.

# Installation and usage

Bibat is installed by running the command `pip install bibat` and then used by
running the command `bibat`.

This command triggers an interactive form which prompts the user for
configuration information including project and repository name, author name, a
short description and choices of open source licence options, documentation
formats and whether to include tests and continuous integration.

Bibat then creates a folder with the chosen repository name in the current
working directory, containing code that implements bibat's example analysis.

# Documentation

Bibat is documented at <https://github.com/teddygroves/bibat/>. The
documentation website includes instructions for getting started, a detailed
explanation of bibat's concepts, an extended vignette illustrating intended
usage, full description of the python API and command line interface,
instructions for contributing and a section discussing accessibility
considerations.

Bibat's documentation website is continuously checked to ensure that it
implements accessibility best practices.

# Software

Bibat is implemented using Python 3. Its main dependencies are cookiecutter
[@greenfeldCookiecutter2021], pydantic [@pydanticdevelopersPydantic2022] and
click [@clickdevelopersClickPythonComposable2022].

Bibat projects have the following dependencies:

* Python 3 [@vanrossumPythonReferenceManual2009] and standard scientific Python
packages for data manipulation
* Stan [@carpenterStanProbabilisticProgramming2017] for specifying statistical
models and performing inferences
* cmdstanpy [@standevelopmentteamCmdStanPy2022] for interfacing between Python
and Stan
* arviz [@kumarArviZUnifiedLibrary2019] for storing and analysing completed
inferences
* pydantic and pandera [@niels_bantilan-proc-scipy-2020] for validation
* make [@stallman1991gnu] for automation.
 * Sphinx [@georgbrandlandthesphinxteamSphinx2022] and Quarto
[@Allaire_Quarto_2022] for documentation.

Bibat is continuously tested to ensure that it works on the operating systems
Linux, macOS and Windows.

Bibat's continuous integration runs on windows and ubuntu and runs bibat's test
suite as well as an end-to-end functional test on all supported Python versions.

# How bibat addresses specific Bayesian workflow difficulties

This section describes some specific problems that often affect Bayesian
workflow projects and which bibat helps to ameliorate.

## Data modelling

Any statistical analysis requires a definition of a prepared dataset, whether
implicitly or explicitly. Ideally the definition is explicit and allows for
arbitrarily many prepared datasets, accommodates both tabular and non-tabular
formats and provides functionality like validation and serialisation.

Bibat achieves this goal by providing abstract models for prepared data,
preconfigured to work well together, integrated into a data preparation pipeline
and able to be easily customised to suit any analysis.

Without a template such as bibat, users must either devise a new data modelling
framework from scratch as part of an analysis, which is time-consuming, or
else to miss out on either the flexibility or features that bibat's approach
provides.

## Reproducibility

Bibat ensures reproducibility by providing a preconfigured makefile with a
target `analysis` triggering creation of an isolated environment, installation
of dependencies, data preparation, statistical computation and analysis of
results. In this way a bibat analysis can be reproduced on most platforms
using a single command.

Bibat also provides its Python code in the form of a package configured using
modern conventions for specifying dependencies and configuring tooling, so that
it is easy to maintain reproducibility as the analysis develops.

## Collaboration

Bibat provides a preconfigured test environment, continuous integration, linting
and pre-commit hooks, making it suitable for collaborative software development.
In addition, including documentation as a first class component of the analysis
addresses a common problem in academic statistics projects where the paper gets
out of sync with the code.

## Installing dependencies

Bibat's makefile detects the current operating system and attempts to install
cmdstan appropriately if necessary. This functionality addresses a common
issue where researchers find it difficult to install Stan, especially on Windows.

## Fitting modes

As part of a Bayesian workflow it is often necessary to fit a model and dataset
in different ways. For example, one might perform MCMC sampling of both the
prior and posterior distributions, perform multiple leave-out-one-fold fits for
cross-validation or need to compare MCMC sampling with an optimisation-based
alternative.

Bibat accommodates this ubiquitous scenario by introducing an abstraction called
"fitting mode" and a corresponding Pydantic base class `FittingMode`, along with
and several subclasses for commonly used cases.

This abstraction allows bibat projects to handle fitting a model and dataset in
different ways appropriately and flexibly. For example, the provided prior
sampling fitting mode creates a Stan input dictionary with the `likelihood`
data variable set to `0`, performs MCMC sampling and writes data to the
`InferenceData` group `prior`.

## Configure fitting concisely.

Bibat provides a Pydantic class `InferenceConfiguration`, parsed from toml files
in the `inferences` directory. By writing a toml file per inference the user can
configure statistical model, dataset, fitting modes and computational settings
(globally or per fitting mode).

This approach appropriately separates configuration from logic and avoids
unnecessary duplication of configuration for runs with the same model and
dataset.

## Storing MCMC samples

Bibat provides code that automatically saves its output `InferenceData` objects
in zarr format. This format distributes information from the same set of samples
into multiple smaller files, thereby helping to avoid file size limits of online
repository hosting services.

# Comparison with alternative software

Other than bibat, there is currently no interactive template that specifically
targets Bayesian workflow projects. There are some templates that arguably
encompass Bayesian workflow as a special case of data analysis project, such
as cookiecutter-data-science [@drivendataCookiecutterdatascience2022], but
these are of limited use compared with a specialised template due to the many
specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating Bayesian
workflow, but differently from bibat. For example, bambi [@capretto2020]
and brms [@burknerBrmsPackageBayesian2017] aim to make implementing Bayesian
workflows easier by trivialising the task of specifying a Bayesian generalised
linear model and fitting it to a single tabular dataset. This approach does not
address several difficulties that bibat does address,

# References

### 2023-12-15 paper/main.tex
start writing jlmr paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bibat: Batteries-include Bayesian Analysis Template},
  pdfauthor={Teddy Groves},
  pdfkeywords={Bayesian inference,},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Bibat: Batteries-include Bayesian Analysis Template}
\author{Teddy Groves}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Despite their suitability for many scientific problems and the existence
of sound theoretical and computational frameworks, adoption of Bayesian
workflow in science remains a challenge. One reason for the difficulty
is that it is difficult to write software that implements non-trivial
Bayesian workflows. We aimed to address this difficulty by developing
Bibat, a Python package providing an interactive template that targets
Bayesian statistical analysis projects. Bibat trivialises the otherwise
painful process of structuring such a project and integrating together
tools that help with individual parts of a Bayesian workflow, without
compromising flexibility, scalability, interoperability or
reproducibility. Bibat is available on the Python Package index,
documented at \url{https://bibat.readthedocs.io/} and developed at
\url{https://github.com/teddygroves/bibat/}.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

Many areas of science would benefit from replacing traditional
statistical methodologies with Bayesian workflow, as described in, for
example (Gelman et al. 2020; Grinsztajn et al. 2021; Gabry et al. 2019).

There are now software tools that address most individual aspects of a
Bayesian workflow, from fetching and manipulating data to specifying,
computing, storing, inspecting and documenting statistical inferences.
However, writing a software project implementing a Bayesian workflow
remains challenging due partly to the difficulty of orchestrating and
configuring these elements.

Bibat is a Python package that addresses this difficulty by providing an
interactive template for Bayesian workflow projects.

\hypertarget{installation-and-usage}{%
\section{Installation and usage}\label{installation-and-usage}}

Bibat is installed by running the command \texttt{pip\ install\ bibat}
and then used by running the command \texttt{bibat}.

This command triggers an interactive form which prompts the user for
configuration information including project and repository name, author
name, a short description and choices of open source licence options,
documentation formats and whether to include tests and continuous
integration.

Bibat then creates a folder with the chosen repository name in the
current working directory, containing code that implements bibat's
example analysis.

\hypertarget{documentation}{%
\section{Documentation}\label{documentation}}

Bibat is documented at \url{https://github.com/teddygroves/bibat/}. The
documentation website includes instructions for getting started, a
detailed explanation of bibat's concepts, an extended vignette
illustrating intended usage, full description of the python API and
command line interface, instructions for contributing and a section
discussing accessibility considerations.

\hypertarget{software}{%
\section{Software}\label{software}}

Bibat is implemented using Python 3. Its main dependencies are
cookiecutter (Greenfeld et al. 2021), pydantic (Pydantic developers
2022) and click (Click Developers 2022). Bibat is continuously tested to
ensure that it works on the operating systems Linux, macOS and Windows.

Bibat projects use the following tools:

\begin{itemize}
\tightlist
\item
  Python 3 (Van Rossum and Drake 2009) and standard scientific Python
  packages for data manipulation
\item
  Stan (Carpenter et al. 2017) for specifying statistical models and
  performing inferences
\item
  cmdstanpy (Stan Development Team 2022) for interfacing between Python
  and Stan
\item
  arviz (Kumar et al. 2019) for storing and analysing completed
  inferences
\item
  pydantic and pandera (Niels Bantilan 2020) for validation
\item
  make (Stallman and McGrath 1991) for automation.
\item
  Sphinx (Georg Brandl and the Sphinx team 2022) and Quarto (Allaire et
  al. 2022) for documentation.
\end{itemize}

\hypertarget{how-bibat-addresses-specific-bayesian-workflow-difficulties}{%
\section{How bibat addresses specific Bayesian workflow
difficulties}\label{how-bibat-addresses-specific-bayesian-workflow-difficulties}}

This section describes some specific problems that often affect Bayesian
workflow projects and which bibat helps to ameliorate.

\hypertarget{data-modelling}{%
\subsection{Data modelling}\label{data-modelling}}

Any statistical analysis requires a definition of a prepared dataset,
whether implicitly or explicitly. Ideally the definition is explicit and
allows for arbitrarily many prepared datasets, accommodates both tabular
and non-tabular formats and provides functionality like validation and
serialisation.

Bibat achieves this goal by providing abstract models for prepared data,
preconfigured to work well together, integrated into a data preparation
pipeline and able to be easily customised to suit any analysis.

Without a template such as bibat, users must either devise a new data
modelling framework from scratch as part of an analysis, which is
time-consuming, or else to miss out on either the flexibility or
features that bibat's approach provides.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

Bibat ensures reproducibility by providing a preconfigured makefile with
a target \texttt{analysis} triggering creation of an isolated
environment, installation of dependencies, data preparation, statistical
computation and analysis of results. In this way a bibat analysis can be
reproduced on most platforms using a single command.

Bibat also provides its Python code in the form of a package configured
using modern conventions for specifying dependencies and configuring
tooling, so that it is easy to maintain reproducibility as the analysis
develops.

\hypertarget{collaboration}{%
\subsection{Collaboration}\label{collaboration}}

Bibat provides a preconfigured test environment, continuous integration,
linting and pre-commit hooks, making it suitable for collaborative
software development. In addition, including documentation as a first
class component of the analysis addresses a common problem in academic
statistics projects where the paper gets out of sync with the code.

\hypertarget{installing-dependencies}{%
\subsection{Installing dependencies}\label{installing-dependencies}}

Bibat's makefile detects the current operating system and attempts to
install cmdstan appropriately if necessary. This functionality addresses
a common issue where researchers find it difficult to install Stan,
especially on Windows.

\hypertarget{fitting-modes}{%
\subsection{Fitting modes}\label{fitting-modes}}

As part of a Bayesian workflow it is often necessary to fit a model and
dataset in different ways. For example, one might perform MCMC sampling
of both the prior and posterior distributions, perform multiple
leave-out-one-fold fits for cross-validation or need to compare MCMC
sampling with an optimisation-based alternative.

Bibat accommodates this ubiquitous scenario by introducing an
abstraction called ``fitting mode'' and a corresponding Pydantic base
class \texttt{FittingMode}, along with and several subclasses for
commonly used cases.

This abstraction allows bibat projects to handle fitting a model and
dataset in different ways appropriately and flexibly. For example, the
provided prior sampling fitting mode creates a Stan input dictionary
with the \texttt{likelihood} data variable set to \texttt{0}, performs
MCMC sampling and writes data to the \texttt{InferenceData} group
\texttt{prior}.

\hypertarget{configure-fitting-concisely.}{%
\subsection{Configure fitting
concisely.}\label{configure-fitting-concisely.}}

Bibat provides a Pydantic class \texttt{InferenceConfiguration}, parsed
from toml files in the \texttt{inferences} directory. By writing a toml
file per inference the user can configure statistical model, dataset,
fitting modes and computational settings (globally or per fitting mode).

This approach appropriately separates configuration from logic and
avoids unnecessary duplication of configuration for runs with the same
model and dataset.

\hypertarget{storing-mcmc-samples}{%
\subsection{Storing MCMC samples}\label{storing-mcmc-samples}}

Bibat provides code that automatically saves its output
\texttt{InferenceData} objects in zarr format. This format distributes
information from the same set of samples into multiple smaller files,
thereby helping to avoid file size limits of online repository hosting
services.

\hypertarget{comparison-with-alternative-software}{%
\section{Comparison with alternative
software}\label{comparison-with-alternative-software}}

Other than bibat, there is currently no interactive template that
specifically targets Bayesian workflow projects. There are some
templates that arguably encompass Bayesian workflow as a special case of
data analysis project, such as cookiecutter-data-science (Driven Data
2022), but these are of limited use compared with a specialised template
due to the many specificities of Bayesian workflow.

There is some software that addresses the general task of facilitating
Bayesian workflow, but differently from bibat. For example, bambi
(Capretto et al. 2020) and brms (Bürkner 2017) aim to make implementing
Bayesian workflows easier by trivialising the task of specifying a
Bayesian generalised linear model and fitting it to a single tabular
dataset. This approach does not address several difficulties that bibat
does address,

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Allaire_Quarto_2022}{}}%
Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
Christophe Dervieux. 2022. {``Quarto.''}
\url{https://doi.org/10.5281/zenodo.5960048}.

\leavevmode\vadjust pre{\hypertarget{ref-burknerBrmsPackageBayesian2017}{}}%
Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-capretto2020}{}}%
Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
{Bayesian} Linear Models in {Python}.''}
\url{https://arxiv.org/abs/2012.10754}.

\leavevmode\vadjust pre{\hypertarget{ref-carpenterStanProbabilisticProgramming2017}{}}%
Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
\url{https://doi.org/10.18637/jss.v076.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-clickdevelopersClickPythonComposable2022}{}}%
Click Developers. 2022. {``Click: {Python} Composable Command Line
Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.

\leavevmode\vadjust pre{\hypertarget{ref-drivendataCookiecutterdatascience2022}{}}%
Driven Data. 2022. {``Cookiecutter-Data-Science.''}
\url{https://github.com/drivendata/cookiecutter-data-science/}.

\leavevmode\vadjust pre{\hypertarget{ref-gabryVisualizationBayesianWorkflow2019}{}}%
Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
\emph{Journal of the Royal Statistical Society Series A: Statistics in
Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.

\leavevmode\vadjust pre{\hypertarget{ref-gelmanBayesianWorkflow2020}{}}%
Gelman, Andrew, Aki Vehtari, Daniel Simpson, Charles C. Margossian, Bob
Carpenter, Yuling Yao, Lauren Kennedy, Jonah Gabry, Paul-Christian
Bürkner, and Martin Modrák. 2020. {``Bayesian {Workflow}.''}
\emph{arXiv:2011.01808 {[}Stat{]}}, November.
\url{http://arxiv.org/abs/2011.01808}.

\leavevmode\vadjust pre{\hypertarget{ref-georgbrandlandthesphinxteamSphinx2022}{}}%
Georg Brandl and the Sphinx team. 2022. {``Sphinx.''}
\url{https://www.sphinx-doc.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-greenfeldCookiecutter2021}{}}%
Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.

\leavevmode\vadjust pre{\hypertarget{ref-grinsztajnBayesianWorkflowDisease2021}{}}%
Grinsztajn, Léo, Elizaveta Semenova, Charles C. Margossian, and Julien
Riou. 2021. {``Bayesian Workflow for Disease Transmission Modeling in
{Stan}.''} \emph{Statistics in Medicine} 40 (27): 6209--34.
\url{https://doi.org/10.1002/sim.9164}.

\leavevmode\vadjust pre{\hypertarget{ref-kumarArviZUnifiedLibrary2019}{}}%
Kumar, Ravin, Colin Carroll, Ari Hartikainen, and Osvaldo Martin. 2019.
{``{ArviZ} a Unified Library for Exploratory Analysis of {Bayesian}
Models in {Python}.''} \emph{Journal of Open Source Software} 4 (33):
1143. \url{https://doi.org/10.21105/joss.01143}.

\leavevmode\vadjust pre{\hypertarget{ref-niels_bantilan-proc-scipy-2020}{}}%
Niels Bantilan. 2020. {``Pandera: {Statistical Data Validation} of
{Pandas Dataframes}.''} In \emph{Proceedings of the 19th {Python} in
{Science Conference}}, edited by Meghann Agarwal, Chris Calloway, Dillon
Niederhut, and David Shupe, 116--24.
\url{https://doi.org/10.25080/Majora-342d178e-010}.

\leavevmode\vadjust pre{\hypertarget{ref-pydanticdevelopersPydantic2022}{}}%
Pydantic developers. 2022. {``Pydantic.''}
\url{https://pypi.org/project/pydantic/}.

\leavevmode\vadjust pre{\hypertarget{ref-stallman1991gnu}{}}%
Stallman, Richard M, and Roland McGrath. 1991. {``{GNU} Make-{A} Program
for Directing Recompilation.''}

\leavevmode\vadjust pre{\hypertarget{ref-standevelopmentteamCmdStanPy2022}{}}%
Stan Development Team. 2022. {``{CmdStanPy}.''}
\url{https://github.com/stan-dev/cmdstanpy}.

\leavevmode\vadjust pre{\hypertarget{ref-vanrossumPythonReferenceManual2009}{}}%
Van Rossum, Guido, and Fred L. Drake. 2009. \emph{Python 3 {Reference
Manual}}. {Scotts Valley, CA}: {CreateSpace}.

\end{CSLReferences}

\end{document}

### 2023-12-15 paper/softwarex_template.tex
start writing jlmr paper

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt, a4paper]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{hyperref}
\setlength{\parindent}{0pt}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%\usepackage{lineno}

\journal{SoftwareX}

\begin{document}
\renewcommand{\labelenumii}{\arabic{enumi}.\arabic{enumii}}

\begin{frontmatter}
%--- INSTRUCTIONS TO BE DELETED OR COMMENTED BEFORE SUBMISSION

  {\large\textbf{SoftwareX article template Version 4 (November 2023)}}
  Before you complete this template, a few important points to note:
  \begin{itemize}
\item	This template is for an original SoftwareX article. If you are submitting an update to a software article that has already been published, please use the Software Update Template found on the Guide for Authors page.
\item	The format of a software article is very different to a traditional research article. To help you write yours, we have created this template. We will only consider software articles submitted using this template.
\item	It is mandatory to publicly share the code/software referred to in your software article. You’ll find information on our software sharing criteria in the SoftwareX \href{https://www.elsevier.com/journals/softwarex/2352-7110/guide-for-authors}{Guide for Authors}.
\item	It’s important to consult the \href{https://www.elsevier.com/journals/softwarex/2352-7110/guide-for-authors}{Guide for Authors} when preparing your manuscript; it highlights mandatory requirements and is packed with useful advice.
\end{itemize}
%
Still got questions?
Email our editorial team at softwarex@elsevier.com.

\textbf{Now you are ready to fill in the template below. As you complete each section, please carefully read the associated instructions. All sections are mandatory, unless marked optional.
Once you have completed the template, delete these instructions. In addition, please delete the instructions in the template (the text written in italics).}
%--- END OF INSTRUCTIONS TO BE DELETED OR COMMENTED BEFORE SUBMISSION

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Title (Name of your software: Then a short title)}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[label1]{A. Author1}
\author[label2]{A. Author2}
\address[label1]{Author1's affiliation, address, email}
\address[label2]{Author2's affiliation, address, email}

\begin{abstract}
%% Text of abstract
\textit{Ca. 100 words. The abstract is followed by a maximum of six keywords
and some mandatory and optional metadata.}

\textit{Your main body of text (sections 1-5 below) should be a maximum 6 pages in total (excluding metadata, tables, figures, references) with a 3000-word limit (we ask that more priority is placed on the word limit versus the page count). Though we strictly insist on the author following the template, in exceptional circumstances, we can be flexible with the page numbers and word limit. In such cases, it should be discussed with the managing editor or publisher prior to submission. All queries regarding the same can be reached at softwarex@elsevier.com.}
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
keyword 1 \sep keyword 2 \sep keyword 3

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%\linenumbers

\section*{Metadata}
\label{}
\textit{The ancillary data table~\ref{codeMetadata} is required for the sub-version of the codebase. Please replace the italicized text in the right column with the correct information about your current code and leave the left column untouched.}

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this column} \\
\hline
C1 & Current code version & For example v42 \\
\hline
C2 & Permanent link to code/repository used for this code version & For example: \url{https://github.com/mozart/mozart2} \\
\hline
C3  & Permanent link to Reproducible Capsule & For example: \url{https://codeocean.com/capsule/0270963/tree/v1}\\
\hline
C4 & Legal Code License   & All software and code must be released under one of the pre-approved licenses listed in the \href{https://www.elsevier.com/journals/softwarex/2352-7110/guide-for-authors}{Guide for Authors}, such as Apache License, GNU General Public License (GPL) or MIT License. Write the name of the license you’ve chosen here. \\
\hline
C5 & Code versioning system used & For example: svn, git, mercurial,
                                   etc. (put none if none is used) \\
\hline
C6 & Software code languages, tools, and services used & For example: C++, python, r, MPI, OpenCL, etc. \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & \\
\hline
C8 & If available Link to developer documentation/manual & For example: \url{http://mozart.github.io/documentation/} \\
\hline
C9 & Support email for questions & \\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{codeMetadata}
\end{table}

\textit{Optionally, you can provide information about the current executable
software version filling in the left column of
Table~\ref{executabelMetadata}. Please leave the first column as it is.}

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{(Executable) software metadata description} & \textbf{Please fill in this column} \\
\hline
S1 & Current software version & For example 1.1, 2.4 etc. \\
\hline
S2 & Permanent link to executables of this version  & For example: \url{https://github.com/combogenomics/DuctApe/releases/tag/DuctApe-0.16.4} \\
\hline
S3  & Permanent link to Reproducible Capsule & \\
\hline
S4 & Legal Software License & List one of the approved licenses \\
\hline
S5 & Computing platforms/Operating Systems & For example Android, BSD, iOS, Linux, OS X, Microsoft Windows, Unix-like , IBM z/OS, distributed/web based etc. \\
\hline
S6 & Installation requirements \& dependencies & \\
\hline
S7 & If available, link to user manual - if formally published include a reference to the publication in the reference list & For example: \url{http://mozart.github.io/documentation/} \\
\hline
S8 & Support email for questions & \\
\hline
\end{tabular}
\caption{Software metadata (optional)}
\label{executabelMetadata}
\end{table}

\section{Motivation and significance}
\textit{In this section, we want you to introduce the scientific background and the motivation for developing the software.}

\begin{itemize}
    \item \textit{Explain why the software is important and describe the exact (scientific) problem(s) it solves.}
    \item \textit{Indicate in what way the software has contributed (or will contribute in the future) to the process of scientific discovery; if available, please cite a research paper using the software.}
    \item \textit{Provide a description of the experimental setting. (How does the user use the software?)}
    \item \textit{Introduce related work in literature (cite or list algorithms used, other software etc.).}
\end{itemize}

\section{Software description}

\textit{Describe the software. Provide enough detail to help the reader understand its impact. }

\subsection{Software architecture}
\textit{  Give a short overview of the overall software architecture; provide a pictorial overview where possible; for example, an image showing the components. If necessary, provide implementation details.}

 \subsection{Software functionalities}
\textit{  Present the major functionalities of the software.}

 \subsection{Sample code snippets analysis (optional)}

\section{Illustrative examples}

\textit{Provide at least one illustrative example to demonstrate the major
functions of your software/code.}

\textit{\textbf{Optional}: you may include one explanatory  video or screencast that will appear next to your article, in the right hand side panel. Please upload any video as a single supplementary file with your article. Only one MP4 formatted, with 150MB maximum size, video is possible per article. Recommended video dimensions are 640 x 480 at a maximum of 30 frames / second. Prior to submission please test and validate your .mp4 file at  \url{http://elsevier-apps.sciverse.com/GadgetVideoPodcastPlayerWeb/verification} . This tool will display your video exactly in the same way as it will appear on ScienceDirect. }

\section{Impact}
\textit{This is the main section of the article and reviewers will weight it appropriately.
Please indicate:}
\begin{itemize}
    \item \textit{Any new research questions that can be pursued as a result of your software.}
    \item \textit{In what way, and to what extent, your software improves the pursuit of existing research questions.}
    \item \textit{Any ways in which your software has changed the daily practice of its users.}
    \item \textit{How widespread the use of the software is within and outside the intended user group (downloads, number of users if your software is a service, citable publications, etc.).}
    \item \textit{How the software is being used in commercial settings and/or how it has led to the creation of spin-off companies.}
    \end{itemize}
\textit{Please note that points 1 and 2 are best demonstrated by
  references to citable publications.}

\section{Conclusions}

\section*{Acknowledgements}
\label{}
\textit{Optional. You can use this section to acknowledge colleagues who don’t qualify as a co-author but helped you in some way. }

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References:
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num}
%%  \bibliography{<your bibdatabase>}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

\bibitem{} Use this style of ordering. References in-text should also use a similar style.

\end{thebibliography}

\textit{If the software repository you used supplied a DOI or another
Persistent IDentifier (PID), please add a reference for your software
here. For more guidance on software citation, please see our guide for
authors or \href{https://f1000research.com/articles/9-1257/v2}{this
  article on the essentials of software citation by FORCE 11}, of
which Elsevier is a member.}

\large{\textbf{Reminder: Before you submit, please delete all
the instructions in this document,
including this paragraph.
Thank you!}}

\end{document}
\endinput
%%
%% End of file `SoftwareX_article_template.tex'.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2023-12-15 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/data_preparation.py
Save prepared data in a single file
This change saves each `PreparedData` object as a single json file,
avoiding the need for complicated saving and loading operations.

@@ -1,61 +1,28 @@
-"""Provides functions prepare_data_x.
+"""Provides function `prepare_data` and runs it.
 
-These functions should take in a dataframe of measurements and return a
-PreparedData object.
+This function should run some other functions with names `prepare_data_x`, which
+each take in a dataframe of measurements and return a PreparedData object.
 
 """
 import json
 import os
+from io import StringIO
+from pathlib import Path
+from typing import Any
 
 import numpy as np
 import pandas as pd
 import pandera as pa
+from pandas._libs.lib import item_from_zerodim
 from pandera.typing import DataFrame, Series
-from pydantic import BaseModel
+from pandera.typing.common import DataFrameBase
+from pydantic import BaseModel, field_serializer, field_validator
 from {{cookiecutter.repo_name}} import util
 
-NAME_FILE = "name.txt"
-COORDS_FILE = "coords.json"
-MEASUREMENTS_FILE = "measurements.csv"
-NEW_COLNAMES = {"yButIThoughtIdAddSomeLetters": "y"}
-DROPNA_COLS = ["y"]
-N_CV_FOLDS = 10
-DIMS = {
-    "b": ["covariate"],
-    "y": ["observation"],
-    "yrep": ["observation"],
-    "llik": ["observation"],
-}
-
-HERE = os.path.dirname(__file__)
-DATA_DIR = os.path.join(HERE, "..", "data")
-RAW_DIR = os.path.join(DATA_DIR, "raw")
-PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
-RAW_DATA_FILES = {
-    "raw_measurements": os.path.join(RAW_DIR, "raw_measurements.csv"),
-}
-
-
-def prepare_data():
-    """Run main function."""
-    print("Reading raw data...")
-    raw_data = {
-        k: pd.read_csv(v, index_col=None) for k, v in RAW_DATA_FILES.items()
-    }
-    data_preparation_functions_to_run = [
-        prepare_data_interaction,
-        prepare_data_no_interaction,
-        prepare_data_fake_interaction
-    ]
-    print("Preparing data...")
-    for dpf in data_preparation_functions_to_run:
-        print(f"Running data preparation function {dpf.__name__}...")
-        prepared_data = dpf(raw_data["raw_measurements"])
-        output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
-        print(f"\twriting files to {output_dir}")
-        if not os.path.exists(PREPARED_DIR):
[…]

### 2023-12-15 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/prepare_data.py

@@ -1,6 +1,0 @@
-"""Run the prepare_data function from the data_preparaion module."""
-
-from {{cookiecutter.repo_name}}.data_preparation import prepare_data
-
-if __name__ == "__main__":
-    prepare_data()

### 2024-01-25 paper/main.qmd
Small paper update

@@
 Bibat accommodates this ubiquitous scenario by introducing an abstraction called
-"fitting mode" and a corresponding Pydantic base class `FittingMode`, along with
-and several subclasses for commonly used cases.
-
-This abstraction allows bibat projects to handle fitting a model and dataset in
-different ways appropriately and flexibly. For example, the provided prior
-sampling fitting mode creates a Stan input dictionary with the `likelihood`
-data variable set to `0`, performs MCMC sampling and writes data to the
-`InferenceData` group `prior`.
+"fitting mode". This abstraction allows bibat projects to handle fitting a
+model and dataset in different ways appropriately and flexibly. For example, the
+provided prior sampling fitting mode creates a Stan input dictionary with the
+`likelihood` data variable set to `0`, performs MCMC sampling and writes data
+to the `InferenceData` group `prior`. Bibat provides fitting modes corresponding
+for prior sampling, posterior sampling and kfold posterior sampling. Users can
+easily add additional fitting modes by copying these examples.
 
@@
 and brms [@burknerBrmsPackageBayesian2017] aim to make implementing Bayesian
-workflows easier by trivialising the task of specifying a Bayesian generalised
-linear model and fitting it to a single tabular dataset. This approach does not
-address several difficulties that bibat does address,
+workflows easier by providing ergonomic ways to specify and fit Bayesian
+generalised models to tabular datasets. Bibat is complementary with these
+packages, as it targets use cases that they do not support, such as analyses
+where complex datasets or models might be required.
+
+#
 

### 2024-01-25 bibat/examples/baseball/docs/report.qmd
Update baseball case study

@@
 How would you like to document your project? (Quarto, Sphinx, No docs) [Quarto]:
-Would you like to create a tests directory? [y]: n
 Would you like to create a .github directory? [y]: n
@@
 │   ├── notebooks
-│   ├── prepare_data.py
 │   ├── readme.md
@@
 │   └── posterior_predictive_comparison.png
-└── pyproject.toml
+├── pyproject.toml
+└── tests
+    ├── test_integration
+    │   └── test_data_preparation.py
+    └── test_unit
+        ├── test_inference_configuration.py
+        └── test_util.py
 ```
@@
 Writing 2006 data to data/raw/2006.csv
-Fetching bdb-main data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Batting.csv
+Fetching bdb-main data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Batting.csv
 Writing bdb-main data to data/raw/bdb-main.csv
-Fetching bdb-post data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/BattingPost.csv
+Fetching bdb-post data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/BattingPost.csv
 Writing bdb-post data to data/raw/bdb-post.csv
-Fetching bdb-apps data from https://raw.githubusercontent.com/chadwickbureau/baseballdatabank/master/core/Appearances.csv
+Fetching bdb-apps data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Appearances.csv
 Writing bdb-apps data to data/raw/bdb-apps.csv
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=78 end-line=162}
+```{.python include="../baseball/data_preparation.py" start-line=74 end-line=159}
 ```
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=36 end-line=56}
+```{.python include="../baseball/data_preparation.py" start-line=25 end-line=32}
 ```
@@
 > source .venv/bin/activate
-(baseball) > python baseball/prepare_data.py
+(baseball) > python baseball/data_preparation.py
 ```
@@
 are some functions that do this for both of our models:
-
+^[These functions require some more imports: `numpy` and `scipy.special.expit`]
 ```{.python include="../baseball/stan_input_functions.py" start-line=36}

### 2024-01-25 bibat/examples/baseball/baseball/data_preparation.py

@@ -1,28 +1,28 @@
-"""Provides functions prepare_data_x.
+"""Provides function `prepare_data` and runs it.
 
-These functions should take in a dataframe of measurements and return a
-PreparedData object.
+This function should run some other functions with names `prepare_data_x`, which
+each take in a dataframe of measurements and return a PreparedData object.
 
 """
 import json
 import os
+from io import StringIO
+from pathlib import Path
+from typing import Any
 
+import numpy as np
 import pandas as pd
 import pandera as pa
 from pandera.typing import DataFrame, Series
-from pydantic import BaseModel
+from pandera.typing.common import DataFrameBase
+from pydantic import BaseModel, field_serializer, field_validator
+from pydantic_core.core_schema import field_after_validator_function
 
 from baseball import util
 
-NAME_FILE = "name.txt"
-COORDS_FILE = "coords.json"
-MEASUREMENTS_FILE = "measurements.csv"
-N_CV_FOLDS = 10
-
-HERE = os.path.dirname(__file__)
-DATA_DIR = os.path.join(HERE, "..", "data")
-RAW_DIR = os.path.join(DATA_DIR, "raw")
-PREPARED_DIR = os.path.join(DATA_DIR, "prepared")
+HERE = Path(__file__).parent
+RAW_DIR = HERE / ".." / "data" / "raw"
+PREPARED_DIR = HERE / ".." / "data" / "prepared"
 RAW_DATA_FILES = {
     "2006": [os.path.join(RAW_DIR, "2006.csv")],
     "bdb": [
@@ -31,28 +31,6 @@ RAW_DATA_FILES = {
         os.path.join(RAW_DIR, "bdb-apps.csv"),
     ],
 }
-
-
-def prepare_data():
-    """Run main function."""
-    print("Reading raw data...")
-    raw_data = {
-        k: [pd.read_csv(file, index_col=None) for file in v]
-        for k, v in RAW_DATA_FILES.items()
-    }
-    data_preparation_functions_to_run = {
-        "2006": prepare_data_2006,
-        "bdb": prepare_data_bdb,
-    }
-    print("Preparing data...")
-    for name, dpf in data_preparation_functions_to_run.items():
-        print(f"Running data preparation function {dpf.__name__}...")
-        prepared_data = dpf(*raw_data[name])
-        output_dir = os.path.join(PREPARED_DIR, prepared_data.name)
-        print(f"\twriting files to {output_dir}")
-        if not os.path.exists(PREPARED_DIR):
[…]

### 2024-01-25 docs/conf.py
small changes

@@ -24,7 +24,7 @@ copyright = "2022, Teddy Groves"
 author = "Teddy Groves"
 
 # The full version, including alpha/beta/rc tags
-release = "0.1.12"
+release = "0.2.0"
 
 root_doc = "index"
 

### 2024-01-25 bibat/examples/baseball/baseball/data_preparation.py
Fix flake8 errors

@@ -10,13 +10,11 @@ from io import StringIO
 from pathlib import Path
 from typing import Any
 
-import numpy as np
 import pandas as pd
 import pandera as pa
 from pandera.typing import DataFrame, Series
 from pandera.typing.common import DataFrameBase
 from pydantic import BaseModel, field_serializer, field_validator
-from pydantic_core.core_schema import field_after_validator_function
 
 from baseball import util
 
@@ -54,6 +52,7 @@ class PreparedData(BaseModel, arbitrary_types_allowed=True):
 
     @field_validator("measurements")
     def validate_measurements(cls, v: Any) -> DataFrameBase[MeasurementsDF]:
+        """Validate the measurements field."""
         if isinstance(v, str):
             v = pd.read_json(StringIO(v))
         return MeasurementsDF.validate(v)
@@ -62,6 +61,7 @@ class PreparedData(BaseModel, arbitrary_types_allowed=True):
     def serialize_measurements(
         self, measurements: DataFrame[MeasurementsDF], _info
     ):
+        """Serialise the measurements field."""
         return measurements.to_json()
 
 

### 2024-01-25 paper/jmlr_cover_letter.qmd
small cover letter changes

@@
 
-XXX VERSION XXX
+0.2.0
 
@@
 
-Bibat has XXX GitHub stars and XXX active devlopers. In addition, bibat is a
-PyOpenSci affiliated package: the PyOpenSci community provides rigorous peer
-review (though not publication), publicity, community engagement and help with
-software maintenance.
+Bibat has 16 GitHub stars and has been actively developed over about three
+years. In addition, bibat is a PyOpenSci affiliated package: the PyOpenSci
+community provides rigorous peer review (though not publication), publicity,
+community engagement and help with software maintenance.
 
@@
 
-The following people have already reviewed bibat for PyOpenSci and therefore
-might find it easy to provide another review:
+These people previously reviewed bibat for PyOpenSci and therefore might find it
+easy to provide another review:
 
@@
 - Alexey Izmailov <alizma@brown.edu>
-
-Editor: @mjhajharia
-Reviewer 1: @OriolAbril
-Reviewer 2: @alizma

### 2024-01-29 paper/main.qmd
Update paper

@@
 title: "Bibat: Batteries-include Bayesian Analysis Template"
-keywords: Bayesian inference,
 abstract: >
-  Despite their suitability for many scientific problems and the existence of
-  sound theoretical and computational frameworks, adoption of Bayesian workflow
-  in science remains a challenge. One reason for the difficulty is that it is
-  difficult to write software that implements non-trivial Bayesian workflows.
-  We aimed to address this difficulty by developing Bibat, a Python package
-  providing  an interactive template that targets Bayesian statistical analysis
-  projects. Bibat trivialises the otherwise painful process of structuring such
-  a project and integrating together tools that help with individual parts of a
-  Bayesian workflow, without compromising flexibility, scalability,
-  interoperability or reproducibility. Bibat is available on the Python
-  Package index,  documented at <https://bibat.readthedocs.io/> and
-  developed at <https://github.com/teddygroves/bibat/>.
+  Bayesian statistical workflow offers a powerful way to learn from data,
+  but software software projects that implement complex Bayesian workflows
+  in practice are unusual, partly due to the difficulty of orchestrating
+  Bayesian statistical software. Bibat addresses this challenge by providing a
+  full-featured, scalable Bayesian statistical analysis project
+  using an interactive template. Bibat is available on the Python Package index,
+  documented at <https://bibat.readthedocs.io/> and developed at
+  <https://github.com/teddygroves/bibat/>. This paper explains the motivation
+  for bibat, briefly describes intended usage, discusses key design choices,
+  and reviews several examples of bibat's use in scientific applications.
 number-sections: true
@@
     affiliations:
-      - id: biosustain
-        name: DTU
+      - id: 1
+        name: Danish Technical University
         department: The Novo Nordisk Foundation Center for Biosustainability
@@
       corresponding: true
+
 bibliography: bibliography.bib
+
 format:
@@
     embed-resources: true
-  pdf:
+  acm-pdf:
     keep-tex: true
+
+acm-metadata:
+  # comment this out to make submission anonymous
+  # anonymous: true
+
+  # comment this out to build a draft version
+  final: true
+
+  # comment this out to specify detailed document options
+  acmart-options: sigconf, review
+
+  # acm preamble information
+  copyright-year: 2024
+  acm-year: 2024
+  copyright: acmcopyright
+  # doi: XXXXXXX.XXXXXXX
+  conference-acronym: "KDD"
+  conference-name: "KDD"
+  conference-date: Aug 25--29, 2024
+  conference-location: Barcelona, Esp
+  keywords:
+    - Python
+    - Bayesian workflow
+    - Methodology
+  teaser:
+    image: img/workflow.png
+    caption: |
+      Schematic representation of a Bayesian workflow implemented using bibat.
+      The author inspects their analysis's results, edits files corresponding
+      to the boxes on the left, runs the command `make analysis`, then repeats.
+      Note that this workflow is modular, accommodates plurality and results in
+      a final analysis that can be fully reproduced using a single command.
+    description: "DESCRIPTION"
+
 ---
 
-Many areas of science would benefit from replacing traditional statistical
-methodologies with Bayesian workflow, as described in, for example
-[@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
-@gabryVisualizationBayesianWorkflow2019].
+# Introduction: the problem of orchestrating Bayesian workflow software
 
-There are now software tools that address most individual aspects of a Bayesian
-workflow, from fetching and manipulating data to specifying, computing, storing,
-inspecting and documenting statistical inferences. However, writing a software
-project implementing a Bayesian workflow remains challenging due partly to the
-difficulty of orchestrating and configuring these elements.
+The idea that Bayesian statistical analysis comprises not just
+inference, but also specific approaches to related activities like data
+preparation, model design, diagnosis, debugging and criticism, dates
+at least to @boxBayesianInferenceStatistical1992. There is increasing
+scholarly recognition of the need for a holistic view of "Bayesian
+workflow" [@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
+@gabryVisualizationBayesianWorkflow2019]. Relatedly, software tools now
+exist that address most individual aspects of a Bayesian workflow: see
+@strumbeljPresentFutureSoftware for a review of the state of the art.
 
-Bibat is a Python package that addresses this difficulty by providing an
-interactive template for Bayesian workflow projects.
+Unfortunately, currently available tools typically address one, or at most
+a handful, of Bayesian workflow activities; it is left to the individual
+project team to orchestrate the components. Writing software that performs
+this orchestration can be time-consuming and tricky, especially in the common
+scenario where it is not initially clear how many, or what kind of, statistical
+models, datasets, data manipulations or investigations an analysis will require.
 
-# Installation and usage
+Bibat addresses this difficulty by providing a full-featured, high-quality
+Bayesian workflow project that can be extended to implement a wide range of
+statistical analyses.
 
-Bibat is installed by running the command `pip install bibat` and then used by
-running the command `bibat`.
+# How bibat works
 
-This command triggers an interactive form which prompts the user for
-configuration information including project and repository name, author name, a
-short description and choices of open source licence options, documentation
-formats and whether to include tests and continuous integration.
+## Installation and usage
 
-Bibat then creates a folder with the chosen repository name in the current
-working directory, containing code that implements bibat's example analysis.
+Bibat can be installed on the Windows, linux or macos command line on by running
+the command `pip install bibat` and then used by running the command `bibat`.
+This command triggers an interactive form which prompts the user to select a
+range of customisation options. Bibat then creates a new directory containing
+code that implements an example analysis, with customisations reflecting the
+user's choices. This analysis works immediately, and can be reproduced with
+the single command `make analysis` without the need for any further action by
+the user: in this sense bibat comes with "batteries included".
 
-# Documentation
+## Documentation
 
@@
 documentation website includes instructions for getting started, a detailed
-explanation of bibat's concepts, an extended vignette illustrating intended
-usage, full description of the python API and command line interface,
-instructions for contributing and a section discussing accessibility
-considerations.
+explanation of bibat's concepts and an extended vignette illustrating how to
+implement a complex statistical analysis starting from bibat's example analysis
+usage. In addition, the documentation site contains a full description of
+bibat's python API and command line interface, instructions for contributing and
+a section discussing accessibility considerations.
 
-Bibat's documentation website is continuously checked to ensure that it
-implements accessibility best practices.
+# Design choices
 
-# Software
+Bibat's design was informed by these considerations.
 
-Bibat is implemented using Python 3. Its main dependencies are cookiecutter
-[@greenfeldCookiecutter2021], pydantic [@pydanticdevelopersPydantic2022] and
-click [@clickdevelopersClickPythonComposable2022].
+## Accommodate a wide range of statistical analyses
 
-Bibat projects have the following dependencies:
+Bibat projects explicitly allow for plurality at the level of input datasets,
+prepared datasets, statistical models, fitting modes, computation methods and
+analyses. In addition, bibat ensures that there are minimal restrictions on the
+kind of components: for example, datasets need not be singular or tabular, and
+statistical models need not be representable in formula syntax.
 
-* Python 3 [@vanrossumPythonReferenceManual2009] and standard scientific Python
-packages for data manipulation
-* Stan [@carpenterStanProbabilisticProgramming2017] for specifying statistical
-models and performing inferences
-* cmdstanpy [@standevelopmentteamCmdStanPy2022] for interfacing between Python
-and Stan
-* arviz [@kumarArviZUnifiedLibrary2019] for storing and analysing completed
-inferences
-* pydantic and pandera [@niels_bantilan-proc-scipy-2020] for validation
-* make [@stallman1991gnu] for automation.
- * Sphinx [@georgbrandlandthesphinxteamSphinx2022] and Quarto
-[@Allaire_Quarto_2022] for documentation.
+Thanks to these accommodations a project team using bibat should typically not
+need to foresee the ultimate requirements of their analysis before starting
+the project.
 
-Bibat is continuously tested to ensure that it works on the operating systems
-Linux, macOS and Windows.
+## Encourage reproducibility
 
-Bibat's continuous integration runs on windows and ubuntu and runs bibat's test
-suite as well as an end-to-end functional test on all supported Python versions.
-
-# How bibat addresses specific Bayesian workflow difficulties
-
-This section describes some specific problems that often affect Bayesian
-workflow projects and which bibat helps to ameliorate.
-
-## Data modelling
-
-Any statistical analysis requires a definition of a prepared dataset, whether
-implicitly or explicitly. Ideally the definition is explicit and allows for
-arbitrarily many prepared datasets, accommodates both tabular and non-tabular
-formats and provides functionality like validation and serialisation.
-
-Bibat achieves this goal by providing abstract models for prepared data,
-preconfigured to work well together, integrated into a data preparation pipeline
-and able to be easily customised to suit any analysis.
-
-Without a template such as bibat, users must either devise a new data modelling
-framework from scratch as part of an analysis, which is time-consuming, or
-else to miss out on either the flexibility or features that bibat's approach
-provides.
-
-## Reproducibility
-
-Bibat ensures reproducibility by providing a preconfigured makefile with a
-target `analysis` triggering creation of an isolated environment, installation
-of dependencies, data preparation, statistical computation and analysis of
-results. In this way a bibat analysis can be reproduced on most platforms
-using a single command.
+Bibat provides a preconfigured makefile with a target `analysis` triggering
+creation of an isolated environment, installation of dependencies, data
+preparation, statistical computation and analysis of results. In this way a
+bibat analysis can be reproduced on most platforms using a single command.
 
@@
 
-## Collaboration
+## Use widely-adopted, open source and actively developed tools
 
-Bibat provides a preconfigured test environment, continuous integration, linting
-and pre-commit hooks, making it suitable for collaborative software development.
-In addition, including documentation as a first class component of the analysis
-addresses a common problem in academic statistics projects where the paper gets
-out of sync with the code.
+Bibat projects are is written in modern Python and uses pydantic
+@pydanticdevelopersPydantic2022 and pandera [@niels_bantilan-proc-scipy-2020]
+for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
+statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
+python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
+analyzing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
+quarto [@Allaire_Quarto_2022] for documentation.
 
-## Installing dependencies
+Bibat itself is also written in modern Python and uses the popular
+tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
+[@clickdevelopersClickPythonComposable2022].
 
@@
 
+## Implement community standards and best practices for collaborative software development
+
+Bibat projects include a preconfigured test environment, continuous integration,
+linting and pre-commit hooks, making them suitable for collaborative software
+development. In addition, including documentation as a first class component of
+the analysis addresses a common problem in academic statistics projects where
+the paper gets out of sync with the code.
+
+Bibat itself is continuously tested to ensure that it works on the operating
+systems Linux, macOS and Windows. Bibat's continuous integration runs a test
+suite as well as an end-to-end functional test on all supported Python versions.
+
+Bibat is part of the PyOpenSci ecosystem, allowing for community help with
+maintenance as well as peer review for code and documentation quality, usability
+and accessibility. The PyOpenSci peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>
+
+# Challenges
+
+This section describes some specific challenges that often affect Bayesian
+workflow projects and how bibat addresses them.
+
+## Address complexity using a modular, file based approach
+
+As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
+featuring plurality, cyclicity and complexity at many levels. Bibat accommodates
+this by separating non-interacting analysis components into modules and by
+serialising data to files wherever possible. Prepared datasets, statistical
+models, inference configurations, inference results, plots and analyses all
+have file representations. Fitting modes, data manipulations and data models are
+modularised in code through the use of appropriately structured data classes
+and functions.
+
+Thanks to this approach it is possible to perform small sub analyses
+individually and to iteratively add components without needing to consider
+everything at once.
+
 ## Fitting modes
 
-As part of a Bayesian workflow it is often necessary to fit a model and dataset
+As part of a Bayesian workflow it can be necessary to fit a model and dataset
 in different ways. For example, one might perform MCMC sampling of both the
@@
 
-Bibat accommodates this ubiquitous scenario by introducing an abstraction called
-"fitting mode". This abstraction allows bibat projects to handle fitting a
-model and dataset in different ways appropriately and flexibly. For example, the
+Bibat accommodates this need by introducing an abstraction called "fitting
+mode". This abstraction allows bibat projects to handle fitting a model
+and dataset in different ways appropriately and flexibly. For example, the
 provided prior sampling fitting mode creates a Stan input dictionary with the
@@
 to the `InferenceData` group `prior`. Bibat provides fitting modes corresponding
-for prior sampling, posterior sampling and kfold posterior sampling. Users can
+for prior sampling, posterior sampling and k-fold posterior sampling. Users can
 easily add additional fitting modes by copying these examples.
 
-## Configure fitting concisely.
+# Post launch use
 
-Bibat provides a Pydantic class `InferenceConfiguration`, parsed from toml files
-in the `inferences` directory. By writing a toml file per inference the user can
-configure statistical model, dataset, fitting modes and computational settings
-(globally or per fitting mode).
+Bibat has been used for several real statistical analyses.
 
-This approach appropriately separates configuration from logic and avoids
-unnecessary duplication of configuration for runs with the same model and
-dataset.
+@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
+approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
+this analysis even though it was not very large by providing the fitting mode
+abstraction, which was very useful for comparing the different methods.
 
-## Storing MCMC samples
+In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
+two datasets, two models and four inferences, demonstrating that the generalised
+Pareto distribution can be used to describe hitting ability in baseball. This
+analysis is now included in bibat as an illustration, along with an accompanying
+tutorial.
 
-Bibat provides code that automatically saves its output `InferenceData` objects
-in zarr format. This format distributes information from the same set of samples
-into multiple smaller files, thereby helping to avoid file size limits of online
-repository hosting services.
+:::{#fig-baseball}
+
+![](img/baseball.png){width=50%}
+
+A graphical posterior predictive check produced as part of a bibat
+analysis that fit two statistical models to two datasets of baseball data.
+The coloured lines show each model's posterior predictive distributions
+and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
+for the full analysis.
+
+:::
+
+In @grovesteddySphincter2024, bibat was used to implement an analysis of
+cerebrovascular data from mice, involving two raw datasets, 6 prepared
+datasets, 15 models and 15 inferences. In this case bibat's modular design made
+it straightforward to iteratively add data transformations and models while
+maintaining reproducibility.
+
+These cases illustrate that bibat can be useful in a variety of real Bayesian
+workflows, with different sizes, subject matters and emphases.
+
+Bibat has a growing user community, with 16 GitHub stars at the time of writing.
 
@@
 
-There is some software that addresses the general task of facilitating Bayesian
-workflow, but differently from bibat. For example, bambi [@capretto2020]
-and brms [@burknerBrmsPackageBayesian2017] aim to make implementing Bayesian
-workflows easier by providing ergonomic ways to specify and fit Bayesian
-generalised models to tabular datasets. Bibat is complementary with these
-packages, as it targets use cases that they do not support, such as analyses
-where complex datasets or models might be required.
+There is some software that addresses the general task of facilitating
+Bayesian workflow, but using a different approach from bibat's. For example,
+bambi [@capretto2020] and brms [@burknerBrmsPackageBayesian2017] aim to make
+implementing Bayesian workflows easier by providing ergonomic ways to specify
+and fit Bayesian regression models to tabular datasets. Bibat is complementary
+with these packages, as it targets use cases that they do not support, such as
+analyses where complex datasets or custom models might be required.
 
-#
+# Statements and Declarations
+
+The authors have no financial or non-financial interests that are directly or
+indirectly related to the work submitted for publication.
 

### 2024-01-29 paper/main.tex
Update paper

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+\documentclass[sigconf, review]{acmart}
 
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else
-    % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+
+%%
+%% This is file `sample-manuscript.tex',
+%% generated with the docstrip utility.
+%%
+%% The original source files were:
+%%
+%% samples.dtx  (with options: `manuscript')
+%%
+%% IMPORTANT NOTICE:
+%%
+%% For the copyright see the source file.
+%%
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-manuscript.tex.
+%%
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%%
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%%
+%% Commands for TeXCount
+%TC:macro \cite [option:text,text]
+%TC:macro \citep [option:text,text]
+%TC:macro \citet [option:text,text]
+%TC:envir table 0 1
+%TC:envir table* 0 1
+%TC:envir tabular [ignore] word
+%TC:envir displaymath 0 word
+%TC:envir math 0 word
+%TC:envir comment 0 0
+%%
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+
+%% PANDOC PREAMBLE BEGINS
 
@@
 \makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-\KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
+\definecolor{mypink}{RGB}{219, 48, 122}
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+%% PANDOC PREAMBLE ENDS
+
+\setlength{\parindent}{10pt}
+\setlength{\parskip}{0pt}
+
 \hypersetup{
@@
   pdfauthor={Teddy Groves},
-  pdfkeywords={Bayesian inference,},
   colorlinks=true,
@@
   citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
+  urlcolor={red},
+  pdfcreator={LaTeX via pandoc, via quarto}}
 
+%% \BibTeX command to typeset BibTeX logo in the docs
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    Bib\TeX}}}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2024}
+\acmYear{2024}
+\acmDOI{}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[KDD]{KDD}{Aug 25--29, 2024}{Barcelona, Esp}
+\acmPrice{}
+\acmISBN{}
+
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+%%\acmSubmissionID{123-A56-BU3}
+
+%%
+%% For managing citations, it is recommended to use bibliography
+%% files in BibTeX format.
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
+%% end of the preamble, start of the body of the document source.
+\begin{document}
+
+%%
+%% The "title" command has an optional parameter,
+%% allowing the author to define a "short title" to be used in page headers.
 \title{Bibat: Batteries-include Bayesian Analysis Template}
+
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+
   \author{Teddy Groves}
-\date{}
+  \orcid{0000-0002-7109-3270}
+            \affiliation{%
+                  \institution{Danish Technical University}
+                                  \city{Kongens Lyngby}
+                                  \country{Denmark}
+                      }
 
-\begin{document}
+%% By default, the full list of authors will be used in the page
+%% headers. Often, this list is too long, and will overlap
+%% other information printed in the page headers. This command allows
+%% the author to define a more concise list
+%% of authors' names for this purpose.
+%\renewcommand{\shortauthors}{Trovato et al.}
+%%
+%% The abstract is a short summary of the work to be presented in the
+%% article.
+\begin{abstract}
+Bayesian statistical workflow offers a powerful way to learn from data,
+but software software projects that implement complex Bayesian workflows
+in practice are unusual, partly due to the difficulty of orchestrating
+Bayesian statistical software. Bibat addresses this challenge by
+providing a full-featured, scalable Bayesian statistical analysis
+project using an interactive template. Bibat is available on the Python
+Package index, documented at \url{https://bibat.readthedocs.io/} and
+developed at \url{https://github.com/teddygroves/bibat/}. This paper
+explains the motivation for bibat, briefly describes intended usage,
+discusses key design choices, and reviews several examples of bibat's
+use in scientific applications.
+\end{abstract}
+
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+\keywords{Python, Bayesian workflow, Methodology}
+
+\begin{teaserfigure}
+  \includegraphics[width=\textwidth]{img/workflow.png}
+  \caption{Schematic representation of a Bayesian workflow implemented
+using bibat. The author inspects their analysis's results, edits files
+corresponding to the boxes on the left, runs the command
+\texttt{make\ analysis}, then repeats. Note that this workflow is
+modular, accommodates plurality and results in a final analysis that can
+be fully reproduced using a single command.}
+  \Description{DESCRIPTION}
+\end{teaserfigure}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
 \maketitle
-\begin{abstract}
-Despite their suitability for many scientific problems and the existence
-of sound theoretical and computational frameworks, adoption of Bayesian
-workflow in science remains a challenge. One reason for the difficulty
-is that it is difficult to write software that implements non-trivial
-Bayesian workflows. We aimed to address this difficulty by developing
-Bibat, a Python package providing an interactive template that targets
-Bayesian statistical analysis projects. Bibat trivialises the otherwise
-painful process of structuring such a project and integrating together
-tools that help with individual parts of a Bayesian workflow, without
-compromising flexibility, scalability, interoperability or
-reproducibility. Bibat is available on the Python Package index,
-documented at \url{https://bibat.readthedocs.io/} and developed at
-\url{https://github.com/teddygroves/bibat/}.
-\end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
-Many areas of science would benefit from replacing traditional
-statistical methodologies with Bayesian workflow, as described in, for
-example (Gelman et al. 2020; Grinsztajn et al. 2021; Gabry et al. 2019).
+\setlength{\parskip}{-0.1pt}
 
-There are now software tools that address most individual aspects of a
-Bayesian workflow, from fetching and manipulating data to specifying,
-computing, storing, inspecting and documenting statistical inferences.
-However, writing a software project implementing a Bayesian workflow
-remains challenging due partly to the difficulty of orchestrating and
-configuring these elements.
+\section{Introduction: the problem of orchestrating Bayesian workflow
+software}\label{introduction-the-problem-of-orchestrating-bayesian-workflow-software}
 
-Bibat is a Python package that addresses this difficulty by providing an
-interactive template for Bayesian workflow projects.
+The idea that Bayesian statistical analysis comprises not just
+inference, but also specific approaches to related activities like data
+preparation, model design, diagnosis, debugging and criticism, dates at
+least to \citet{boxBayesianInferenceStatistical1992}. There is
+increasing scholarly recognition of the need for a holistic view of
+``Bayesian workflow''
+\citep{gelmanBayesianWorkflow2020, grinsztajnBayesianWorkflowDisease2021, gabryVisualizationBayesianWorkflow2019}.
+Relatedly, software tools now exist that address most individual aspects
+of a Bayesian workflow: see \citet{strumbeljPresentFutureSoftware} for a
+review of the state of the art.
 
-\hypertarget{installation-and-usage}{%
-\section{Installation and usage}\label{installation-and-usage}}
+Unfortunately, currently available tools typically address one, or at
+most a handful, of Bayesian workflow activities; it is left to the
+individual project team to orchestrate the components. Writing software
+that performs this orchestration can be time-consuming and tricky,
+especially in the common scenario where it is not initially clear how
+many, or what kind of, statistical models, datasets, data manipulations
+or investigations an analysis will require.
 
-Bibat is installed by running the command \texttt{pip\ install\ bibat}
-and then used by running the command \texttt{bibat}.
+Bibat addresses this difficulty by providing a full-featured,
+high-quality Bayesian workflow project that can be extended to implement
+a wide range of statistical analyses.
 
-This command triggers an interactive form which prompts the user for
-configuration information including project and repository name, author
-name, a short description and choices of open source licence options,
-documentation formats and whether to include tests and continuous
-integration.
+\section{How bibat works}\label{how-bibat-works}
 
-Bibat then creates a folder with the chosen repository name in the
-current working directory, containing code that implements bibat's
-example analysis.
+\subsection{Installation and usage}\label{installation-and-usage}
 
-\hypertarget{documentation}{%
-\section{Documentation}\label{documentation}}
+Bibat can be installed on the Windows, linux or macos command line on by
+running the command \texttt{pip\ install\ bibat} and then used by
+running the command \texttt{bibat}. This command triggers an interactive
+form which prompts the user to select a range of customisation options.
+Bibat then creates a new directory containing code that implements an
+example analysis, with customisations reflecting the user's choices.
+This analysis works immediately, and can be reproduced with the single
+command \texttt{make\ analysis} without the need for any further action
+by the user: in this sense bibat comes with ``batteries included''.
+
+\subsection{Documentation}\label{documentation}
 
@@
 documentation website includes instructions for getting started, a
-detailed explanation of bibat's concepts, an extended vignette
-illustrating intended usage, full description of the python API and
-command line interface, instructions for contributing and a section
-discussing accessibility considerations.
+detailed explanation of bibat's concepts and an extended vignette
+illustrating how to implement a complex statistical analysis starting
+from bibat's example analysis usage. In addition, the documentation site
+contains a full description of bibat's python API and command line
+interface, instructions for contributing and a section discussing
+accessibility considerations.
 
-\hypertarget{software}{%
-\section{Software}\label{software}}
+\section{Design choices}\label{design-choices}
 
-Bibat is implemented using Python 3. Its main dependencies are
-cookiecutter (Greenfeld et al. 2021), pydantic (Pydantic developers
-2022) and click (Click Developers 2022). Bibat is continuously tested to
-ensure that it works on the operating systems Linux, macOS and Windows.
+Bibat's design was informed by these considerations.
 
-Bibat projects use the following tools:
+\subsection{Accommodate a wide range of statistical
+analyses}\label{accommodate-a-wide-range-of-statistical-analyses}
 
-\begin{itemize}
-\tightlist
-\item
-  Python 3 (Van Rossum and Drake 2009) and standard scientific Python
-  packages for data manipulation
-\item
-  Stan (Carpenter et al. 2017) for specifying statistical models and
-  performing inferences
-\item
-  cmdstanpy (Stan Development Team 2022) for interfacing between Python
-  and Stan
-\item
-  arviz (Kumar et al. 2019) for storing and analysing completed
-  inferences
-\item
-  pydantic and pandera (Niels Bantilan 2020) for validation
-\item
-  make (Stallman and McGrath 1991) for automation.
-\item
-  Sphinx (Georg Brandl and the Sphinx team 2022) and Quarto (Allaire et
-  al. 2022) for documentation.
-\end{itemize}
+Bibat projects explicitly allow for plurality at the level of input
+datasets, prepared datasets, statistical models, fitting modes,
+computation methods and analyses. In addition, bibat ensures that there
+are minimal restrictions on the kind of components: for example,
+datasets need not be singular or tabular, and statistical models need
+not be representable in formula syntax.
 
-\hypertarget{how-bibat-addresses-specific-bayesian-workflow-difficulties}{%
-\section{How bibat addresses specific Bayesian workflow
-difficulties}\label{how-bibat-addresses-specific-bayesian-workflow-difficulties}}
+Thanks to these accommodations a project team using bibat should
+typically not need to foresee the ultimate requirements of their
+analysis before starting the project.
 
-This section describes some specific problems that often affect Bayesian
-workflow projects and which bibat helps to ameliorate.
+\subsection{Encourage reproducibility}\label{encourage-reproducibility}
 
-\hypertarget{data-modelling}{%
-\subsection{Data modelling}\label{data-modelling}}
-
-Any statistical analysis requires a definition of a prepared dataset,
-whether implicitly or explicitly. Ideally the definition is explicit and
-allows for arbitrarily many prepared datasets, accommodates both tabular
-and non-tabular formats and provides functionality like validation and
-serialisation.
-
-Bibat achieves this goal by providing abstract models for prepared data,
-preconfigured to work well together, integrated into a data preparation
-pipeline and able to be easily customised to suit any analysis.
-
-Without a template such as bibat, users must either devise a new data
-modelling framework from scratch as part of an analysis, which is
-time-consuming, or else to miss out on either the flexibility or
-features that bibat's approach provides.
-
-\hypertarget{reproducibility}{%
-\subsection{Reproducibility}\label{reproducibility}}
-
-Bibat ensures reproducibility by providing a preconfigured makefile with
-a target \texttt{analysis} triggering creation of an isolated
-environment, installation of dependencies, data preparation, statistical
-computation and analysis of results. In this way a bibat analysis can be
-reproduced on most platforms using a single command.
+Bibat provides a preconfigured makefile with a target \texttt{analysis}
+triggering creation of an isolated environment, installation of
+dependencies, data preparation, statistical computation and analysis of
+results. In this way a bibat analysis can be reproduced on most
+platforms using a single command.
 
@@
 
-\hypertarget{collaboration}{%
-\subsection{Collaboration}\label{collaboration}}
+\subsection{Use widely-adopted, open source and actively developed
+tools}\label{use-widely-adopted-open-source-and-actively-developed-tools}
 
-Bibat provides a preconfigured test environment, continuous integration,
-linting and pre-commit hooks, making it suitable for collaborative
-software development. In addition, including documentation as a first
-class component of the analysis addresses a common problem in academic
-statistics projects where the paper gets out of sync with the code.
+Bibat projects are is written in modern Python and uses pydantic
+\citet{pydanticdevelopersPydantic2022} and pandera
+\citep{niels_bantilan-proc-scipy-2020} for data modelling, Stan
+\citep{carpenterStanProbabilisticProgramming2017} for statistical
+inference, cmdstanpy \citep{standevelopmentteamCmdStanPy2022} for
+python-Stan interface, arviz \citep{kumarArviZUnifiedLibrary2019} for
+storing and analyzing inferences and sphinx
+\citep{georgbrandlandthesphinxteamSphinx2022} and quarto
+\citep{Allaire_Quarto_2022} for documentation.
 
-\hypertarget{installing-dependencies}{%
-\subsection{Installing dependencies}\label{installing-dependencies}}
+Bibat itself is also written in modern Python and uses the popular tools
+cookiecutter \citep{greenfeldCookiecutter2021}, pydantic and click
+\citep{clickdevelopersClickPythonComposable2022}.
 
@@
 
-\hypertarget{fitting-modes}{%
-\subsection{Fitting modes}\label{fitting-modes}}
+\subsection{Implement community standards and best practices for
+collaborative software
+development}\label{implement-community-standards-and-best-practices-for-collaborative-software-development}
 
-As part of a Bayesian workflow it is often necessary to fit a model and
+Bibat projects include a preconfigured test environment, continuous
+integration, linting and pre-commit hooks, making them suitable for
+collaborative software development. In addition, including documentation
+as a first class component of the analysis addresses a common problem in
+academic statistics projects where the paper gets out of sync with the
+code.
+
+Bibat itself is continuously tested to ensure that it works on the
+operating systems Linux, macOS and Windows. Bibat's continuous
+integration runs a test suite as well as an end-to-end functional test
+on all supported Python versions.
+
+Bibat is part of the PyOpenSci ecosystem, allowing for community help
+with maintenance as well as peer review for code and documentation
+quality, usability and accessibility. The PyOpenSci peer review for
+bibat can be found here:
+\url{https://github.com/pyOpenSci/software-submission/issues/83}
+
+\section{Challenges}\label{challenges}
+
+This section describes some specific challenges that often affect
+Bayesian workflow projects and how bibat addresses them.
+
+\subsection{Address complexity using a modular, file based
+approach}\label{address-complexity-using-a-modular-file-based-approach}
+
+As discussed in \citet{gelmanBayesianWorkflow2020}, Bayesian workflows
+are complicated, featuring plurality, cyclicity and complexity at many
+levels. Bibat accommodates this by separating non-interacting analysis
+components into modules and by serialising data to files wherever
+possible. Prepared datasets, statistical models, inference
+configurations, inference results, plots and analyses all have file
+representations. Fitting modes, data manipulations and data models are
+modularised in code through the use of appropriately structured data
+classes and functions.
+
+Thanks to this approach it is possible to perform small sub analyses
+individually and to iteratively add components without needing to
+consider everything at once.
+
+\subsection{Fitting modes}\label{fitting-modes}
+
+As part of a Bayesian workflow it can be necessary to fit a model and
 dataset in different ways. For example, one might perform MCMC sampling
@@
 
-Bibat accommodates this ubiquitous scenario by introducing an
-abstraction called ``fitting mode'' and a corresponding Pydantic base
-class \texttt{FittingMode}, along with and several subclasses for
-commonly used cases.
+Bibat accommodates this need by introducing an abstraction called
+``fitting mode''. This abstraction allows bibat projects to handle
+fitting a model and dataset in different ways appropriately and
+flexibly. For example, the provided prior sampling fitting mode creates
+a Stan input dictionary with the \texttt{likelihood} data variable set
+to \texttt{0}, performs MCMC sampling and writes data to the
+\texttt{InferenceData} group \texttt{prior}. Bibat provides fitting
+modes corresponding for prior sampling, posterior sampling and k-fold
+posterior sampling. Users can easily add additional fitting modes by
+copying these examples.
 
-This abstraction allows bibat projects to handle fitting a model and
-dataset in different ways appropriately and flexibly. For example, the
-provided prior sampling fitting mode creates a Stan input dictionary
-with the \texttt{likelihood} data variable set to \texttt{0}, performs
-MCMC sampling and writes data to the \texttt{InferenceData} group
-\texttt{prior}.
+\section{Post launch use}\label{post-launch-use}
 
-\hypertarget{configure-fitting-concisely.}{%
-\subsection{Configure fitting
-concisely.}\label{configure-fitting-concisely.}}
+Bibat has been used for several real statistical analyses.
 
-Bibat provides a Pydantic class \texttt{InferenceConfiguration}, parsed
-from toml files in the \texttt{inferences} directory. By writing a toml
-file per inference the user can configure statistical model, dataset,
-fitting modes and computational settings (globally or per fitting mode).
+\citet{grovesteddyDgfreg2023} used bibat to compare a Bayesian and two
+non-Bayesian approaches to modelling a biochemical thermodynamics
+dataset. Bibat facilitated this analysis even though it was not very
+large by providing the fitting mode abstraction, which was very useful
+for comparing the different methods.
 
-This approach appropriately separates configuration from logic and
-avoids unnecessary duplication of configuration for runs with the same
-model and dataset.
+In \citet{grovesBaseball2022}, Bibat was used to implement a sports
+analysis involving two datasets, two models and four inferences,
+demonstrating that the generalised Pareto distribution can be used to
+describe hitting ability in baseball. This analysis is now included in
+bibat as an illustration, along with an accompanying tutorial.
 
-\hypertarget{storing-mcmc-samples}{%
-\subsection{Storing MCMC samples}\label{storing-mcmc-samples}}
+\begin{figure}
 
-Bibat provides code that automatically saves its output
-\texttt{InferenceData} objects in zarr format. This format distributes
-information from the same set of samples into multiple smaller files,
-thereby helping to avoid file size limits of online repository hosting
-services.
+\centering{
 
-\hypertarget{comparison-with-alternative-software}{%
+\includegraphics[width=0.5\textwidth,height=\textheight]{img/baseball.png}
+
+}
+
+\caption{\label{fig-baseball}A graphical posterior predictive check
+produced as part of a bibat analysis that fit two statistical models to
+two datasets of baseball data. The coloured lines show each model's
+posterior predictive distributions and the black dots show the two
+observed datasets. See
+\url{https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball}
+for the full analysis.}
+
+\end{figure}%
+
+In \citet{grovesteddySphincter2024}, bibat was used to implement an
+analysis of cerebrovascular data from mice, involving two raw datasets,
+6 prepared datasets, 15 models and 15 inferences. In this case bibat's
+modular design made it straightforward to iteratively add data
+transformations and models while maintaining reproducibility.
+
+These cases illustrate that bibat can be useful in a variety of real
+Bayesian workflows, with different sizes, subject matters and emphases.
+
+Bibat has a growing user community, with 16 GitHub stars at the time of
+writing.
+
 \section{Comparison with alternative
-software}\label{comparison-with-alternative-software}}
+software}\label{comparison-with-alternative-software}
 
@@
 templates that arguably encompass Bayesian workflow as a special case of
-data analysis project, such as cookiecutter-data-science (Driven Data
-2022), but these are of limited use compared with a specialised template
-due to the many specificities of Bayesian workflow.
+data analysis project, such as cookiecutter-data-science
+\citep{drivendataCookiecutterdatascience2022}, but these are of limited
+use compared with a specialised template due to the many specificities
+of Bayesian workflow.
 
 There is some software that addresses the general task of facilitating
-Bayesian workflow, but differently from bibat. For example, bambi
-(Capretto et al. 2020) and brms (Bürkner 2017) aim to make implementing
-Bayesian workflows easier by trivialising the task of specifying a
-Bayesian generalised linear model and fitting it to a single tabular
-dataset. This approach does not address several difficulties that bibat
-does address,
+Bayesian workflow, but using a different approach from bibat's. For
+example, bambi \citep{capretto2020} and brms
+\citep{burknerBrmsPackageBayesian2017} aim to make implementing Bayesian
+workflows easier by providing ergonomic ways to specify and fit Bayesian
+regression models to tabular datasets. Bibat is complementary with these
+packages, as it targets use cases that they do not support, such as
+analyses where complex datasets or custom models might be required.
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{Statements and Declarations}\label{statements-and-declarations}
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Allaire_Quarto_2022}{}}%
-Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
-Christophe Dervieux. 2022. {``Quarto.''}
-\url{https://doi.org/10.5281/zenodo.5960048}.
+The authors have no financial or non-financial interests that are
+directly or indirectly related to the work submitted for publication.
 
-\leavevmode\vadjust pre{\hypertarget{ref-burknerBrmsPackageBayesian2017}{}}%
-Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
-Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
-Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bibliography.bib}
 
-\leavevmode\vadjust pre{\hypertarget{ref-capretto2020}{}}%
-Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
-and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
-{Bayesian} Linear Models in {Python}.''}
-\url{https://arxiv.org/abs/2012.10754}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-carpenterStanProbabilisticProgramming2017}{}}%
-Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
-Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
-and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
-Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
-\url{https://doi.org/10.18637/jss.v076.i01}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-clickdevelopersClickPythonComposable2022}{}}%
-Click Developers. 2022. {``Click: {Python} Composable Command Line
-Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-drivendataCookiecutterdatascience2022}{}}%
-Driven Data. 2022. {``Cookiecutter-Data-Science.''}
-\url{https://github.com/drivendata/cookiecutter-data-science/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gabryVisualizationBayesianWorkflow2019}{}}%
-Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
-Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
-\emph{Journal of the Royal Statistical Society Series A: Statistics in
-Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelmanBayesianWorkflow2020}{}}%
-Gelman, Andrew, Aki Vehtari, Daniel Simpson, Charles C. Margossian, Bob
-Carpenter, Yuling Yao, Lauren Kennedy, Jonah Gabry, Paul-Christian
-Bürkner, and Martin Modrák. 2020. {``Bayesian {Workflow}.''}
-\emph{arXiv:2011.01808 {[}Stat{]}}, November.
-\url{http://arxiv.org/abs/2011.01808}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-georgbrandlandthesphinxteamSphinx2022}{}}%
-Georg Brandl and the Sphinx team. 2022. {``Sphinx.''}
-\url{https://www.sphinx-doc.org/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-greenfeldCookiecutter2021}{}}%
-Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
-2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-grinsztajnBayesianWorkflowDisease2021}{}}%
-Grinsztajn, Léo, Elizaveta Semenova, Charles C. Margossian, and Julien
-Riou. 2021. {``Bayesian Workflow for Disease Transmission Modeling in
-{Stan}.''} \emph{Statistics in Medicine} 40 (27): 6209--34.
-\url{https://doi.org/10.1002/sim.9164}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kumarArviZUnifiedLibrary2019}{}}%
-Kumar, Ravin, Colin Carroll, Ari Hartikainen, and Osvaldo Martin. 2019.
-{``{ArviZ} a Unified Library for Exploratory Analysis of {Bayesian}
-Models in {Python}.''} \emph{Journal of Open Source Software} 4 (33):
-1143. \url{https://doi.org/10.21105/joss.01143}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-niels_bantilan-proc-scipy-2020}{}}%
-Niels Bantilan. 2020. {``Pandera: {Statistical Data Validation} of
-{Pandas Dataframes}.''} In \emph{Proceedings of the 19th {Python} in
-{Science Conference}}, edited by Meghann Agarwal, Chris Calloway, Dillon
-Niederhut, and David Shupe, 116--24.
-\url{https://doi.org/10.25080/Majora-342d178e-010}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pydanticdevelopersPydantic2022}{}}%
-Pydantic developers. 2022. {``Pydantic.''}
-\url{https://pypi.org/project/pydantic/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-stallman1991gnu}{}}%
-Stallman, Richard M, and Roland McGrath. 1991. {``{GNU} Make-{A} Program
-for Directing Recompilation.''}
-
-\leavevmode\vadjust pre{\hypertarget{ref-standevelopmentteamCmdStanPy2022}{}}%
-Stan Development Team. 2022. {``{CmdStanPy}.''}
-\url{https://github.com/stan-dev/cmdstanpy}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vanrossumPythonReferenceManual2009}{}}%
-Van Rossum, Guido, and Fred L. Drake. 2009. \emph{Python 3 {Reference
-Manual}}. {Scotts Valley, CA}: {CreateSpace}.
-
-\end{CSLReferences}
+%% begin pandoc before-bib
+%% end pandoc before-bib
+%% begin pandoc biblio
+%% end pandoc biblio
+%% begin pandoc include-after
+%% end pandoc include-after
+%% begin pandoc after-body
+%% end pandoc after-body
 
 \end{document}
+\endinput
+%%
+%% End of file `sample-manuscript.tex'.

### 2024-01-30 paper/main.qmd
Update paper

@@
   in practice are unusual, partly due to the difficulty of orchestrating
-  Bayesian statistical software. Bibat addresses this challenge by providing a
-  full-featured, scalable Bayesian statistical analysis project
-  using an interactive template. Bibat is available on the Python Package index,
-  documented at <https://bibat.readthedocs.io/> and developed at
-  <https://github.com/teddygroves/bibat/>. This paper explains the motivation
-  for bibat, briefly describes intended usage, discusses key design choices,
-  and reviews several examples of bibat's use in scientific applications.
+  Bayesian statistical software. Bibat addresses this challenge by providing
+  a full-featured, scalable Bayesian statistical analysis project using
+  an interactive template. Bibat is available on the Python Package index,
+  documented at <https://bibat.readthedocs.io/> and developed at <https://github.com/teddygroves/bibat/>. This paper explains the motivation for bibat,
+  describes intended usage, discusses bibat's design, compares bibat with similar software, highlights several examples of bibat's use in science and provides links to community resources associated with bibat.
 number-sections: true
@@
       corresponding: true
-
 bibliography: bibliography.bib
-
+keywords: [Bayesian workflow, Methodology, Software]
 format:
+  typst: default
   html:
     embed-resources: true
-  acm-pdf:
-    keep-tex: true
-
-acm-metadata:
-  # comment this out to make submission anonymous
-  # anonymous: true
-
-  # comment this out to build a draft version
-  final: true
-
-  # comment this out to specify detailed document options
-  acmart-options: sigconf, review
-
-  # acm preamble information
-  copyright-year: 2024
-  acm-year: 2024
-  copyright: acmcopyright
-  # doi: XXXXXXX.XXXXXXX
-  conference-acronym: "KDD"
-  conference-name: "KDD"
-  conference-date: Aug 25--29, 2024
-  conference-location: Barcelona, Esp
-  keywords:
-    - Python
-    - Bayesian workflow
-    - Methodology
-  teaser:
-    image: img/workflow.png
-    caption: |
-      Schematic representation of a Bayesian workflow implemented using bibat.
-      The author inspects their analysis's results, edits files corresponding
-      to the boxes on the left, runs the command `make analysis`, then repeats.
-      Note that this workflow is modular, accommodates plurality and results in
-      a final analysis that can be fully reproduced using a single command.
-    description: "DESCRIPTION"
-
+  latex: default
 ---
@@
 
-Bibat can be installed on the Windows, linux or macos command line on by running
+Bibat can be installed on the Windows, Linux or macOS command line on by running
 the command `pip install bibat` and then used by running the command `bibat`.
@@
 the single command `make analysis` without the need for any further action by
-the user: in this sense bibat comes with "batteries included".
+the user: in this sense bibat comes with batteries included.
+
+@fig-workflow illustrates the components of a bibat-based Bayesian workflow and
+shows how it proceeds: the project team edits the code components, then runs
+`make analysis` triggering creation of the result components. After inspecting
+these they repeat the process, leading to a cycle that ultimately results in a
+complete, easily reproducible analysis.
+
+:::{#fig-workflow}
+
+![](img/workflow.png){width=100%}
+
+Schematic representation of a Bayesian workflow implemented using bibat. The
+author inspects their analysis's results, edits code corresponding to the
+boxes on the left, runs the command `make analysis`, then repeats. The diagram
+illustrates several key features of bibat: inference components are modular and
+plural, the overall workflow is iterative and cyclical and the whole analysis
+can be executed with a single command.
+
+:::
 
@@
 
-Bibat's design was informed by these considerations.
+Bibat's design was informed by the aims to accommodate the many sources of
+complexity and in a Bayesian workflow project, to ensure easy reproducibility
+and to integrate many open-source, widely-adopted and powerful Bayesian workflow
+tools and to encourage collaborative development.
 
-## Accommodate a wide range of statistical analyses
-
-Bibat projects explicitly allow for plurality at the level of input datasets,
-prepared datasets, statistical models, fitting modes, computation methods and
-analyses. In addition, bibat ensures that there are minimal restrictions on the
-kind of components: for example, datasets need not be singular or tabular, and
-statistical models need not be representable in formula syntax.
-
+As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
+featuring plurality, cyclicity and complexity at many levels. As a specialised
+Bayesian workflow template, a key goal for bibat was to manage this complexity.
+Bibat achieves this aim by separating non-interacting analysis components into
+separate, potentially plural modules and by serialising data to files wherever
+possible. Prepared datasets, statistical models, inference configurations,
+inference results, plots and analyses all have file representations. Fitting
+modes, data manipulations and data models are modularised in code through the
+use of appropriately structured data classes and functions. Thanks to this
+modular approach it is possible to perform small sub analyses individually
+and to iteratively expand the analysis by adding components without needing to
+consider everything at once. In addition, bibat ensures that there are minimal
+restrictions on the components: for example, datasets need not be singular or
+tabular and statistical models need not be representable in formula syntax.
 Thanks to these accommodations a project team using bibat should typically not
@@
 
-## Encourage reproducibility
+Bibat encourages reproducibility mainly by providing a preconfigured makefile
+with a target `analysis` that triggers creation of an isolated environment,
+installation of dependencies, data preparation, statistical computation and
+analysis of results. In this way a bibat analysis can be reproduced on most
+platforms using a single command. In particular, this target attempts to install
+cmdstan if necessary, using a recipe tailored to the host operating system.
+This functionality addresses a common issue where researchers find it difficult
+to install Stan, especially on Windows. A second way in which bibat encourages
+reproducibility is by providing its Python code in the form of a preconfigured
+package. Thanks to the use of modern conventions for specifying dependencies
+and configuring tooling, we expect that this will make bibat analyses easier
+for other researchers to work with, thereby facilitating replication while also
+making initial development more convenient.
 
-Bibat provides a preconfigured makefile with a target `analysis` triggering
-creation of an isolated environment, installation of dependencies, data
-preparation, statistical computation and analysis of results. In this way a
-bibat analysis can be reproduced on most platforms using a single command.
-
-Bibat also provides its Python code in the form of a package configured using
-modern conventions for specifying dependencies and configuring tooling, so that
-it is easy to maintain reproducibility as the analysis develops.
-
-## Use widely-adopted, open source and actively developed tools
-
-Bibat projects are is written in modern Python and uses pydantic
-@pydanticdevelopersPydantic2022 and pandera [@niels_bantilan-proc-scipy-2020]
-for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
-statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
-python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
-analyzing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
-quarto [@Allaire_Quarto_2022] for documentation.
-
-Bibat itself is also written in modern Python and uses the popular
-tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
+Bibat integrates many widely-adopted, open-source and powerful tools to
+implement the components of a Bayesian workflow. Bibat projects are is
+written in modern Python and uses pydantic @pydanticdevelopersPydantic2022
+and pandera [@niels_bantilan-proc-scipy-2020] for data modelling, Stan
+[@carpenterStanProbabilisticProgramming2017] for statistical inference,
+cmdstanpy [@standevelopmentteamCmdStanPy2022] for python-Stan interface, arviz
+[@kumarArviZUnifiedLibrary2019] for storing and analyzing inferences and sphinx
+[@georgbrandlandthesphinxteamSphinx2022] and quarto [@Allaire_Quarto_2022]
+for documentation.  Bibat itself is also written in modern Python and uses the
+popular tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
 [@clickdevelopersClickPythonComposable2022].
 
-Bibat's makefile detects the current operating system and attempts to install
-cmdstan appropriately if necessary. This functionality addresses a common
-issue where researchers find it difficult to install Stan, especially on Windows.
+To encourage collaborative development of Bayesian workflow projects, Bibat
+projects include a preconfigured test environment, continuous integration,
+linting and pre-commit hooks. In addition, including documentation as a
+first class component of the analysis addresses a common problem in academic
+statistics projects where the paper gets out of sync with the code. Bibat itself
+is continuously tested to ensure that it works on the operating systems Linux,
+macOS and Windows. Bibat's continuous integration runs a test suite as well as
+an end-to-end functional test on all supported Python versions.
 
-## Implement community standards and best practices for collaborative software development
+# Fitting modes
 
-Bibat projects include a preconfigured test environment, continuous integration,
-linting and pre-commit hooks, making them suitable for collaborative software
-development. In addition, including documentation as a first class component of
-the analysis addresses a common problem in academic statistics projects where
-the paper gets out of sync with the code.
+The most novel part of bibat's design is the introduction of an abstraction
+called "fitting mode", which allows bibat project to handle different ways
+of fitting a model to a dataset appropriately and flexibly.  This is often
+necessary as part of a Bayesian workflow: for example, one might perform
+MCMC sampling of both the prior and posterior distributions, perform multiple
+leave-out-one-fold fits for cross-validation or need to compare MCMC sampling
+with an optimisation-based alternative.
 
-Bibat itself is continuously tested to ensure that it works on the operating
-systems Linux, macOS and Windows. Bibat's continuous integration runs a test
-suite as well as an end-to-end functional test on all supported Python versions.
+Fitting modes in bibat projects take the form of instances of the class
+`FittingMode`. Each fitting mode contains a name, a function that fits a
+prepared dataset and instructions for how and where to save the results.
+For example, the provided prior sampling fitting mode creates a Stan input
+dictionary with the `likelihood` data variable set to `0`, performs MCMC
+sampling and writes data to the `InferenceData` group `prior`. This design
+allows multiple fits to be stored alongside each other naturally, facilitating
+many common Bayesian workflow analyses, and can easily be configured to prevent
+unnecessary computations.
 
-Bibat is part of the PyOpenSci ecosystem, allowing for community help with
-maintenance as well as peer review for code and documentation quality, usability
-and accessibility. The PyOpenSci peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>
-
-# Challenges
-
-This section describes some specific challenges that often affect Bayesian
-workflow projects and how bibat addresses them.
-
-## Address complexity using a modular, file based approach
-
-As discussed in @gelmanBayesianWorkflow2020, Bayesian workflows are complicated,
-featuring plurality, cyclicity and complexity at many levels. Bibat accommodates
-this by separating non-interacting analysis components into modules and by
-serialising data to files wherever possible. Prepared datasets, statistical
-models, inference configurations, inference results, plots and analyses all
-have file representations. Fitting modes, data manipulations and data models are
-modularised in code through the use of appropriately structured data classes
-and functions.
-
-Thanks to this approach it is possible to perform small sub analyses
-individually and to iteratively add components without needing to consider
-everything at once.
-
-## Fitting modes
-
-As part of a Bayesian workflow it can be necessary to fit a model and dataset
-in different ways. For example, one might perform MCMC sampling of both the
-prior and posterior distributions, perform multiple leave-out-one-fold fits for
-cross-validation or need to compare MCMC sampling with an optimisation-based
-alternative.
-
-Bibat accommodates this need by introducing an abstraction called "fitting
-mode". This abstraction allows bibat projects to handle fitting a model
-and dataset in different ways appropriately and flexibly. For example, the
-provided prior sampling fitting mode creates a Stan input dictionary with the
-`likelihood` data variable set to `0`, performs MCMC sampling and writes data
-to the `InferenceData` group `prior`. Bibat provides fitting modes corresponding
-for prior sampling, posterior sampling and k-fold posterior sampling. Users can
-easily add additional fitting modes by copying these examples.
-
-# Post launch use
-
-Bibat has been used for several real statistical analyses.
-
-@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
-approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
-this analysis even though it was not very large by providing the fitting mode
-abstraction, which was very useful for comparing the different methods.
-
-In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
-two datasets, two models and four inferences, demonstrating that the generalised
-Pareto distribution can be used to describe hitting ability in baseball. This
-analysis is now included in bibat as an illustration, along with an accompanying
-tutorial.
-
-:::{#fig-baseball}
-
-![](img/baseball.png){width=50%}
-
-A graphical posterior predictive check produced as part of a bibat
-analysis that fit two statistical models to two datasets of baseball data.
-The coloured lines show each model's posterior predictive distributions
-and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
-for the full analysis.
-
-:::
-
-In @grovesteddySphincter2024, bibat was used to implement an analysis of
-cerebrovascular data from mice, involving two raw datasets, 6 prepared
-datasets, 15 models and 15 inferences. In this case bibat's modular design made
-it straightforward to iteratively add data transformations and models while
-maintaining reproducibility.
-
-These cases illustrate that bibat can be useful in a variety of real Bayesian
-workflows, with different sizes, subject matters and emphases.
-
-Bibat has a growing user community, with 16 GitHub stars at the time of writing.
+Bibat provides fitting modes corresponding for prior sampling, posterior
+sampling and k-fold posterior sampling. Users can easily add additional fitting
+modes by copying these examples, or modify the `FittingMode` class to achieve
+even richer functionality.
 
@@
 
-# Statements and Declarations
+# Case studies
 
-The authors have no financial or non-financial interests that are directly or
-indirectly related to the work submitted for publication.
+The following cases illustrate how bibat has been used in practice to facilitate
+Bayesian workflow projects.
 
-# References
+@grovesteddyDgfreg2023 used bibat to compare a Bayesian and two non-Bayesian
+approaches to modelling a biochemical thermodynamics dataset. Bibat facilitated
+this analysis even though it was not very large---the final analysis contained
+one dataset, three models and three inferences---because the fitting mode
+abstraction allowed for straightforward comparison of the different methods.
+Additionally, bibat made it easier to iteratively investigate and discard models
+that did not form part of the final analysis.
+
+In @grovesBaseball2022, Bibat was used to implement a sports analysis involving
+two datasets, two models and four inferences, demonstrating that the generalised
+Pareto distribution can be used to describe hitting ability in baseball. This
+analysis is now included in bibat as an illustration, along with an accompanying
+tutorial. An illustrative graphic from this analysis is shown in @fig-baseball.
+
+In this case bibat was useful because of its ability to implement arbitrary
+statistical models, as latent generalised Pareto distributions are not supported
+by generic regression packages like brms and bambi. Further, bibat's modular
+design made it easier to implement this medium-sized analysis with two datasets,
+two models and six inferences.
+
+:::{#fig-baseball}
+
+![](img/baseball.png){width=100%}
+
+A graphical posterior predictive check produced as part of a bibat
+analysis that fit two statistical models to two datasets of baseball data.
+The coloured lines show each model's posterior predictive distributions
+and the black dots show the two observed datasets.  See <https://github.com/teddygroves/bibat/tree/main/bibat/examples/baseball>
+for the full analysis.
+
+:::
+
+In @grovesteddySphincter2024, bibat was used to implement a large analysis
+of cerebrovascular data from mice, involving two raw datasets, 6 prepared
+datasets, 15 models and 15 inferences. Again it was not possible to implement
+the analysis using formula-based packages due to the need for custom modelling
+approaches to take into account expert knowledge about both the target system
+and the measurement apparatus. Due to the number of datasets and models that
+needed to be considered, implementing this analysis without using bibat or a
+similar structured approach would have required a combination of more time and/
+or compromises in software quality.
+
+These cases illustrate that bibat can be useful in a variety of real Bayesian
+workflows, with different sizes, subject matters and emphases.
+
+# Community
+
+Bibat is developed in public and encourages community contribution.
+Please see the contributing page <https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst> and code of conduct <https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md> if you would like to help develop bibat.
+
+Bibat has a growing user community, with 16 GitHub stars at the time of
+writing, and is affiliated with cmdstanpy through a link on its documentation
+website. Bibat is also affiliated with the Python scientific software community
+PyOpenSci, allowing for help with maintenance as well as peer review for
+code and documentation quality, usability and accessibility. The PyOpenSci
+peer review for bibat can be found here: <https://github.com/pyOpenSci/software-submission/issues/83>.
+
+# Author contributions
+
+# Acknowledgements

### 2024-01-30 paper/main.tex
Update paper

@@
-\documentclass[sigconf, review]{acmart}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
 
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \makeatother
-
-%%
-%% This is file `sample-manuscript.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `manuscript')
-%%
-%% IMPORTANT NOTICE:
-%%
-%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-manuscript.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
-
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-
-%% PANDOC PREAMBLE BEGINS
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
+  % set entry spacing
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-\definecolor{mypink}{RGB}{219, 48, 122}
+\KOMAoption{captions}{tableheading}
 \makeatletter
@@
 \makeatother
-%% PANDOC PREAMBLE ENDS
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
 
-\setlength{\parindent}{10pt}
-\setlength{\parskip}{0pt}
-
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
 \hypersetup{
@@
   pdfauthor={Teddy Groves},
+  pdfkeywords={Bayesian workflow, Methodology, Software},
   colorlinks=true,
@@
   citecolor={Blue},
-  urlcolor={red},
-  pdfcreator={LaTeX via pandoc, via quarto}}
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
+\title{Bibat: Batteries-include Bayesian Analysis Template}
+\author{Teddy Groves}
+\date{}
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2024}
-\acmYear{2024}
-\acmDOI{}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[KDD]{KDD}{Aug 25--29, 2024}{Barcelona, Esp}
-\acmPrice{}
-\acmISBN{}
-
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%% end of the preamble, start of the body of the document source.
 \begin{document}
-
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{Bibat: Batteries-include Bayesian Analysis Template}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-
-  \author{Teddy Groves}
-  \orcid{0000-0002-7109-3270}
-            \affiliation{%
-                  \institution{Danish Technical University}
-                                  \city{Kongens Lyngby}
-                                  \country{Denmark}
-                      }
-
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato et al.}
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
+\maketitle
 \begin{abstract}
@@
 developed at \url{https://github.com/teddygroves/bibat/}. This paper
-explains the motivation for bibat, briefly describes intended usage,
-discusses key design choices, and reviews several examples of bibat's
-use in scientific applications.
+explains the motivation for bibat, describes intended usage, discusses
+bibat's design goals and how key challenges that were overcome in the
+course of development, compares bibat with similar software, highlights
+several examples of bibat's use in science and provides links to
+community resources associated with bibat.
 \end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Python, Bayesian workflow, Methodology}
-
-\begin{teaserfigure}
-  \includegraphics[width=\textwidth]{img/workflow.png}
-  \caption{Schematic representation of a Bayesian workflow implemented
-using bibat. The author inspects their analysis's results, edits files
-corresponding to the boxes on the left, runs the command
-\texttt{make\ analysis}, then repeats. Note that this workflow is
-modular, accommodates plurality and results in a final analysis that can
-be fully reproduced using a single command.}
-  \Description{DESCRIPTION}
-\end{teaserfigure}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
-
-\setlength{\parskip}{-0.1pt}
 
@@
 preparation, model design, diagnosis, debugging and criticism, dates at
-least to \citet{boxBayesianInferenceStatistical1992}. There is
-increasing scholarly recognition of the need for a holistic view of
-``Bayesian workflow''
-\citep{gelmanBayesianWorkflow2020, grinsztajnBayesianWorkflowDisease2021, gabryVisualizationBayesianWorkflow2019}.
-Relatedly, software tools now exist that address most individual aspects
-of a Bayesian workflow: see \citet{strumbeljPresentFutureSoftware} for a
-review of the state of the art.
+least to Box and Tiao (1992). There is increasing scholarly recognition
+of the need for a holistic view of ``Bayesian workflow'' (Gelman et al.
+2020; Grinsztajn et al. 2021; Gabry et al. 2019). Relatedly, software
+tools now exist that address most individual aspects of a Bayesian
+workflow: see Štrumbelj et al. (2024) for a review of the state of the
+art.
 
@@
 
-Bibat can be installed on the Windows, linux or macos command line on by
+Bibat can be installed on the Windows, Linux or macOS command line on by
 running the command \texttt{pip\ install\ bibat} and then used by
@@
 command \texttt{make\ analysis} without the need for any further action
-by the user: in this sense bibat comes with ``batteries included''.
+by the user: in this sense bibat comes with batteries included.
+
+Figure~\ref{fig-workflow} illustrates the components of a bibat-based
+Bayesian workflow and shows how it proceeds: the project team edits the
+code components, then runs \texttt{make\ analysis} triggering creation
+of the result components. After inspecting these they repeat the
+process, leading to a cycle that ultimately results in a complete,
+easily reproducible analysis.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/workflow.png}
+
+}
+
+\caption{\label{fig-workflow}Schematic representation of a Bayesian
+workflow implemented using bibat. The author inspects their analysis's
+results, edits code corresponding to the boxes on the left, runs the
+command \texttt{make\ analysis}, then repeats. The diagram illustrates
+several key features of bibat: inference components are modular and
+plural, the overall workflow is iterative and cyclical and the whole
+analysis can be executed with a single command.}
+
+\end{figure}%
 
@@
 
+In particular, this target attempts to install cmdstan if necessary,
+using a recipe tailored to the host operating system. This functionality
+addresses a common issue where researchers find it difficult to install
+Stan, especially on Windows.
+
 Bibat also provides its Python code in the form of a package configured
@@
 Bibat projects are is written in modern Python and uses pydantic
-\citet{pydanticdevelopersPydantic2022} and pandera
-\citep{niels_bantilan-proc-scipy-2020} for data modelling, Stan
-\citep{carpenterStanProbabilisticProgramming2017} for statistical
-inference, cmdstanpy \citep{standevelopmentteamCmdStanPy2022} for
-python-Stan interface, arviz \citep{kumarArviZUnifiedLibrary2019} for
-storing and analyzing inferences and sphinx
-\citep{georgbrandlandthesphinxteamSphinx2022} and quarto
-\citep{Allaire_Quarto_2022} for documentation.
+Pydantic developers (2022) and pandera (Niels Bantilan 2020) for data
+modelling, Stan (Carpenter et al. 2017) for statistical inference,
+cmdstanpy (Stan Development Team 2022) for python-Stan interface, arviz
+(Kumar et al. 2019) for storing and analyzing inferences and sphinx
+(Georg Brandl and the Sphinx team 2022) and quarto (Allaire et al. 2022)
+for documentation.
 
 Bibat itself is also written in modern Python and uses the popular tools
-cookiecutter \citep{greenfeldCookiecutter2021}, pydantic and click
-\citep{clickdevelopersClickPythonComposable2022}.
-
-Bibat's makefile detects the current operating system and attempts to
-install cmdstan appropriately if necessary. This functionality addresses
-a common issue where researchers find it difficult to install Stan,
-especially on Windows.
+cookiecutter (Greenfeld et al. 2021), pydantic and click (Click
+Developers 2022).
 
@@
 
-Bibat is part of the PyOpenSci ecosystem, allowing for community help
-with maintenance as well as peer review for code and documentation
-quality, usability and accessibility. The PyOpenSci peer review for
-bibat can be found here:
-\url{https://github.com/pyOpenSci/software-submission/issues/83}
-
 \section{Challenges}\label{challenges}
@@
 
-As discussed in \citet{gelmanBayesianWorkflow2020}, Bayesian workflows
-are complicated, featuring plurality, cyclicity and complexity at many
+As discussed in Gelman et al. (2020), Bayesian workflows are
+complicated, featuring plurality, cyclicity and complexity at many
 levels. Bibat accommodates this by separating non-interacting analysis
@@
 
-\section{Post launch use}\label{post-launch-use}
+\section{Comparison with alternative
+software}\label{comparison-with-alternative-software}
 
-Bibat has been used for several real statistical analyses.
+Other than bibat, there is currently no interactive template that
+specifically targets Bayesian workflow projects. There are some
+templates that arguably encompass Bayesian workflow as a special case of
+data analysis project, such as cookiecutter-data-science (Driven Data
+2022), but these are of limited use compared with a specialised template
+due to the many specificities of Bayesian workflow.
 
-\citet{grovesteddyDgfreg2023} used bibat to compare a Bayesian and two
+There is some software that addresses the general task of facilitating
+Bayesian workflow, but using a different approach from bibat's. For
+example, bambi (Capretto et al. 2020) and brms (Bürkner 2017) aim to
+make implementing Bayesian workflows easier by providing ergonomic ways
+to specify and fit Bayesian regression models to tabular datasets. Bibat
+is complementary with these packages, as it targets use cases that they
+do not support, such as analyses where complex datasets or custom models
+might be required.
+
+\section{Case studies}\label{case-studies}
+
+The following cases illustrate how bibat has been used in practice to
+facilitate Bayesian workflow projects.
+
+Groves and Jooste (2023) used bibat to compare a Bayesian and two
 non-Bayesian approaches to modelling a biochemical thermodynamics
 dataset. Bibat facilitated this analysis even though it was not very
-large by providing the fitting mode abstraction, which was very useful
-for comparing the different methods.
+large---the final analysis contained one dataset, three models and three
+inferences---because the fitting mode abstraction allowed for
+straightforward comparison of the different methods. Additionally, bibat
+made it easier to iteratively investigate and discard models that did
+not form part of the final analysis.
 
-In \citet{grovesBaseball2022}, Bibat was used to implement a sports
-analysis involving two datasets, two models and four inferences,
-demonstrating that the generalised Pareto distribution can be used to
-describe hitting ability in baseball. This analysis is now included in
-bibat as an illustration, along with an accompanying tutorial.
+In Groves (2022), Bibat was used to implement a sports analysis
+involving two datasets, two models and four inferences, demonstrating
+that the generalised Pareto distribution can be used to describe hitting
+ability in baseball. This analysis is now included in bibat as an
+illustration, along with an accompanying tutorial. An illustrative
+graphic from this analysis is shown in Figure~\ref{fig-baseball}.
+
+In this case bibat was useful because of its ability to implement
+arbitrary statistical models, as latent generalised Pareto distributions
+are not supported by generic regression packages like brms and bambi.
+Further, bibat's modular design made it easier to implement this
+medium-sized analysis with two datasets, two models and six inferences.
 
@@
 
-\includegraphics[width=0.5\textwidth,height=\textheight]{img/baseball.png}
+\includegraphics[width=1\textwidth,height=\textheight]{img/baseball.png}
 
@@
 
-In \citet{grovesteddySphincter2024}, bibat was used to implement an
-analysis of cerebrovascular data from mice, involving two raw datasets,
-6 prepared datasets, 15 models and 15 inferences. In this case bibat's
-modular design made it straightforward to iteratively add data
-transformations and models while maintaining reproducibility.
+In Groves (2024), bibat was used to implement a large analysis of
+cerebrovascular data from mice, involving two raw datasets, 6 prepared
+datasets, 15 models and 15 inferences. Again it was not possible to
+implement the analysis using formula-based packages due to the need for
+custom modelling approaches to take into account expert knowledge about
+both the target system and the measurement apparatus. Due to the number
+of datasets and models that needed to be considered, implementing this
+analysis without using bibat or a similar structured approach would have
+required a combination of more time and/ or compromises in software
+quality.
 
@@
 
+\section{Community}\label{community}
+
+Bibat is developed in public and encourages community contribution.
+Please see the contributing page
+\url{https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst}
+and code of conduct
+\url{https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md}
+if you would like to help develop bibat.
+
 Bibat has a growing user community, with 16 GitHub stars at the time of
-writing.
+writing, and is affiliated with cmdstanpy through a link on its
+documentation website. Bibat is also affiliated with the Python
+scientific software community PyOpenSci, allowing for help with
+maintenance as well as peer review for code and documentation quality,
+usability and accessibility. The PyOpenSci peer review for bibat can be
+found here:
+\url{https://github.com/pyOpenSci/software-submission/issues/83}.
 
-\section{Comparison with alternative
-software}\label{comparison-with-alternative-software}
+\section{Author contributions}\label{author-contributions}
 
-Other than bibat, there is currently no interactive template that
-specifically targets Bayesian workflow projects. There are some
-templates that arguably encompass Bayesian workflow as a special case of
-data analysis project, such as cookiecutter-data-science
-\citep{drivendataCookiecutterdatascience2022}, but these are of limited
-use compared with a specialised template due to the many specificities
-of Bayesian workflow.
+\section*{Acknowledgements}\label{acknowledgements}
+\addcontentsline{toc}{section}{Acknowledgements}
 
-There is some software that addresses the general task of facilitating
-Bayesian workflow, but using a different approach from bibat's. For
-example, bambi \citep{capretto2020} and brms
-\citep{burknerBrmsPackageBayesian2017} aim to make implementing Bayesian
-workflows easier by providing ergonomic ways to specify and fit Bayesian
-regression models to tabular datasets. Bibat is complementary with these
-packages, as it targets use cases that they do not support, such as
-analyses where complex datasets or custom models might be required.
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
+Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
+Christophe Dervieux. 2022. {``Quarto.''}
+\url{https://doi.org/10.5281/zenodo.5960048}.
 
-\section{Statements and Declarations}\label{statements-and-declarations}
+\bibitem[\citeproctext]{ref-boxBayesianInferenceStatistical1992}
+Box, George E. P., and George C. Tiao. 1992. \emph{Bayesian Inference in
+Statistical Analysis}. Wiley classics library ed. A {Wiley-Interscience}
+Publication. {New York}: {Wiley}.
 
-The authors have no financial or non-financial interests that are
-directly or indirectly related to the work submitted for publication.
+\bibitem[\citeproctext]{ref-burknerBrmsPackageBayesian2017}
+Bürkner, Paul-Christian. 2017. {``Brms: {An R} Package for {Bayesian}
+Multilevel Models Using {Stan}.''} \emph{Journal of Statistical
+Software} 80 (1): 1--28. \url{https://doi.org/10.18637/jss.v080.i01}.
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography.bib}
+\bibitem[\citeproctext]{ref-capretto2020}
+Capretto, Tomás, Camen Piho, Ravin Kumar, Jacob Westfall, Tal Yarkoni,
+and Osvaldo A. Martin. 2020. {``Bambi: {A} Simple Interface for Fitting
+{Bayesian} Linear Models in {Python}.''}
+\url{https://arxiv.org/abs/2012.10754}.
 
-%% begin pandoc before-bib
-%% end pandoc before-bib
-%% begin pandoc biblio
-%% end pandoc biblio
-%% begin pandoc include-after
-%% end pandoc include-after
-%% begin pandoc after-body
-%% end pandoc after-body
+\bibitem[\citeproctext]{ref-carpenterStanProbabilisticProgramming2017}
+Carpenter, Bob, Andrew Gelman, Matthew D. Hoffman, Daniel Lee, Ben
+Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li,
+and Allen Riddell. 2017. {``Stan: {A Probabilistic Programming
+Language}.''} \emph{Journal of Statistical Software} 76 (1): 1--32.
+\url{https://doi.org/10.18637/jss.v076.i01}.
+
+\bibitem[\citeproctext]{ref-clickdevelopersClickPythonComposable2022}
+Click Developers. 2022. {``Click: {Python} Composable Command Line
+Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.
+
+\bibitem[\citeproctext]{ref-drivendataCookiecutterdatascience2022}
+Driven Data. 2022. {``Cookiecutter-Data-Science.''}
+\url{https://github.com/drivendata/cookiecutter-data-science/}.
+
+\bibitem[\citeproctext]{ref-gabryVisualizationBayesianWorkflow2019}
+Gabry, Jonah, Daniel Simpson, Aki Vehtari, Michael Betancourt, and
+Andrew Gelman. 2019. {``Visualization in {Bayesian Workflow}.''}
+\emph{Journal of the Royal Statistical Society Series A: Statistics in
+Society} 182 (2): 389--402. \url{https://doi.org/10.1111/rssa.12378}.
+
+\bibitem[\citeproctext]{ref-gelmanBayesianWorkflow2020}
+Gelman, Andrew, Aki Vehtari, Daniel Simpson, Charles C. Margossian, Bob
+Carpenter, Yuling Yao, Lauren Kennedy, Jonah Gabry, Paul-Christian
+Bürkner, and Martin Modrák. 2020. {``Bayesian {Workflow}.''}
+\emph{arXiv:2011.01808 {[}Stat{]}}, November.
+\url{http://arxiv.org/abs/2011.01808}.
+
+\bibitem[\citeproctext]{ref-georgbrandlandthesphinxteamSphinx2022}
+Georg Brandl and the Sphinx team. 2022. {``Sphinx.''}
+\url{https://www.sphinx-doc.org/}.
+
+\bibitem[\citeproctext]{ref-greenfeldCookiecutter2021}
+Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
+2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.
+
+\bibitem[\citeproctext]{ref-grinsztajnBayesianWorkflowDisease2021}
+Grinsztajn, Léo, Elizaveta Semenova, Charles C. Margossian, and Julien
+Riou. 2021. {``Bayesian Workflow for Disease Transmission Modeling in
+{Stan}.''} \emph{Statistics in Medicine} 40 (27): 6209--34.
+\url{https://doi.org/10.1002/sim.9164}.
+
+\bibitem[\citeproctext]{ref-grovesBaseball2022}
+Groves, Teddy. 2022. {``Baseball.''}
+\url{https://github.com/teddygroves/baseball}.
+
+\bibitem[\citeproctext]{ref-grovesteddySphincter2024}
+---------. 2024. {``Sphincter.''}
+\url{https://github.com/teddygroves/sphincter}.
+
+\bibitem[\citeproctext]{ref-grovesteddyDgfreg2023}
+Groves, Teddy, and Jason Jooste. 2023. {``Dgfreg.''} DTU Biosustain.
+\url{https://github.com/biosustain/dgfreg}.
+
+\bibitem[\citeproctext]{ref-kumarArviZUnifiedLibrary2019}
+Kumar, Ravin, Colin Carroll, Ari Hartikainen, and Osvaldo Martin. 2019.
+{``{ArviZ} a Unified Library for Exploratory Analysis of {Bayesian}
+Models in {Python}.''} \emph{Journal of Open Source Software} 4 (33):
+1143. \url{https://doi.org/10.21105/joss.01143}.
+
+\bibitem[\citeproctext]{ref-niels_bantilan-proc-scipy-2020}
+Niels Bantilan. 2020. {``Pandera: {Statistical Data Validation} of
+{Pandas Dataframes}.''} In \emph{Proceedings of the 19th {Python} in
+{Science Conference}}, edited by Meghann Agarwal, Chris Calloway, Dillon
+Niederhut, and David Shupe, 116--24.
+\url{https://doi.org/10.25080/Majora-342d178e-010}.
+
+\bibitem[\citeproctext]{ref-pydanticdevelopersPydantic2022}
+Pydantic developers. 2022. {``Pydantic.''}
+\url{https://pypi.org/project/pydantic/}.
+
+\bibitem[\citeproctext]{ref-standevelopmentteamCmdStanPy2022}
+Stan Development Team. 2022. {``{CmdStanPy}.''}
+\url{https://github.com/stan-dev/cmdstanpy}.
+
+\bibitem[\citeproctext]{ref-strumbeljPresentFutureSoftware}
+Štrumbelj, Erik, Alexandre Bouchard-Côté, Jukka Corander, Andrew Gelman,
+Håvard Rue, Lawrence Murray, Henri Pesonen, Martyn Plummer, and Aki
+Vehtari. 2024. {``Past, {Present}, and {Future} of {Software} for
+{Bayesian Inference}.''}
+
+\end{CSLReferences}
 
 \end{document}
-\endinput
-%%
-%% End of file `sample-manuscript.tex'.

### 2024-02-02 example_projects/baseball/docs/report.qmd
copier

--- 2024-01-25 bibat/examples/baseball/docs/report.qmd
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=74 end-line=159}
+```{.python include="../baseball/data_preparation.py" start-line=75 end-line=159}
 ```
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=25 end-line=32}
+```{.python include="../baseball/data_preparation.py" start-line=162 end-line=182}
 ```
@@
 
-```{.python include="../baseball/stan_input_functions.py" start-line=11 end-line=34}
+```{.python include="../baseball/stan_input_functions.py" start-line=22 end-line=46}
 ```
@@
 ^[These functions require some more imports: `numpy` and `scipy.special.expit`]
-```{.python include="../baseball/stan_input_functions.py" start-line=36}
+```{.python include="../baseball/stan_input_functions.py" start-line=49}
 ```

### 2024-02-02 paper/main.qmd
copier

@@
 
-The idea that Bayesian statistical analysis comprises not just
-inference, but also specific approaches to related activities like data
-preparation, model design, diagnosis, debugging and criticism, dates
-at least to @boxBayesianInferenceStatistical1992. There is increasing
-scholarly recognition of the need for a holistic view of "Bayesian
-workflow" [@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
-@gabryVisualizationBayesianWorkflow2019]. Relatedly, software tools now
-exist that address most individual aspects of a Bayesian workflow: see
+The term "Bayesian workflow" captures the idea that Bayesian statistical
+analysis comprises not just inference, but also specific approaches to related
+activities like data preparation, model design, diagnosis, debugging and
+criticism. This idea can be found in @boxBayesianInferenceStatistical1992
+and has recently received increasing scholarly recognition of
+[@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
+@gabryVisualizationBayesianWorkflow2019]. Software tools now exist
+that address most individual aspects of a Bayesian workflow: see
 @strumbeljPresentFutureSoftware for a review of the state of the art.
 
-Unfortunately, currently available tools typically address one, or at most
-a handful, of Bayesian workflow activities; it is left to the individual
-project team to orchestrate the components. Writing software that performs
-this orchestration can be time-consuming and tricky, especially in the common
-scenario where it is not initially clear how many, or what kind of, statistical
-models, datasets, data manipulations or investigations an analysis will require.
+Unfortunately, each tool typically addresses one, or at most a few, of the many
+activities that comprise a real Bayesian workflow software project; it is left
+to the individual project team to orchestrate all of the tools they require.
+Writing software that performs this orchestration can be time-consuming and
+tricky, especially in the common scenario where it is not initially clear how
+many, or what kind of, statistical models, datasets, data manipulations or
+investigations an analysis will require.
 
-Bibat addresses this difficulty by providing a full-featured, high-quality
-Bayesian workflow project that can be extended to implement a wide range of
-statistical analyses.
+Bibat is a new tool that addresses the difficulty of orchestrating Bayesian
+workflow software by providing a full-featured, high-quality project that can be
+extended to implement a wide range of statistical analyses.
 
 # How bibat works
-
-## Installation and usage
 
@@
 :::
-
-## Documentation
 

### 2024-02-02 paper/main.tex
copier

@@
 explains the motivation for bibat, describes intended usage, discusses
-bibat's design goals and how key challenges that were overcome in the
-course of development, compares bibat with similar software, highlights
-several examples of bibat's use in science and provides links to
-community resources associated with bibat.
+bibat's design, compares bibat with similar software, highlights several
+examples of bibat's use in science and provides links to community
+resources associated with bibat.
 \end{abstract}
@@
 
-The idea that Bayesian statistical analysis comprises not just
-inference, but also specific approaches to related activities like data
-preparation, model design, diagnosis, debugging and criticism, dates at
-least to Box and Tiao (1992). There is increasing scholarly recognition
-of the need for a holistic view of ``Bayesian workflow'' (Gelman et al.
-2020; Grinsztajn et al. 2021; Gabry et al. 2019). Relatedly, software
-tools now exist that address most individual aspects of a Bayesian
-workflow: see Štrumbelj et al. (2024) for a review of the state of the
-art.
+The term ``Bayesian workflow'' captures the idea that Bayesian
+statistical analysis comprises not just inference, but also specific
+approaches to related activities like data preparation, model design,
+diagnosis, debugging and criticism. This idea can be found in Box and
+Tiao (1992) and has recently received increasing scholarly recognition
+of (Gelman et al. 2020; Grinsztajn et al. 2021; Gabry et al. 2019).
+Software tools now exist that address most individual aspects of a
+Bayesian workflow: see Štrumbelj et al. (2024) for a review of the state
+of the art.
 
-Unfortunately, currently available tools typically address one, or at
-most a handful, of Bayesian workflow activities; it is left to the
-individual project team to orchestrate the components. Writing software
-that performs this orchestration can be time-consuming and tricky,
-especially in the common scenario where it is not initially clear how
-many, or what kind of, statistical models, datasets, data manipulations
-or investigations an analysis will require.
+Unfortunately, each tool typically addresses one, or at most a few, of
+the many activities that comprise a real Bayesian workflow software
+project; it is left to the individual project team to orchestrate all of
+the tools they require. Writing software that performs this
+orchestration can be time-consuming and tricky, especially in the common
+scenario where it is not initially clear how many, or what kind of,
+statistical models, datasets, data manipulations or investigations an
+analysis will require.
 
-Bibat addresses this difficulty by providing a full-featured,
-high-quality Bayesian workflow project that can be extended to implement
-a wide range of statistical analyses.
+Bibat is a new tool that addresses the difficulty of orchestrating
+Bayesian workflow software by providing a full-featured, high-quality
+project that can be extended to implement a wide range of statistical
+analyses.
 
 \section{How bibat works}\label{how-bibat-works}
-
-\subsection{Installation and usage}\label{installation-and-usage}
 
@@
 
-\subsection{Documentation}\label{documentation}
-
 Bibat is documented at \url{https://github.com/teddygroves/bibat/}. The
@@
 
-Bibat's design was informed by these considerations.
+Bibat's design was informed by the aims to accommodate the many sources
+of complexity and in a Bayesian workflow project, to ensure easy
+reproducibility and to integrate many open-source, widely-adopted and
+powerful Bayesian workflow tools and to encourage collaborative
+development.
 
-\subsection{Accommodate a wide range of statistical
-analyses}\label{accommodate-a-wide-range-of-statistical-analyses}
-
-Bibat projects explicitly allow for plurality at the level of input
-datasets, prepared datasets, statistical models, fitting modes,
-computation methods and analyses. In addition, bibat ensures that there
-are minimal restrictions on the kind of components: for example,
-datasets need not be singular or tabular, and statistical models need
-not be representable in formula syntax.
-
+As discussed in Gelman et al. (2020), Bayesian workflows are
+complicated, featuring plurality, cyclicity and complexity at many
+levels. As a specialised Bayesian workflow template, a key goal for
+bibat was to manage this complexity. Bibat achieves this aim by
+separating non-interacting analysis components into separate,
+potentially plural modules and by serialising data to files wherever
+possible. Prepared datasets, statistical models, inference
+configurations, inference results, plots and analyses all have file
+representations. Fitting modes, data manipulations and data models are
+modularised in code through the use of appropriately structured data
+classes and functions. Thanks to this modular approach it is possible to
+perform small sub analyses individually and to iteratively expand the
+analysis by adding components without needing to consider everything at
+once. In addition, bibat ensures that there are minimal restrictions on
+the components: for example, datasets need not be singular or tabular
+and statistical models need not be representable in formula syntax.
 Thanks to these accommodations a project team using bibat should
@@
 
-\subsection{Encourage reproducibility}\label{encourage-reproducibility}
+Bibat encourages reproducibility mainly by providing a preconfigured
+makefile with a target \texttt{analysis} that triggers creation of an
+isolated environment, installation of dependencies, data preparation,
+statistical computation and analysis of results. In this way a bibat
+analysis can be reproduced on most platforms using a single command. In
+particular, this target attempts to install cmdstan if necessary, using
+a recipe tailored to the host operating system. This functionality
+addresses a common issue where researchers find it difficult to install
+Stan, especially on Windows. A second way in which bibat encourages
+reproducibility is by providing its Python code in the form of a
+preconfigured package. Thanks to the use of modern conventions for
+specifying dependencies and configuring tooling, we expect that this
+will make bibat analyses easier for other researchers to work with,
+thereby facilitating replication while also making initial development
+more convenient.
 
-Bibat provides a preconfigured makefile with a target \texttt{analysis}
-triggering creation of an isolated environment, installation of
-dependencies, data preparation, statistical computation and analysis of
-results. In this way a bibat analysis can be reproduced on most
-platforms using a single command.
+Bibat integrates many widely-adopted, open-source and powerful tools to
+implement the components of a Bayesian workflow. Bibat projects are is
+written in modern Python and uses pydantic Pydantic developers (2022)
+and pandera (Niels Bantilan 2020) for data modelling, Stan (Carpenter et
+al. 2017) for statistical inference, cmdstanpy (Stan Development Team
+2022) for python-Stan interface, arviz (Kumar et al. 2019) for storing
+and analyzing inferences and sphinx (Georg Brandl and the Sphinx team
+2022) and quarto (Allaire et al. 2022) for documentation. Bibat itself
+is also written in modern Python and uses the popular tools cookiecutter
+(Greenfeld et al. 2021), pydantic and click (Click Developers 2022).
 
-In particular, this target attempts to install cmdstan if necessary,
-using a recipe tailored to the host operating system. This functionality
-addresses a common issue where researchers find it difficult to install
-Stan, especially on Windows.
+To encourage collaborative development of Bayesian workflow projects,
+Bibat projects include a preconfigured test environment, continuous
+integration, linting and pre-commit hooks. In addition, including
+documentation as a first class component of the analysis addresses a
+common problem in academic statistics projects where the paper gets out
+of sync with the code. Bibat itself is continuously tested to ensure
+that it works on the operating systems Linux, macOS and Windows. Bibat's
+continuous integration runs a test suite as well as an end-to-end
+functional test on all supported Python versions.
 
-Bibat also provides its Python code in the form of a package configured
-using modern conventions for specifying dependencies and configuring
-tooling, so that it is easy to maintain reproducibility as the analysis
-develops.
+\section{Fitting modes}\label{fitting-modes}
 
-\subsection{Use widely-adopted, open source and actively developed
-tools}\label{use-widely-adopted-open-source-and-actively-developed-tools}
+The most novel part of bibat's design is the introduction of an
+abstraction called ``fitting mode'', which allows bibat project to
+handle different ways of fitting a model to a dataset appropriately and
+flexibly. This is often necessary as part of a Bayesian workflow: for
+example, one might perform MCMC sampling of both the prior and posterior
+distributions, perform multiple leave-out-one-fold fits for
+cross-validation or need to compare MCMC sampling with an
+optimisation-based alternative.
 
-Bibat projects are is written in modern Python and uses pydantic
-Pydantic developers (2022) and pandera (Niels Bantilan 2020) for data
-modelling, Stan (Carpenter et al. 2017) for statistical inference,
-cmdstanpy (Stan Development Team 2022) for python-Stan interface, arviz
-(Kumar et al. 2019) for storing and analyzing inferences and sphinx
-(Georg Brandl and the Sphinx team 2022) and quarto (Allaire et al. 2022)
-for documentation.
+Fitting modes in bibat projects take the form of instances of the class
+\texttt{FittingMode}. Each fitting mode contains a name, a function that
+fits a prepared dataset and instructions for how and where to save the
+results. For example, the provided prior sampling fitting mode creates a
+Stan input dictionary with the \texttt{likelihood} data variable set to
+\texttt{0}, performs MCMC sampling and writes data to the
+\texttt{InferenceData} group \texttt{prior}. This design allows multiple
+fits to be stored alongside each other naturally, facilitating many
+common Bayesian workflow analyses, and can easily be configured to
+prevent unnecessary computations.
 
-Bibat itself is also written in modern Python and uses the popular tools
-cookiecutter (Greenfeld et al. 2021), pydantic and click (Click
-Developers 2022).
-
-\subsection{Implement community standards and best practices for
-collaborative software
-development}\label{implement-community-standards-and-best-practices-for-collaborative-software-development}
-
-Bibat projects include a preconfigured test environment, continuous
-integration, linting and pre-commit hooks, making them suitable for
-collaborative software development. In addition, including documentation
-as a first class component of the analysis addresses a common problem in
-academic statistics projects where the paper gets out of sync with the
-code.
-
-Bibat itself is continuously tested to ensure that it works on the
-operating systems Linux, macOS and Windows. Bibat's continuous
-integration runs a test suite as well as an end-to-end functional test
-on all supported Python versions.
-
-\section{Challenges}\label{challenges}
-
-This section describes some specific challenges that often affect
-Bayesian workflow projects and how bibat addresses them.
-
-\subsection{Address complexity using a modular, file based
-approach}\label{address-complexity-using-a-modular-file-based-approach}
-
-As discussed in Gelman et al. (2020), Bayesian workflows are
-complicated, featuring plurality, cyclicity and complexity at many
-levels. Bibat accommodates this by separating non-interacting analysis
-components into modules and by serialising data to files wherever
-possible. Prepared datasets, statistical models, inference
-configurations, inference results, plots and analyses all have file
-representations. Fitting modes, data manipulations and data models are
-modularised in code through the use of appropriately structured data
-classes and functions.
-
-Thanks to this approach it is possible to perform small sub analyses
-individually and to iteratively add components without needing to
-consider everything at once.
-
-\subsection{Fitting modes}\label{fitting-modes}
-
-As part of a Bayesian workflow it can be necessary to fit a model and
-dataset in different ways. For example, one might perform MCMC sampling
-of both the prior and posterior distributions, perform multiple
-leave-out-one-fold fits for cross-validation or need to compare MCMC
-sampling with an optimisation-based alternative.
-
-Bibat accommodates this need by introducing an abstraction called
-``fitting mode''. This abstraction allows bibat projects to handle
-fitting a model and dataset in different ways appropriately and
-flexibly. For example, the provided prior sampling fitting mode creates
-a Stan input dictionary with the \texttt{likelihood} data variable set
-to \texttt{0}, performs MCMC sampling and writes data to the
-\texttt{InferenceData} group \texttt{prior}. Bibat provides fitting
-modes corresponding for prior sampling, posterior sampling and k-fold
-posterior sampling. Users can easily add additional fitting modes by
-copying these examples.
+Bibat provides fitting modes corresponding for prior sampling, posterior
+sampling and k-fold posterior sampling. Users can easily add additional
+fitting modes by copying these examples, or modify the
+\texttt{FittingMode} class to achieve even richer functionality.
 

### 2024-02-02 bibat/__init__.py

@@ -1,5 +1,0 @@
-"""bibat, the Batteries-included Bayesian Analysis Template."""
-
-from bibat import cli, wizarding
-
-__all__ = ["cli", "wizarding"]

### 2024-02-02 bibat/cli.py

@@ -1,85 +1,0 @@
-"""Cli for using bibat.
-
-This module contains functions that are exposed to bibat's command line
-interface using click.
-
-"""
-
-import os
-from importlib.metadata import version
-from typing import Optional
-
-import click
-from cookiecutter.main import cookiecutter
-
-from bibat.wizarding import WizardChoice, WizardStr, prompt_user
-
-THIS_DIR = os.path.dirname(__file__)
-
-WIZARD_FIELDS = [
-    WizardStr("project_name", "What is your project called?", "Project Name"),
-    WizardStr(
-        "repo_name",
-        "What should the project repository be called",
-        default_function=lambda context: context["project_name"]
-        .lower()
-        .replace(" ", "_"),
-    ),
-    WizardStr("author_name", "What is your name?", default="Author name"),
-    WizardStr(
-        "author_email", "What is your email?", default="author@email.com"
-    ),
-    WizardStr(
-        "coc_contact",
-        "Who should be the code of conduct contact?",
-        default_function=lambda context: context["author_email"],
-    ),
-    WizardStr(
-        "description",
-        "Please briefly describe your project",
-        default="A short description of the project.",
-    ),
-    WizardChoice(
-        "open_source_license",
-        "Choose an open source license from these options:",
-        ["MIT", "BSD-3-Clause", "No license file"],
-        default="MIT",
-    ),
-    WizardChoice(
-        "docs_format",
-        "How would you like to document your project?",
-        ["Quarto", "Sphinx", "No docs"],
-        default="Quarto",
-    ),
-    WizardStr(
-        "create_dotgithub_directory",
-        "Would you like to create a .github directory?",
-        default="y",
-    ),
-]
-
-
-@click.command()
-@click.option(
-    "--config-file",
-    type=click.Path(exists=True),
-    default=None,
-    help="Path to a yaml file with prefilled cookiecutter fields.",
-)
-def generate_project(config_file: Optional[str]):
-    """Generate a Bayesian statistical analysis project.
-
-    :param config_file: Optional string specifying a relative path to a
-        configuration file. See cookiecutter documentation for more about
-        configuration files:
-        https://cookiecutter.readthedocs.io/en/2.0.2/advanced/user_config.html
-    """
-    if config_file is not None:
-        cookiecutter(THIS_DIR, no_input=True, config_file=config_file)
-        return
[…]

### 2024-02-02 bibat/examples/baseball/baseball/data_preparation.py

@@ -1,186 +1,0 @@
-"""Provides function `prepare_data` and runs it.
-
-This function should run some other functions with names `prepare_data_x`, which
-each take in a dataframe of measurements and return a PreparedData object.
-
-"""
-import json
-import os
-from io import StringIO
-from pathlib import Path
-from typing import Any
-
-import pandas as pd
-import pandera as pa
-from pandera.typing import DataFrame, Series
-from pandera.typing.common import DataFrameBase
-from pydantic import BaseModel, field_serializer, field_validator
-
-from baseball import util
-
-HERE = Path(__file__).parent
-RAW_DIR = HERE / ".." / "data" / "raw"
-PREPARED_DIR = HERE / ".." / "data" / "prepared"
-RAW_DATA_FILES = {
-    "2006": [os.path.join(RAW_DIR, "2006.csv")],
-    "bdb": [
-        os.path.join(RAW_DIR, "bdb-main.csv"),
-        os.path.join(RAW_DIR, "bdb-post.csv"),
-        os.path.join(RAW_DIR, "bdb-apps.csv"),
-    ],
-}
-
-
-class MeasurementsDF(pa.SchemaModel):
-    """A PreparedData should have a measurements dataframe like this.
-
-    Other columns are also allowed!
-    """
-
-    player_season: Series[str]
-    season: Series[str] = pa.Field(coerce=True)
-    n_attempt: Series[int] = pa.Field(ge=1)
-    n_success: Series[int] = pa.Field(ge=0)
-
-
-class PreparedData(BaseModel, arbitrary_types_allowed=True):
-    """What prepared data looks like in this analysis."""
-
-    name: str
-    coords: util.CoordDict
-    measurements: Any
-
-    @field_validator("measurements")
-    def validate_measurements(cls, v: Any) -> DataFrameBase[MeasurementsDF]:
-        """Validate the measurements field."""
-        if isinstance(v, str):
-            v = pd.read_json(StringIO(v))
-        return MeasurementsDF.validate(v)
-
-    @field_serializer("measurements")
-    def serialize_measurements(
-        self, measurements: DataFrame[MeasurementsDF], _info
-    ):
-        """Serialise the measurements field."""
-        return measurements.to_json()
-
-
-def load_prepared_data(path_to_data: str) -> PreparedData:
-    """Load a dataset."""
-    with open(path_to_data) as f:
-        raw = json.load(f)
-    return PreparedData(**raw)
-
-
[…]

### 2024-02-02 bibat/examples/baseball/baseball/stan_input_functions.py

@@ -1,103 +1,0 @@
-"""Functions for generating input to Stan from prepared data."""
-
-
-from typing import Any, Callable, Dict
-
-import numpy as np
-from scipy.special import expit, logit
-from stanio.json import process_dictionary
-
-from baseball.data_preparation import PreparedData
-
-
-def returns_stan_input(func: Callable[[Any], Dict]) -> Callable[[Any], Dict]:
-    """Decorate a function so it returns a json-serialisable dictionary."""
-
-    def wrapper(*args, **kwargs):
-        return process_dictionary(func(*args, **kwargs))
-
-    return wrapper
-
-
-@returns_stan_input
-def get_stan_input_normal(ppd: PreparedData) -> Dict:
-    """General function for creating a Stan input."""
-    return {
-        "N": len(ppd.measurements),
-        "K": ppd.measurements["n_attempt"].values,
-        "y": ppd.measurements["n_success"].values,
-        "prior_mu": [logit(0.25), 0.2],
-        "prior_tau": [0.2, 0.1],
-        "prior_b_K": [0, 0.03],
-    }
-
-
-@returns_stan_input
-def get_stan_input_gpareto(ppd: PreparedData) -> Dict:
-    """General function for creating a Stan input."""
-    return {
-        "N": len(ppd.measurements),
-        "K": ppd.measurements["n_attempt"].values,
-        "y": ppd.measurements["n_success"].values,
-        "min_alpha": logit(0.07),
-        "max_alpha": logit(0.5),
-        "prior_sigma": [1.5, 0.4],
-        "prior_k": [-0.5, 1],
-    }
-
-
-@returns_stan_input
-def get_stan_input_normal_fake(ppd: PreparedData) -> Dict:
-    """Generate fake Stan input consistent with the normal model."""
-    N = len(ppd.measurements)
-    rng = np.random.default_rng(seed=1234)
-    true_param_values = {
-        "mu": logit(0.25),
-        "tau": 0.18,  # 2sds is 0.19 to 0.32 batting average
-        "b_K": 0.04,  # slight effect of more attempts
-        "alpha_std": rng.random.normal(loc=0, scale=1, size=N),
-    }
-    K = ppd.measurements["n_attempt"].values
-    log_K_std = (np.log(K) - np.log(K).mean()) / np.log(K).std()
-    alpha = (
-        true_param_values["mu"]
-        + true_param_values["b_K"] * log_K_std
[…]

### 2024-02-02 bibat/wizarding.py

@@ -1,93 +1,0 @@
-"""Dataclasses and functions for running bibat's command line wizard."""
-
-from typing import Callable, Dict, List, Optional, Union
-
-import click
-from pydantic import model_validator
-from pydantic.dataclasses import dataclass
-
-
-@dataclass
-class WizardStr:
-    """A wizard question where the answer is a string.
-
-    :param name: name of the field
-    :param prompt: string that the user will be prompted with
-    :param default: default answer
-    :param default_function: function that returns a default answer, given a
-        dictionary.
-    """
-
-    name: str
-    prompt: str
-    default: Optional[str] = None
-    default_function: Optional[Callable] = None
-
-    @model_validator(mode="after")
-    def default_or_default_function_exists(cls, m: "WizardStr"):
-        """Either the default or the default function must not be None."""
-        assert (m.default is not None) or (
-            m.default_function is not None
-        ), "Either the default or the default function must not be None."
-        return m
-
-
-@dataclass
-class WizardChoice:
-    """A wizard question where the answer comes from a list of choices.
-
-    :param name: name of the field
-    :param prompt: string that the user will be prompted with
-    :param default: default answer
-    :param default_function: function that returns a default answer, given a
-        dictionary.
-    """
-
-    name: str
-    prompt: str
-    options: List[str]
-    default: Optional[str] = None
-    default_function: Optional[Callable] = None
-
-    @model_validator(mode="after")
-    def default_is_an_option(cls, m: "WizardChoice"):
-        """Check that the default is one of the options."""
-        if isinstance(m.default, str):
-            msg = f"default {m.default} not in options {m.options}"
-            assert m.default in m.options, msg
-            return m
-
-    @model_validator(mode="after")
-    def default_or_default_function_exists(cls, m: "WizardChoice"):
-        """Either the default or the default function must not be None."""
-        assert (m.default is not None) or (
-            m.default_function is not None
-        ), "Either the default or the default function must not be None."
-        return m
-
-
-def prompt_user(
-    wq: Union[WizardStr, WizardChoice], context: Optional[Dict]
-) -> str:
-    """Prompt the user for an input and parse it with click.
-
-    :param wq: A wizard question: should be one of the classes defined in the
-        module wizarding.py.
-    :param context: an optional dictionary containing the answers to previous
-        prompts.
-    """
-    if not (isinstance(wq, WizardStr) or isinstance(wq, WizardChoice)):
[…]

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/docs/conf.py

@@ -1,62 +1,0 @@
-"""Configuration file for the Sphinx documentation builder."""
-
-# This file only contains a selection of the most common options. For a full
-# list see the documentation:
-# https://www.sphinx-doc.org/en/master/usage/configuration.html
-
-# -- Path setup --------------------------------------------------------------
-
-# If extensions (or modules to document with autodoc) are in another directory,
-# add these directories to sys.path here. If the directory is relative to the
-# documentation root, use os.path.abspath to make it absolute, like shown here.
-#
-# import os
-# import sys
-# sys.path.insert(0, os.path.abspath('.'))
-
-
-# -- Project information -----------------------------------------------------
-
-project = "{{ cookiecutter.project_name }}"
-copyright = "2022, {{ cookiecutter.author_name }}"
-author = "{{ cookiecutter.author_name }}"
-
-
-# -- General configuration ---------------------------------------------------
-
-# Add any Sphinx extension module names here, as strings. They can be
-# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
-# ones.
-extensions = [
-    "sphinx.ext.autodoc",
-    "sphinx.ext.doctest",
-    "sphinx.ext.intersphinx",
-    "sphinx.ext.todo",
-    "sphinx.ext.coverage",
-    "sphinx.ext.mathjax",
-    "sphinx.ext.ifconfig",
-    "sphinx.ext.viewcode",
-    "sphinx.ext.napoleon",
-]
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ["_templates"]
-
-# List of patterns, relative to source directory, that match files and
-# directories to ignore when looking for source files.
-# This pattern also affects html_static_path and html_extra_path.
-exclude_patterns = ["_build", "Thumbs.db", ".DS_Store"]
-html_static_path = ["_static"]
-
-
-# -- Options for HTML output -------------------------------------------------
-
-# The theme to use for HTML and HTML Help pages.  See the documentation for
-# a list of builtin themes.
-#
-html_theme = "alabaster"
-
-# Add any paths that contain custom static files (such as style sheets) here,
-# relative to this directory. They are copied after the builtin static files,
-# so a file named "default.css" will overwrite the builtin "default.css".
-html_static_path = ["_static"]

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/tests/test_unit/test_inference_configuration.py

@@ -1,98 +1,0 @@
-"""Unit tests for the InferenceConfiguration class."""
-
-import os
-
-import pytest
-from {{cookiecutter.repo_name}}.inference_configuration import (
-    InferenceConfiguration,
-)
-
-SAMPLE_KWARGS = {
-    "iter_warmup": 50,
-    "iter_sampling": 50,
-    "chains": 2,
-    "step_size": 0.01,
-    "adapt_delta": 0.7,
-    "show_progress": False,
-}
-MODES_GOOD = ["prior", "posterior", "kfold"]
-MODES_BAD = ["prio", "prserior", "cross-validation"]
-
-
-def test_model_configuration_good_modes():
-    """Check that an inference configuration with good modes initialises."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_interaction",
-        sample_kwargs=SAMPLE_KWARGS,
-        mode_options={"kfold": {"n_folds": 10}},
-        modes=MODES_GOOD,
-        cpp_options=None,
-        stanc_options=None,
-    )
-
-
-@pytest.mark.xfail
-def test_model_configuration_bad_modes():
-    """Check that an inference configuration with bad modes fails."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_interaction",
-        sample_kwargs=SAMPLE_KWARGS,
-        modes=MODES_BAD,
-        mode_options={"kfold": {"n_folds": 10}},
-        cpp_options=None,
-        stanc_options=None,
-    )
-
-
-@pytest.mark.xfail
-def test_model_configuration_no_k():
-    """Check that an inference configuration with no kfold options fails."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_interaction",
-        sample_kwargs=SAMPLE_KWARGS,
-        mode_options={"kfold": None},  # This is the bad mode!
-        modes=MODES_GOOD,  # it would be ok if 'kfold' weren't in here.
[…]

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/tests/test_unit/test_util.py

@@ -1,79 +1,0 @@
-"""Unit tests for functions in src/util.py."""
-
-import numpy as np
-import pandas as pd
-import pytest
-from pandas.testing import assert_frame_equal, assert_series_equal
-from {{cookiecutter.repo_name}}.util import (
-    StanInputDict,
-    make_columns_lower_case,
-    one_encode,
-    stanify_dict,
-)
-
-
-@pytest.mark.parametrize(
-    "s_in,expected",
-    [
-        (
-            pd.Series(["8", 1, "????"], index=["a", "b", "c"]),
-            pd.Series([1, 2, 3], index=["a", "b", "c"]),
-        ),
-        (
-            pd.Series([1, "????", "????"], index=["a", "b", "c"]),
-            pd.Series([1, 2, 2], index=["a", "b", "c"]),
-        ),
-    ],
-)
-def test_one_encode(s_in: pd.Series, expected: pd.Series):
-    """Check that the function one_encode works as expected."""
-    assert_series_equal(one_encode(s_in), expected)
-
-
-@pytest.mark.parametrize(
-    "df_in,expected",
-    [
-        (
-            pd.DataFrame({"A": [1, 2, 3], "B": ["a", "b", "c"]}),
-            pd.DataFrame({"a": [1, 2, 3], "b": ["a", "b", "c"]}),
-        ),
-        (
-            pd.DataFrame(
-                [[1, 1]],
-                columns=pd.MultiIndex.from_product([["A"], ["B", "C"]]),
-            ),
-            pd.DataFrame(
-                [[1, 1]],
-                columns=pd.MultiIndex.from_product([["a"], ["b", "c"]]),
-            ),
-        ),
-    ],
-)
-def test_make_columns_lower_case(df_in: pd.DataFrame, expected: pd.DataFrame):
-    """Check that the function make_columns_lower_case works as expected."""
-    assert_frame_equal(make_columns_lower_case(df_in), expected)
-
-
-@pytest.mark.parametrize(
-    "d_in,expected",
-    [
-        ({"a": pd.Series([1, 2, 3])}, {"a": [1, 2, 3]}),
-        (
-            {"a": pd.DataFrame([[1, 2, 3], [4, 5, 6]])},
-            {"a": [[1, 2, 3], [4, 5, 6]]},
-        ),
-        (
-            {
-                "a": np.array([[1, 2, 3], [4, 5, 6]]),
-            },
[…]

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/__init__.py

@@ -1,1 +1,0 @@
-"""Source code for {{cookiecutter.project_name}}."""

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/data_preparation.py

@@ -1,167 +1,0 @@
-"""Provides function `prepare_data` and runs it.
-
-This function should run some other functions with names `prepare_data_x`, which
-each take in a dataframe of measurements and return a PreparedData object.
-
-"""
-import json
-import os
-from io import StringIO
-from pathlib import Path
-from typing import Any
-
-import numpy as np
-import pandas as pd
-import pandera as pa
-from pandera.typing import DataFrame, Series
-from pandera.typing.common import DataFrameBase
-from pydantic import BaseModel, field_serializer, field_validator
-from {{cookiecutter.repo_name}} import util
-
-HERE = Path(__file__).parent
-RAW_DIR = HERE /  ".." / "data" / "raw"
-PREPARED_DIR = HERE / ".." / "data" / "prepared"
-RAW_DATA_FILES = {"measurements": RAW_DIR / "raw_measurements.csv"}
-
-
-class MeasurementsDF(pa.SchemaModel):
-    """A PreparedData should have a measurements dataframe like this.
-
-    Other columns are also allowed!
-    """
-
-    x1: Series[float]
-    x2: Series[float]
-    x1colonx2: Series[float] = pa.Field(alias="x1:x2")
-    y: Series[float]
-
-
-class PreparedData(BaseModel, arbitrary_types_allowed=True):
-    """What prepared data looks like in this analysis."""
-
-    name: str
-    coords: util.CoordDict
-    measurements: Any
-
-    @field_validator("measurements")
-    def validate_measurements(
-        cls, v: Any
-    ) -> DataFrameBase[MeasurementsDF]:
-        if isinstance(v, str):
-            v = pd.read_json(StringIO(v))
-        return MeasurementsDF.validate(v)
-
-    @field_serializer("measurements")
-    def serialize_measurements(
-        self, measurements: DataFrame[MeasurementsDF], _info
-    ):
-        return measurements.to_json()
-
-
-def load_prepared_data(path_to_data: str) -> PreparedData:
-    """Load a dataset."""
-    with open(path_to_data) as f:
-        raw = json.load(f)
-    return PreparedData(**raw)
-
-
-def prepare_data_interaction(measurements_raw: pd.DataFrame) -> PreparedData:
-    """Prepare data with an interaction column."""
-    measurements = process_measurements(measurements_raw)
-    return PreparedData(
-        name="interaction",
-        coords=util.CoordDict(
-            {
-                "covariate": ["x1", "x2", "x1:x2"],
[…]

### 2024-02-02 bibat/{{cookiecutter.repo_name}}/{{cookiecutter.repo_name}}/sample.py

@@ -1,66 +1,0 @@
-"""Run all the inferences in the inferences folder."""
-
-import os
-
-import arviz as az
-import cmdstanpy
-from {{cookiecutter.repo_name}}.data_preparation import load_prepared_data
-from {{cookiecutter.repo_name}}.inference_configuration import (
-    load_inference_configuration,
-)
-
-HERE = os.path.dirname(__file__)
-RUNS_DIR = os.path.join(HERE, "..", "inferences")
-STAN_DIR = os.path.join(HERE, "stan")
-
-def main():
-    """Fit all inferences in all modes."""
-    run_dirs = [
-        os.path.join(RUNS_DIR, d)
-        for d in os.listdir(RUNS_DIR)
-        if os.path.isdir(os.path.join(RUNS_DIR, d))
-    ]
-    for run_dir in sorted(run_dirs):
-        ic = load_inference_configuration(os.path.join(run_dir, "config.toml"))
-        stan_file = os.path.join(STAN_DIR, ic.stan_file)
-        prepared_data_file = os.path.join(
-            "data", "prepared", ic.prepared_data_dir + ".json"
-        )
-        model = cmdstanpy.CmdStanModel(
-            stan_file=stan_file,
-            cpp_options=ic.cpp_options,
-            stanc_options=ic.stanc_options,
-        )
-        prepared_data = load_prepared_data(prepared_data_file)
-        stan_input_base = ic.stan_input_function(prepared_data)
-        idata_kwargs = {
-            "observed_data": stan_input_base,
-            "log_likelihood": "llik",
-            "coords": prepared_data.coords,
-            "dims": ic.dims,
-        }
-        llik_outputs = {}
-        for mode in ic.fitting_modes:
-            fit_kwargs = ic.sample_kwargs
-            if ic.mode_options is not None and mode.name in ic.mode_options.keys():
-                fit_kwargs |= ic.mode_options[mode.name]
-            output = mode.fit(model, stan_input_base, fit_kwargs)
-            if mode.idata_target in ["prior", "posterior"]:
-                idata_kwargs[mode.idata_target] = output
-                idata_kwargs[f"{mode.idata_target.value}_predictive"] = "yrep"
-            elif mode.idata_target == "log_likelihood":
-                llik_outputs[f"llik_{mode.name}"] = output
-            else:
-                raise ValueError(
-                    f"idata_target {mode.idata_target} is not yet supported"
-                )
[…]

### 2024-02-02 example_projects/baseball/baseball/data_preparation.py

@@ -1,0 +1,187 @@
+"""Provides function `prepare_data` and runs it.
+
+This function should run some other functions with names `prepare_data_x`, which
+each take in a dataframe of measurements and return a PreparedData object.
+
+"""
+
+import json
+import os
+from io import StringIO
+from pathlib import Path
+from typing import Any
+
+import pandas as pd
+import pandera as pa
+from pandera.typing import DataFrame, Series
+from pandera.typing.common import DataFrameBase
+from pydantic import BaseModel, field_serializer, field_validator
+
+from baseball import util
+
+HERE = Path(__file__).parent
+RAW_DIR = HERE / ".." / "data" / "raw"
+PREPARED_DIR = HERE / ".." / "data" / "prepared"
+RAW_DATA_FILES = {
+    "2006": [os.path.join(RAW_DIR, "2006.csv")],
+    "bdb": [
+        os.path.join(RAW_DIR, "bdb-main.csv"),
+        os.path.join(RAW_DIR, "bdb-post.csv"),
+        os.path.join(RAW_DIR, "bdb-apps.csv"),
+    ],
+}
+
+
+class MeasurementsDF(pa.SchemaModel):
+    """A PreparedData should have a measurements dataframe like this.
+
+    Other columns are also allowed!
+    """
+
+    player_season: Series[str]
+    season: Series[str] = pa.Field(coerce=True)
+    n_attempt: Series[int] = pa.Field(ge=1)
+    n_success: Series[int] = pa.Field(ge=0)
+
+
+class PreparedData(BaseModel, arbitrary_types_allowed=True):
+    """What prepared data looks like in this analysis."""
+
+    name: str
+    coords: util.CoordDict
+    measurements: Any
+
+    @field_validator("measurements")
+    def validate_measurements(cls, v: Any) -> DataFrameBase[MeasurementsDF]:
+        """Validate the measurements field."""
+        if isinstance(v, str):
+            v = pd.read_json(StringIO(v))
+        return MeasurementsDF.validate(v)
+
+    @field_serializer("measurements")
+    def serialize_measurements(
+        self, measurements: DataFrame[MeasurementsDF], _info
+    ):
+        """Serialise the measurements field."""
+        return measurements.to_json()
+
+
+def load_prepared_data(path_to_data: str) -> PreparedData:
+    """Load a dataset."""
+    with open(path_to_data) as f:
+        raw = json.load(f)
+    return PreparedData(**raw)
[…]

### 2024-02-02 example_projects/baseball/baseball/stan_input_functions.py

@@ -1,0 +1,102 @@
+"""Functions for generating input to Stan from prepared data."""
+
+from typing import Any, Callable, Dict
+
+import numpy as np
+from scipy.special import expit, logit
+from stanio.json import process_dictionary
+
+from baseball.data_preparation import PreparedData
+
+
+def returns_stan_input(func: Callable[[Any], Dict]) -> Callable[[Any], Dict]:
+    """Decorate a function so it returns a json-serialisable dictionary."""
+
+    def wrapper(*args, **kwargs):
+        return process_dictionary(func(*args, **kwargs))
+
+    return wrapper
+
+
+@returns_stan_input
+def get_stan_input_normal(ppd: PreparedData) -> Dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"].values,
+        "y": ppd.measurements["n_success"].values,
+        "prior_mu": [logit(0.25), 0.2],
+        "prior_tau": [0.2, 0.1],
+        "prior_b_K": [0, 0.03],
+    }
+
+
+@returns_stan_input
+def get_stan_input_gpareto(ppd: PreparedData) -> Dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"].values,
+        "y": ppd.measurements["n_success"].values,
+        "min_alpha": logit(0.07),
+        "max_alpha": logit(0.5),
+        "prior_sigma": [1.5, 0.4],
+        "prior_k": [-0.5, 1],
+    }
+
+
+@returns_stan_input
+def get_stan_input_normal_fake(ppd: PreparedData) -> Dict:
+    """Generate fake Stan input consistent with the normal model."""
+    N = len(ppd.measurements)
+    rng = np.random.default_rng(seed=1234)
+    true_param_values = {
+        "mu": logit(0.25),
+        "tau": 0.18,  # 2sds is 0.19 to 0.32 batting average
+        "b_K": 0.04,  # slight effect of more attempts
+        "alpha_std": rng.random.normal(loc=0, scale=1, size=N),
+    }
+    K = ppd.measurements["n_attempt"].values
+    log_K_std = (np.log(K) - np.log(K).mean()) / np.log(K).std()
+    alpha = (
+        true_param_values["mu"]
+        + true_param_values["b_K"] * log_K_std
[…]

### 2024-02-02 template/src/data_preparation.py

@@ -1,0 +1,170 @@
+"""Provides function `prepare_data` and runs it.
+
+This function should run some other functions with names `prepare_data_x`, which
+each take in a dataframe of measurements and return a PreparedData object.
+
+"""
+
+import json
+import os
+from io import StringIO
+from pathlib import Path
+from typing import Any
+
+import numpy as np
+import pandas as pd
+import pandera as pa
+from pandera.typing import DataFrame, Series
+from pandera.typing.common import DataFrameBase
+from pydantic import BaseModel, field_serializer, field_validator
+from src import util
+
+HERE = Path(__file__).parent
+RAW_DIR = HERE / ".." / "data" / "raw"
+PREPARED_DIR = HERE / ".." / "data" / "prepared"
+RAW_DATA_FILES = {"measurements": RAW_DIR / "raw_measurements.csv"}
+
+
+class MeasurementsDF(pa.SchemaModel):
+    """A PreparedData should have a measurements dataframe like this.
+
+    Other columns are also allowed!
+    """
+
+    x1: Series[float]
+    x2: Series[float]
+    x1colonx2: Series[float] = pa.Field(alias="x1:x2")
+    y: Series[float]
+
+
+class PreparedData(BaseModel, arbitrary_types_allowed=True):
+    """What prepared data looks like in this analysis."""
+
+    name: str
+    coords: util.CoordDict
+    measurements: Any
+
+    @field_validator("measurements")
+    def validate_measurements(cls, v: Any) -> DataFrameBase[MeasurementsDF]:
+        """Validate the measurements table."""
+        if isinstance(v, str):
+            v = pd.read_json(StringIO(v))
+        return MeasurementsDF.validate(v)
+
+    @field_serializer("measurements")
+    def serialize_measurements(
+        self, measurements: DataFrame[MeasurementsDF], _info
+    ):
+        """Convert the measurements table to json."""
+        return measurements.to_json()
+
+
+def load_prepared_data(path_to_data: str) -> PreparedData:
+    """Load a dataset."""
+    with open(path_to_data) as f:
+        raw = json.load(f)
+    return PreparedData(**raw)
+
+
+def prepare_data_interaction(measurements_raw: pd.DataFrame) -> PreparedData:
+    """Prepare data with an interaction column."""
+    measurements = process_measurements(measurements_raw)
+    return PreparedData(
+        name="interaction",
+        coords=util.CoordDict(
+            {
[…]

### 2024-02-02 template/src/inference_configuration.py

@@ -1,0 +1,125 @@
+"""Definition of the InferenceConfiguration class."""
+
+import os
+from typing import Callable, Dict, List, Optional
+
+import toml
+from pydantic import BaseModel, Field, field_validator, model_validator
+from src import fitting_mode, stan_input_functions
+
+HERE = os.path.dirname(os.path.abspath(__file__))
+STAN_DIR = os.path.join(HERE, "stan")
+DEFAULT_DIMS = {"llik": ["observation"], "yrep": ["observation"]}
+DEFAULT_SAMPLE_KWARGS = {"show_progress": False}
+
+
+class InferenceConfiguration(BaseModel):
+    """Container for a path to a Stan model and some configuration.
+
+    For example, you may want to compare how well two stan programs fit the
+    same data, or how well the same model fits the data with different
+    covariates.
+
+    :param name: string name identifying the model configuration
+
+    :param stan_file: Path to a Stan program, with "/" even on windows
+
+    :param prepared_data_dir: Path to a directory containing prepared data
+
+    :param stan_input_function: function from src.stan_input_functions used to
+    get a Stan input dictionary from a PreparedData object.
+
+    :param sample_kwargs: dictionary of keyword arguments to
+    cmdstanpy.CmdStanModel.sample.
+
+    :param modes: which modes to run the model in. Choose one or more of the
+    AVAILABLE_MODES.
+
+    :param dims: map from parameter names to lists of coordinate names. See the
+    field names of the file "coords.json" in the prepared_data_dir for possible
+    coordinate names.
+
+    :param kfold_folds: How many kfold folds to run
+
+    :param cpp_options: valid choices for the `cpp_options` argument to
+    CmdStanModel
+
+    :param stanc_options: valid choices for the `cpp_options` argument to
+    CmdStanModel
+
+    """
+
+    name: str
+    stan_file: str
+    prepared_data_dir: str
+    stan_input_function: Callable
+    fitting_mode_names: List[str] = Field(alias="modes")
+    fitting_modes: List[fitting_mode.FittingMode]
+    sample_kwargs: dict = Field(default_factory=lambda: DEFAULT_SAMPLE_KWARGS)
+    dims: Dict[str, List[str]] = Field(default_factory=lambda: DEFAULT_DIMS)
+    mode_options: Optional[Dict[str, dict]] = None
+    kfold_folds: Optional[int] = None
+    cpp_options: Optional[dict] = None
[…]

### 2024-02-02 template/src/sample.py

@@ -1,0 +1,68 @@
+"""Run all the inferences in the inferences folder."""
+
+import os
+
+import arviz as az
+import cmdstanpy
+from src.data_preparation import load_prepared_data
+from src.inference_configuration import load_inference_configuration
+
+HERE = os.path.dirname(__file__)
+RUNS_DIR = os.path.join(HERE, "..", "inferences")
+STAN_DIR = os.path.join(HERE, "stan")
+
+
+def main():
+    """Fit all inferences in all modes."""
+    run_dirs = [
+        os.path.join(RUNS_DIR, d)
+        for d in os.listdir(RUNS_DIR)
+        if os.path.isdir(os.path.join(RUNS_DIR, d))
+    ]
+    for run_dir in sorted(run_dirs):
+        ic = load_inference_configuration(os.path.join(run_dir, "config.toml"))
+        stan_file = os.path.join(STAN_DIR, ic.stan_file)
+        prepared_data_file = os.path.join(
+            "data", "prepared", ic.prepared_data_dir + ".json"
+        )
+        model = cmdstanpy.CmdStanModel(
+            stan_file=stan_file,
+            cpp_options=ic.cpp_options,
+            stanc_options=ic.stanc_options,
+        )
+        prepared_data = load_prepared_data(prepared_data_file)
+        stan_input_base = ic.stan_input_function(prepared_data)
+        idata_kwargs = {
+            "observed_data": stan_input_base,
+            "log_likelihood": "llik",
+            "coords": prepared_data.coords,
+            "dims": ic.dims,
+        }
+        llik_outputs = {}
+        for mode in ic.fitting_modes:
+            fit_kwargs = ic.sample_kwargs
+            if (
+                ic.mode_options is not None
+                and mode.name in ic.mode_options.keys()
+            ):
+                fit_kwargs |= ic.mode_options[mode.name]
+            output = mode.fit(model, stan_input_base, fit_kwargs)
+            if mode.idata_target in ["prior", "posterior"]:
+                idata_kwargs[mode.idata_target] = output
+                idata_kwargs[f"{mode.idata_target.value}_predictive"] = "yrep"
+            elif mode.idata_target == "log_likelihood":
+                llik_outputs[f"llik_{mode.name}"] = output
+            else:
+                raise ValueError(
+                    f"idata_target {mode.idata_target} is not yet supported"
+                )
+        idata = az.from_cmdstanpy(**idata_kwargs)
[…]

### 2024-02-02 template/src/stan_input_functions.py

@@ -1,0 +1,43 @@
+"""Functions for generating input to Stan from prepared data."""
+
+from typing import Any, Callable, Dict, List
+
+import pandas as pd
+from src.data_preparation import PreparedData
+from stanio.json import process_dictionary
+
+
+def returns_stan_input(func: Callable[[Any], Dict]) -> Callable[[Any], Dict]:
+    """Decorate a function so it returns a json-serialisable dictionary."""
+
+    def wrapper(*args, **kwargs):
+        return process_dictionary(func(*args, **kwargs))
+
+    return wrapper
+
+
+@returns_stan_input
+def get_stan_input(measurements: pd.DataFrame, x_cols: List[str]) -> Dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(measurements),
+        "N_train": len(measurements),
+        "N_test": len(measurements),
+        "K": len(x_cols),
+        "x": measurements[x_cols].values.tolist(),
+        "y": measurements["y"].tolist(),
+        "ix_train": [i + 1 for i in range(len(measurements))],
+        "ix_test": [i + 1 for i in range(len(measurements))],
+    }
+
+
+@returns_stan_input
+def get_stan_input_interaction(prepared_data: PreparedData) -> Dict:
+    """Get a Stan input with an interaction predictor."""
+    return get_stan_input(prepared_data.measurements, ["x1", "x2", "x1:x2"])
+
+
+@returns_stan_input
+def get_stan_input_no_interaction(prepared_data: PreparedData) -> Dict:
+    """Get a Stan input without an interaction predictor."""
+    return get_stan_input(prepared_data.measurements, ["x1", "x2"])

### 2024-02-02 template/tests/test_integration/test_data_preparation.py

@@ -1,0 +1,74 @@
+"""Integration tests for functions in src/data_preparation.py."""
+
+from typing import Callable
+
+import pandas as pd
+import pytest
+from pandas.testing import assert_frame_equal
+from src.data_preparation import (
+    prepare_data_interaction,
+    prepare_data_no_interaction,
+)
+from src.util import CoordDict
+
+EXAMPLE_RAW_MEASUREMENTS = pd.DataFrame(
+    {
+        "X1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+        "X2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+        "yButIThoughtIdAddSomeLetters": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+    }
+)
+
+
+@pytest.mark.parametrize(
+    "prepare_data_function,name,raw_measurements,expected_measurements,"
+    "expected_coords",
+    [
+        (
+            prepare_data_interaction,
+            "interaction",
+            EXAMPLE_RAW_MEASUREMENTS,
+            pd.DataFrame(
+                {
+                    "x1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "y": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x1:x2": [1, 4, 9, 1, 4, 9, 1, 4, 9, 1, 4],
+                }
+            ).apply(lambda s: s.astype(float)),
+            {
+                "covariate": ["x1", "x2", "x1:x2"],
+                "observation": [str(i) for i in range(11)],
+            },
+        ),
+        (
+            prepare_data_no_interaction,
+            "no_interaction",
+            EXAMPLE_RAW_MEASUREMENTS,
+            pd.DataFrame(
+                {
+                    "x1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "y": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x1:x2": [1, 4, 9, 1, 4, 9, 1, 4, 9, 1, 4],
+                }
+            ).apply(lambda s: s.astype(float)),
+            {
+                "covariate": ["x1", "x2"],
[…]

### 2024-02-02 template/tests/test_unit/test_inference_configuration.py

@@ -1,0 +1,96 @@
+"""Unit tests for the InferenceConfiguration class."""
+
+import os
+
+import pytest
+from src.inference_configuration import InferenceConfiguration
+
+SAMPLE_KWARGS = {
+    "iter_warmup": 50,
+    "iter_sampling": 50,
+    "chains": 2,
+    "step_size": 0.01,
+    "adapt_delta": 0.7,
+    "show_progress": False,
+}
+MODES_GOOD = ["prior", "posterior", "kfold"]
+MODES_BAD = ["prio", "prserior", "cross-validation"]
+
+
+def test_model_configuration_good_modes():
+    """Check that an inference configuration with good modes initialises."""
+    _ = InferenceConfiguration(
+        name="my_mc",
+        stan_file="multilevel-linear-regression.stan",
+        prepared_data_dir=os.path.join("hi", "hello", "hey"),
+        stan_input_function="get_stan_input_interaction",
+        sample_kwargs=SAMPLE_KWARGS,
+        mode_options={"kfold": {"n_folds": 10}},
+        modes=MODES_GOOD,
+        cpp_options=None,
+        stanc_options=None,
+    )
+
+
+@pytest.mark.xfail
+def test_model_configuration_bad_modes():
+    """Check that an inference configuration with bad modes fails."""
+    _ = InferenceConfiguration(
+        name="my_mc",
+        stan_file="multilevel-linear-regression.stan",
+        prepared_data_dir=os.path.join("hi", "hello", "hey"),
+        stan_input_function="get_stan_input_interaction",
+        sample_kwargs=SAMPLE_KWARGS,
+        modes=MODES_BAD,
+        mode_options={"kfold": {"n_folds": 10}},
+        cpp_options=None,
+        stanc_options=None,
+    )
+
+
+@pytest.mark.xfail
+def test_model_configuration_no_k():
+    """Check that an inference configuration with no kfold options fails."""
+    _ = InferenceConfiguration(
+        name="my_mc",
+        stan_file="multilevel-linear-regression.stan",
+        prepared_data_dir=os.path.join("hi", "hello", "hey"),
+        stan_input_function="get_stan_input_interaction",
+        sample_kwargs=SAMPLE_KWARGS,
+        mode_options={"kfold": None},  # This is the bad mode!
+        modes=MODES_GOOD,  # it would be ok if 'kfold' weren't in here.
+        cpp_options=None,
+        stanc_options=None,
[…]

### 2024-02-02 template/tests/test_unit/test_util.py

@@ -1,0 +1,79 @@
+"""Unit tests for functions in src/util.py."""
+
+import numpy as np
+import pandas as pd
+import pytest
+from pandas.testing import assert_frame_equal, assert_series_equal
+from src.util import (
+    StanInputDict,
+    make_columns_lower_case,
+    one_encode,
+    stanify_dict,
+)
+
+
+@pytest.mark.parametrize(
+    "s_in,expected",
+    [
+        (
+            pd.Series(["8", 1, "????"], index=["a", "b", "c"]),
+            pd.Series([1, 2, 3], index=["a", "b", "c"]),
+        ),
+        (
+            pd.Series([1, "????", "????"], index=["a", "b", "c"]),
+            pd.Series([1, 2, 2], index=["a", "b", "c"]),
+        ),
+    ],
+)
+def test_one_encode(s_in: pd.Series, expected: pd.Series):
+    """Check that the function one_encode works as expected."""
+    assert_series_equal(one_encode(s_in), expected)
+
+
+@pytest.mark.parametrize(
+    "df_in,expected",
+    [
+        (
+            pd.DataFrame({"A": [1, 2, 3], "B": ["a", "b", "c"]}),
+            pd.DataFrame({"a": [1, 2, 3], "b": ["a", "b", "c"]}),
+        ),
+        (
+            pd.DataFrame(
+                [[1, 1]],
+                columns=pd.MultiIndex.from_product([["A"], ["B", "C"]]),
+            ),
+            pd.DataFrame(
+                [[1, 1]],
+                columns=pd.MultiIndex.from_product([["a"], ["b", "c"]]),
+            ),
+        ),
+    ],
+)
+def test_make_columns_lower_case(df_in: pd.DataFrame, expected: pd.DataFrame):
+    """Check that the function make_columns_lower_case works as expected."""
+    assert_frame_equal(make_columns_lower_case(df_in), expected)
+
+
+@pytest.mark.parametrize(
+    "d_in,expected",
+    [
+        ({"a": pd.Series([1, 2, 3])}, {"a": [1, 2, 3]}),
+        (
+            {"a": pd.DataFrame([[1, 2, 3], [4, 5, 6]])},
+            {"a": [[1, 2, 3], [4, 5, 6]]},
+        ),
+        (
+            {
+                "a": np.array([[1, 2, 3], [4, 5, 6]]),
+            },
+            {"a": [[1, 2, 3], [4, 5, 6]]},
[…]

### 2024-02-02 tests/test_unit/test_copier.py

@@ -1,0 +1,26 @@
+"""Unit tests for the copier template."""
+
+from pathlib import Path
+from tempfile import TemporaryDirectory
+
+import copier
+
+HERE = Path(__file__).parent
+EXAMPLE_DATA = {
+    "project_name": "My Cool Project",
+    "project_name_no_spaces": "my_cool_project",
+    "description": "bla bla bla",
+    "author_name": "Author",
+    "author_email": "author@email.com",
+    "coc_contact": "author@email.com",
+    "open_source_license": "MIT",
+    "docs_format": "Quarto",
+    "create_dotgithub_directory": True,
+}
+
+
+def test_copier_runs():
+    """Test that copier runs."""
+    source = HERE / ".." / ".."
+    with TemporaryDirectory() as destination:
+        copier.run_copy(str(source), destination, data=EXAMPLE_DATA)

### 2024-02-02 tests/test_unit/test_wizarding.py

@@ -1,54 +1,0 @@
-"""Test the wizarding module."""
-
-import click
-import pytest
-from click.testing import CliRunner
-
-from bibat.wizarding import WizardChoice, WizardStr, prompt_user
-
-
-def test_wizard_str_good():
-    """Test that the WizardStr class instantiates as expected."""
-    WizardStr("name", "What's your name?", "Sasha")
-
-
-def test_wizard_choice_good():
-    """Test that the WizardChoice class instantiates as expected."""
-    WizardChoice("name", "What's your name?", ["Sasha", "Alex"], "Sasha")
-
-
-@pytest.mark.xfail
-def test_wizard_choice_bad():
-    """Test that the WizardChoice class fails as expected."""
-    WizardChoice("name", "What's your name?", ["Sasha", "Alex"], "Lex")
-
-
-def test_prompt_user_good():
-    """Test that the prompt_user function works as expected."""
-    runner = CliRunner()
-    c = WizardChoice("name", "What's your name?", ["Sasha", "Alex"], "Sasha")
-
-    @click.command()
-    def prompt_c():
-        return prompt_user(c)
-
-    runner.invoke(prompt_c, input="Alex")
-
-
-@pytest.mark.xfail
-def test_prompt_user_bad():
-    """Test that the prompt_user function fails as expected."""
-    runner = CliRunner()
-    c = "asdfasdf"
-
-    @click.command()
-    def prompt_c():
-        return prompt_user(c)
-
-    runner.invoke(prompt_c, input="Alex")
-
-
-@pytest.mark.xfail
-def test_prompt_user_bad_input():
-    """Test that the prompt_user function fails with incorrect default."""
-    prompt_user(123, context=None)

### 2024-02-15 example_projects/baseball/docs/report.qmd
Update baseball example project

--- 2024-01-25 bibat/examples/baseball/docs/report.qmd
@@
 ---
-title: "Modelling at-bats in baseball using the generalised Pareto distribution"
-subtitle: "A report created using bibat version 0.1.6"
+title: "Baseball"
+resource_path: ["../"]
+subtitle: "A report created using bibat version 0.3.0"
 author: "Teddy Groves"
@@
     embed-resources: true
+  pdf: default
 filters:
@@
 
-Even though I don't work with sports data very much any more, the problem still
-haunted me, so when I read [this great case
+Even though I don't work with sports data very much any more, these issues still
+lingered in my mind, so when I read [this great case
 study](https://mc-stan.org/users/documentation/case-studies/gpareto_functions.html#conclusion-on-the-data-analysis)
@@
 
-The idea was this: just as data about intense storms that cause electricity
-problems tell us about a tail of the bigger solar magnetism distribution, maybe
-data about professional sportspeople is best thought about as coming from a tail
-of the general sports ability distribution. If so, maybe something like the
-[generalised pareto
+The idea was this: just as data about intense storms tell us about a tail
+of the bigger solar magnetism distribution, maybe data about professional
+sportspeople is best thought about as coming from a tail of the general sports
+ability distribution. If so, perhaps something like the [generalised pareto
 distribution](https://en.wikipedia.org/wiki/Generalized_Pareto_distribution)
-might be better than the bog standard normal distribution for describing the
+might be better than the bog-standard normal distribution for describing the
 pros' abilities.
@@
 
-With a few datasets and at least two different statistical models to consider,
-the full analysis looked like it would be a little too big to fit in a trivial
-file structure, so it seemed like a good opportunity to showcase my
-batteries-included Bayesian analysis template package
-[bibat](https://github.com/teddygroves/bibat).
+With a few datasets and statistical models to consider, the full analysis looked
+like it would be too big to fit in a single file, so this seemed like a job for a
+batteries-included Bayesian analysis template like [bibat](https://github.com/teddygroves/bibat).
 
@@
 
-Check out the analysis's [github
-repository](https://github.com/teddygroves/baseball) for all the details.
+Check out [the full analysis](https://github.com/teddygroves/bibat/example_projects/baseball) for all the details.
 
@@
 
-First I installed bibat in my global Python 3.11 environment with this command:
+First I installed [copier](https://copier.readthedocs.io/en/stable/) in my
+global Python 3.12 environment with this command:
 
 ```zsh
-> pip install bibat
+> pipx install copier
 ```
@@
 ```{.zsh}
-> bibat
-Welcome to the Batteries-Included Bayesian Analysis Template!
-What is your project called? [Project Name]: Baseball
-What should the project repository be called [baseball]: baseball
-What is your name? [Author name]: Teddy Groves
-What is your email? [author@email.com]:
-Who should be the code of conduct contact? [author@email.com]:
-Please briefly describe your project [A short description of the project.]: Comparison of distributions for modelling baseball hitting
-Choose an open source license from these options: (MIT, BSD-3-Clause, No license file) [MIT]:
-How would you like to document your project? (Quarto, Sphinx, No docs) [Quarto]:
-Would you like to create a .github directory? [y]: n
->
+☁  teddygroves  copier copy gh:teddygroves/bibat baseballl
+🎤 Name of your project
+   Baseball
+🎤 Name of your project, with no spaces (used for venv and package names)
+   baseball
+🎤 A short description of the project.
+   Is the generalised Pareto distribution good for modelling latent hitting
+ability?
+🎤 Your name (or your organization/company/team)
+   Teddy Groves
+🎤 Author email (will be included in pyproject.toml)
+   groves.teddy@gmail.com
+🎤 Code of conduct contact
+   groves.teddy@gmail.com
+🎤 open_source_license
+   MIT
+🎤 docs_format
+   Quarto
+🎤 create_dotgithub_directory (bool)
+   No
 ```
@@
 ```{.zsh}
+☁  teddygroves  tree baseball
 baseball
@@
 ├── README.md
-├── bibat_version.txt
 ├── data
@@
 │   │   └── readme.md
-│   ├── report.html
-│   ├── report.pdf
 │   └── report.qmd
-├── baseball
-│   ├── __init__.py
-│   ├── data_preparation.py
-│   ├── fitting_mode.py
-│   ├── inference_configuration.py
-│   ├── investigate.ipynb
-│   ├── notebooks
-│   ├── readme.md
-│   ├── sample.py
-│   ├── stan
-│   │   ├── custom_functions.stan
-│   │   ├── multilevel-linear-regression.stan
-│   │   └── readme.md
-│   ├── stan_input_functions.py
-│   └── util.py
 ├── inferences
@@
 │       └── config.toml
+├── notebooks
+│   └── investigate.ipynb
 ├── plots
-│   ├── posterior_ll_comparison.png
-│   └── posterior_predictive_comparison.png
+│   └── readme.md
 ├── pyproject.toml
+├── src
+│   ├── __init__.py
+│   ├── data_preparation.py
+│   ├── fitting.py
+│   ├── stan
+│   │   ├── custom_functions.stan
+│   │   ├── multilevel-linear-regression.stan
+│   │   └── readme.md
+│   └── stan_input_functions.py
 └── tests
-    ├── test_integration
-    │   └── test_data_preparation.py
-    └── test_unit
-        ├── test_inference_configuration.py
-        └── test_util.py
+    └── test_integration
+        └── test_data_preparation.py
 ```
 
-This folder implements bibat's example analysis – a comparison of linear
-regression with two different design matrices. To check that everything was
+This folder implements bibat's example analysis, which compares linear
+regression models with different design matrices. To check that everything was
 working correctly I ran the analysis like this:
@@
 
-This ran without errors, running some data preparation functions and then analysing the data with Stan in a few configurations.
+This ran without errors, running some data preparation functions and then
+analysing the data with Stan in a few configurations.
 
@@
 
-```{.python include="../baseball/fetch_data.py"}
+```{.python include="../src/fetch_data.py"}
 ```
@@
 ```zsh
-> source .venv/bin/activate
-(baseball) > python baseball/fetch_data.py
-Fetching 2006 data from https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv
-Writing 2006 data to data/raw/2006.csv
-Fetching bdb-main data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Batting.csv
-Writing bdb-main data to data/raw/bdb-main.csv
-Fetching bdb-post data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/BattingPost.csv
-Writing bdb-post data to data/raw/bdb-post.csv
-Fetching bdb-apps data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Appearances.csv
-Writing bdb-apps data to data/raw/bdb-apps.csv
+(baseball) ☁  baseball  python src/fetch_data.py --verbose
+/Users/tedgro/repos/teddygroves/baseball/src/fetch_data.py:6: DeprecationWarning:
+Pyarrow will become a required dependency of pandas in the next major release of pandas (pandas 3.0),
+(to allow more performant data types, such as the Arrow string type, and better interoperability with other libraries)
+but was not found to be installed on your system.
+If this would cause problems for you,
+please provide us feedback at https://github.com/pandas-dev/pandas/issues/54466
+
+  import pandas as pd
+INFO:root:Fetching 2006 data from https://raw.githubusercontent.com/stan-dev/example-models/master/knitr/pool-binary-trials/baseball-hits-2006.csv
+INFO:root:Writing 2006 data to data/raw/2006.csv
+INFO:root:Fetching bdb-main data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Batting.csv
+INFO:root:Writing bdb-main data to data/raw/bdb-main.csv
+INFO:root:Fetching bdb-post data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/BattingPost.csv
+INFO:root:Writing bdb-post data to data/raw/bdb-post.csv
+INFO:root:Fetching bdb-apps data from https://raw.githubusercontent.com/cbwinslow/baseballdatabank/master/core/Appearances.csv
+INFO:root:Writing bdb-apps data to data/raw/bdb-apps.csv
 ```
@@
 ```zsh
-(baseball) > rm data/raw/raw_measurements.csv
+> rm data/raw/raw_measurements.csv
 ```
 
-# Preparing the data
+# Data Preparation
 
-The first step in preparing data is to decide what prepared data looks like for
-the purposes of our analysis. Bibat provides dataclasses called `PreparedData`
-and `MeasurementsDF` to help get started with this, which I found in the file
-`baseball/data_preparation.py`.
+The example project already has some data preparation code in the file
+`src/data_preparation.py`: I mostly just had to adapt what was already there.
 
-As it happens, prepared data looks very similar in this analysis and the
-example. All I had to do was change the `MeasurementsDF` definition a
-little^[note that this class uses [pandera](https://pandera.readthedocs.io), a
-handy library for defining what a pandas dataframe should look like]:
+The first data preparation step was to decide what prepared data should
+look like for my analysis and represent this definition using a subclass of
+bibat's `PreparedData` [Pydantic `BaseModel`](https://docs.pydantic.dev/latest/api/base_model/). Looking at the example analysis's
+`ExamplePreparedData` class, I could see that it already had the components
+that I needed, namely a name, a dictionary of coordinates and a table of
+measurements:
 
-```python
-class MeasurementsDF(pa.SchemaModel):
-    """A PreparedData should have a measurements dataframe like this.
-
-    Other columns are also allowed!
-    """
-
-    player_season: Series[str]
-    season: Series[str] = pa.Field(coerce=True)
-    n_attempt: Series[int] = pa.Field(ge = 1)
-    n_success: Series[int] = pa.Field(ge = 0)
+```{.python include="../../../bibat/template/src/data_preparation.py" start-line=44 end-line=57}
 ```
 
-The next step is to write functions that return `PreparedData` objects. In this
-case I wrote a couple of data preparation functions: `prepare_data_2006` and
-`prepare_data_bdb`:
+There are a few interesting things about this model. First, it uses some
+specialised types from bibat's `util` module, namely `CoordDict` and
+`DfInPydanticModel`. The latter is quite handy as it ensures that the table of
+measurements can be saved to json and then safely read back again. Second, the
+measurement table has a validator that refers to a [pandera `DataFrameModel`](https://pandera.readthedocs.io/en/stable/reference/generated/pandera.api.pandas.model.DataFrameModel.html?highlight=SchemaModel) called `ExampleMeasurementsDF` that defines what a measurement table should be like.
+It is defined just before `ExamplePreparedData`:
 
-```{.python include="../baseball/data_preparation.py" start-line=74 end-line=159}
+```{.python include="../../../bibat/template/src/data_preparation.py" start-line=32 end-line=41}
 ```
 
-To take into account the inconsistency between the two raw data sources, I
-first had to change the variable `RAW_DATA_FILES`:
+I only had to change a few things to make new definitions for my analysis:
 
-```{.python include="../baseball/data_preparation.py" start-line=26 end-line=33}
+```{.python include="../src/data_preparation.py" start-line=39 end-line=61}
+```
+
+The next step is to write functions that return `BaseballPreparedData` objects.
+In this case I wrote a couple of data preparation functions: `prepare_data_2006`
+and `prepare_data_bdb`:
+
+```{.python include="../src/data_preparation.py" start-line=64 end-line=157}
+```
+
+To take into account the inconsistency between my two raw data sources, I first
+had to change the variable RAW_DATA_FILES:
+
+```{.python include="../src/data_preparation.py" start-line=29 end-line=36}
 ```
@@
 
-```{.python include="../baseball/data_preparation.py" start-line=25 end-line=32}
+```{.python include="../src/data_preparation.py" start-line=159 end-line=177}
 ```
 
-To finish off I deleted the unused global variables `NEW_COLNAMES`,
-`DROPNA_COLS` and `DIMS`, then checked if the function `load_prepared_data`
-needed any changes: I was pretty sure it didn't.
+To finish off the data preparation step I deleted any unused code from the
+example analysis, and updated the function `load_prepared_data`'s signature:
+
+```{.python include="../src/data_preparation.py" start-line=180 end-line=185}
+```
 
@@
 
-Now the folder `data/prepared/bdb` contained a file
-`data/prepared/bdb/measurements.csv` that looked like this:
+Now the folder `data/prepared/` contained json files `2006.json` and `bdb.json`
+that looked like this:
 
-```{csv}
-,player_season,season,n_attempt,n_success
-0,abreujo02,2018,553,180
-1,acunaro01,2018,487,178
-3,adamewi01,2018,322,112
-6,adamsla01,2018,29,10
-7,adamsma01,2018,337,104
-8,adamsma01,2018,277,92
-9,adamsma01,2018,60,12
+```zsh
+(baseball) > jq . data/prepared/bdb.json | head -n 10
+{
+  "name": "bdb",
+  "coords": {
+    "player_season": [
+      "abreujo022017",
+      "abreujo022018",
+      "abreujo022019",
+      "abreujo022020",
+      "abreujo022021",
+      "acunaro012018",
 ```
@@
 
-```{.stan include="../baseball/stan/gpareto.stan"}
+```{.stan include="../src/stan/gpareto.stan"}
 ```
@@
 
-```{.stan include="../baseball/stan/normal.stan"}
+```{.stan include="../src/stan/normal.stan"}
 ```
@@
 Next I needed to tell the analysis how to turn some prepared data into a
-dictionary that can be used as input for Stan. Bibat assumes that this task is
-handled by functions that live in the file `baseball/stan_input_functions.py`,
-each of which takes in a `PreparedData` and returns a Python dictionary. You
-can write as many Stan input functions as you like and choose which one to run
-for any given inference.
+dictionary that can be used as input for Stan. The example analysis includes
+some functions that do this in the file `src/stan_input_functions.py`.
 
-I started by defining some Stan input functions that pass arbitrary prepared
-data on to each of the models:^[Note that this code uses the scipy function
-`logit`, which it imported like this: `from scipy.special import logit`]
+I followed the examples to create similar functions for my two Stan
+models:^[Note that this code uses the scipy function `logit`, which it imported
+like this: `from scipy.special import logit`]
 
-```{.python include="../baseball/stan_input_functions.py" start-line=11 end-line=34}
+```{.python include="../src/stan_input_functions.py" end-line=33}
 ```
 
-But why stop there? It can also be useful to generate Stan input data with a
-model, by running it in simulation mode with hardcoded parameter values. Here
-are some functions that do this for both of our models:
-^[These functions require some more imports: `numpy` and `scipy.special.expit`]
-```{.python include="../baseball/stan_input_functions.py" start-line=36}
+<!-- todo: add link for returns_stan_input below -->
+
+These function use a handy decorator called `returns_stan_input` that uses the
+power of [`stanio`](https://github.com/WardBrian/stanio) to convert a dictionary
+to a json serialisable, Stan-friendly form. Thanks to this decorator I didn't
+have to spend time appending `.values.tolist()` to the inputs that came from
+pandas objects.
+
+I now had functions that turned my prepared data into input for either of my
+Stan models. But why stop there? It can also be useful to generate fake input
+data, by running a model in simulation mode using hardcoded parameter values.
+Here are some functions that do this for both of my models: ^[These functions
+require some more imports: `numpy` and `scipy.special.expit`]
+
+```{.python include="../src/stan_input_functions.py" start-line=37}
+```
+
+# Configuring fitting
+
+To tell my analysis how to run inferences, I had to edit the file `src/
+fitting.py`. The only changes I made from the example analysis were to remove
+the `kfold` fitting mode, which I didn't need to use in this analysis, and to
+add my Stan input functions to the `LOCAL_FUNCTIONS` constant:
+
+```{.python include="../src/fitting.py"}
 ```
@@
 Now all the building blocks for making statistical inferences – raw data, data
-preparation rules, statistical models and recipes for turning prepared data
-into model inputs – were in place. The last step before actually running Stan
-was to write down how put the blocks together. Bibat has another concept for
-this, called 'inferences'.
+preparation rules, statistical models, recipes for turning prepared data into
+model inputs and a procedure for running inferences – were in place. The last
+step before actually running my analysis was to write down how put the blocks
+together. Bibat has another concept for this, called 'inferences'.
 
@@
   strings. The analysis should raise an error if you enter a non-existing value.
-- Both inferences are set to run in "prior" and "posterior" modes – the other
-  pre-defined mode is "kfold", but you can also write your own!
+- Both inferences are set to run in `prior` and `posterior` modes
 - You can enter arbitrary arguments to cmdstanpy's `CmdStanModel.sample` method
@@
 │   ├── config.toml
-│   └── idata.json
+│   └── idata
 └── normal2006
     ├── config.toml
-    └── idata.json
+    └── idata
 ```
@@
 Now that the inferences are ready it's time to check them out. Bibat provides a
-Jupyter notebook at `baseball/investigate.ipynb` for exactly this purpose. The
+Jupyter notebook at `notebooks/investigate.ipynb` for exactly this purpose. The
 notebook's main job is to create plots and save them in the `plots` directory
-when it is executed with the command `jupyter execute investigate.ipynb`, which
-is the final step in the chain of commands that is triggered by `make
+when it is executed with the command `jupyter execute notebooks/investigate.ipynb`,
+ which is the final step in the chain of commands that is triggered by `make
 analysis`.
@@
 everything I did, see
-[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/baseball/investigate.ipynb).
+[here](https://github.com/teddygroves/bibat/blob/main/bibat/example_projects/baseball/notebooks/investigate.ipynb).
 
@@
 To model the more recent data, all I had to do was create some new inference
-folders with appropriate `prepared_data_dir` fields in their `config.toml`
+folders with appropriate `prepared_data` fields in their `config.toml`
 files. For example, here is the `config.toml` file for the `gparetobdb`
@@
 After running `make analysis` one more time, I went back to the notebook
-`baseball/investigate.ipynb` and made plots of both models' posterior 1%-99%
+`notebooks/investigate.ipynb` and made plots of both models' posterior 1%-99%
 success rate intervals for both datasets:
@@
 `report.qmd` file
-[here](https://github.com/teddygroves/bibat/blob/main/bibat/examples/baseball/docs/report.qmd).
+[here](https://github.com/teddygroves/bibat/blob/main/example_projects/baseball/docs/report.qmd).

### 2024-02-15 example_projects/baseball/baseball/fitting_mode.py

@@ -1,154 +1,0 @@
-"""A general definition of a fitting mode, plus some mode instances."""
-
-import textwrap
-from enum import Enum
-from typing import Callable, Dict, Union
-
-import numpy as np
-import xarray as xr
-from cmdstanpy import CmdStanMCMC, CmdStanModel
-from pydantic import BaseModel
-from sklearn.model_selection import KFold
-
-
-class IdataTarget(str, Enum):
-    """An enum for choosing the group that a fitting mode writes to."""
-
-    prior = "prior"
-    posterior = "posterior"
-    log_likelihood = "log_likelihood"
-
-
-class FittingMode(BaseModel):
-    """A way of fitting a statistical model.
-
-    Each fitting mode has a name, an 'idata_target', i.e. the group that this
-    mode will create in the output InferenceData, and a function that performs
-    the fitting.
-
-    Inferences can be configured to use any of the fitting modes defined here
-    by including them by name in the top-level list 'modes'. For example:
-
-      ...
-      modes = ['prior', 'posterior', 'kfold']
-      ...
-
-    The idata_target must be one out of 'prior', 'posterior' and
-    'log_likelihood', and specifies which out of these InferenceData groups the
-    output will be written to.
-
-    Each fitting mode's fit function must match the signature specified in the
-    FittingMode class. This can also be changed, but must agree with the
-    'sample' module.
-
-
-    """
-
-    name: str
-    idata_target: IdataTarget
-    fit: Callable[
-        [CmdStanModel, Dict, Dict[str, str]],
-        Union[CmdStanMCMC, xr.DataArray],
-    ]
-
-
-def fit_prior(model: CmdStanModel, input_dict: dict, kwargs) -> CmdStanMCMC:
-    """Create a CmdStanMCMC from a model, data and config, in prior mode.
-
-    :param model: a CmdStanModel. It must have a data variable called
-    'likelihood': setting this variable to 0 should trigger prior mode,
-    typically by causing the code that performs likelihood evaluations not to
-    run.
-
-    :param input_dict: a Stan input dictionary. It doesn't need to have a
-    'likelihood' entry: if it does, it will be overwritten.
-
-    :param kwargs: keyword arguments for CmdStanModel.sample
-
-    """
-    input_dict_final = input_dict | {"likelihood": 0}
-    return model.sample(input_dict_final, **kwargs)
-
-
[…]

### 2024-02-15 example_projects/baseball/baseball/inference_configuration.py

@@ -1,128 +1,0 @@
-"""Definition of the InferenceConfiguration class."""
-
-import os
-from typing import Callable, Dict, List, Optional
-
-import toml
-from pydantic import BaseModel, Field, field_validator, model_validator
-
-from baseball import fitting_mode, stan_input_functions
-
-HERE = os.path.dirname(os.path.abspath(__file__))
-STAN_DIR = os.path.join(HERE, "stan")
-DEFAULT_DIMS = {"llik": ["observation"], "yrep": ["observation"]}
-DEFAULT_SAMPLE_KWARGS = {"show_progress": False}
-
-
-class InferenceConfiguration(BaseModel):
-    """Container for a path to a Stan model and some configuration.
-
-    For example, you may want to compare how well two stan programs fit the
-    same data, or how well the same model fits the data with different
-    covariates.
-
-    :param name: string name identifying the model configuration
-
-    :param stan_file: Path to a Stan program, with "/" even on windows
-
-    :param prepared_data_dir: Path to a directory containing prepared data
-
-    :param stan_input_function: function from src.stan_input_functions used to
-    get a Stan input dictionary from a PreparedData object.
-
-    :param sample_kwargs: dictionary of keyword arguments to
-    cmdstanpy.CmdStanModel.sample.
-
-    :param modes: which modes to run the model in. Choose one or more of the
-    AVAILABLE_MODES.
-
-    :param dims: map from parameter names to lists of coordinate names. See the
-    field names of the file "coords.json" in the prepared_data_dir for possible
-    coordinate names.
-
-    :param kfold_folds: How many kfold folds to run
-
-    :param cpp_options: valid choices for the `cpp_options` argument to
-    CmdStanModel
-
-    :param stanc_options: valid choices for the `cpp_options` argument to
-    CmdStanModel
-
-    """
-
-    name: str
-    stan_file: str
-    prepared_data_dir: str
-    stan_input_function: Callable
-    fitting_mode_names: List[str] = Field(alias="modes")
-    fitting_modes: List[fitting_mode.FittingMode]
-    sample_kwargs: dict = Field(default_factory=lambda: DEFAULT_SAMPLE_KWARGS)
-    dims: Dict[str, List[str]] = Field(default_factory=lambda: DEFAULT_DIMS)
-    mode_options: Optional[Dict[str, dict]] = None
-    kfold_folds: Optional[int] = None
-    cpp_options: Optional[dict] = None
[…]

### 2024-02-15 example_projects/baseball/baseball/sample.py

@@ -1,71 +1,0 @@
-"""Run all the inferences in the inferences folder."""
-
-import os
-
-import arviz as az
-import cmdstanpy
-
-from baseball.data_preparation import load_prepared_data
-from baseball.inference_configuration import load_inference_configuration
-
-HERE = os.path.dirname(__file__)
-RUNS_DIR = os.path.join(HERE, "..", "inferences")
-STAN_DIR = os.path.join(HERE, "stan")
-
-
-def main():
-    """Fit all inferences in all modes."""
-    run_dirs = [
-        os.path.join(RUNS_DIR, d)
-        for d in os.listdir(RUNS_DIR)
-        if os.path.isdir(os.path.join(RUNS_DIR, d))
-    ]
-    for run_dir in sorted(run_dirs):
-        ic = load_inference_configuration(os.path.join(run_dir, "config.toml"))
-        stan_file = os.path.join(STAN_DIR, ic.stan_file)
-        prepared_data_file = os.path.join(
-            "data",
-            "prepared",
-            ic.prepared_data_dir + ".json",
-        )
-        model = cmdstanpy.CmdStanModel(
-            stan_file=stan_file,
-            cpp_options=ic.cpp_options,
-            stanc_options=ic.stanc_options,
-        )
-        prepared_data = load_prepared_data(prepared_data_file)
-        stan_input_base = ic.stan_input_function(prepared_data)
-        idata_kwargs = {
-            "observed_data": stan_input_base,
-            "log_likelihood": "llik",
-            "coords": prepared_data.coords,
-            "dims": ic.dims,
-        }
-        llik_outputs = {}
-        for mode in ic.fitting_modes:
-            fit_kwargs = ic.sample_kwargs
-            if (
-                ic.mode_options is not None
-                and mode.name in ic.mode_options.keys()
-            ):
-                fit_kwargs |= ic.mode_options[mode.name]
-            output = mode.fit(model, stan_input_base, fit_kwargs)
-            if mode.idata_target in ["prior", "posterior"]:
-                idata_kwargs[mode.idata_target] = output
-                idata_kwargs[f"{mode.idata_target.value}_predictive"] = "yrep"
-            elif mode.idata_target == "log_likelihood":
-                llik_outputs[f"llik_{mode.name}"] = output
-            else:
-                raise ValueError(
-                    f"idata_target {mode.idata_target} is not yet supported",
-                )
[…]

### 2024-02-15 example_projects/baseball/baseball/stan_input_functions.py

@@ -1,106 +1,0 @@
-"""Functions for generating input to Stan from prepared data."""
-
-from typing import Any, Callable, Dict
-
-import numpy as np
-from scipy.special import expit, logit
-from stanio.json import process_dictionary
-
-from baseball.data_preparation import PreparedData
-
-
-def returns_stan_input(func: Callable[[Any], Dict]) -> Callable[[Any], Dict]:
-    """Decorate a function so it returns a json-serialisable dictionary."""
-
-    def wrapper(*args, **kwargs):
-        return process_dictionary(func(*args, **kwargs))
-
-    return wrapper
-
-
-@returns_stan_input
-def get_stan_input_normal(ppd: PreparedData) -> Dict:
-    """General function for creating a Stan input."""
-    return {
-        "N": len(ppd.measurements),
-        "K": ppd.measurements["n_attempt"].values,
-        "y": ppd.measurements["n_success"].values,
-        "prior_mu": [logit(0.25), 0.2],
-        "prior_tau": [0.2, 0.1],
-        "prior_b_K": [0, 0.03],
-    }
-
-
-@returns_stan_input
-def get_stan_input_gpareto(ppd: PreparedData) -> Dict:
-    """General function for creating a Stan input."""
-    return {
-        "N": len(ppd.measurements),
-        "K": ppd.measurements["n_attempt"].values,
-        "y": ppd.measurements["n_success"].values,
-        "min_alpha": logit(0.07),
-        "max_alpha": logit(0.5),
-        "prior_sigma": [1.5, 0.4],
-        "prior_k": [-0.5, 1],
-    }
-
-
-@returns_stan_input
-def get_stan_input_normal_fake(ppd: PreparedData) -> Dict:
-    """Generate fake Stan input consistent with the normal model."""
-    N = len(ppd.measurements)
-    rng = np.random.default_rng(seed=1234)
-    true_param_values = {
-        "mu": logit(0.25),
-        "tau": 0.18,  # 2sds is 0.19 to 0.32 batting average
-        "b_K": 0.04,  # slight effect of more attempts
-        "alpha_std": rng.random.normal(loc=0, scale=1, size=N),
-    }
-    K = ppd.measurements["n_attempt"].values
-    log_K_std = (np.log(K) - np.log(K).mean()) / np.log(K).std()
-    alpha = (
-        true_param_values["mu"]
-        + true_param_values["b_K"] * log_K_std
[…]

### 2024-02-15 example_projects/baseball/baseball/util.py

@@ -1,85 +1,0 @@
-"""Some handy python functions."""
-
-from typing import Dict, List, NewType, Optional, Union
-
-import numpy as np
-import pandas as pd
-
-CoordDict = NewType("CoordDict", Dict[str, List[str]])
-StanInputDict = Dict["str", Union[int, float, List]]
-
-
-def one_encode(s: pd.Series) -> pd.Series:
-    """Replace a series's values with 1-indexed integer factors.
-
-    :param s: a pandas Series that you want to factorise.
-
-    """
-    return pd.Series(pd.factorize(s)[0] + 1, index=s.index)
-
-
-def make_columns_lower_case(df: pd.DataFrame) -> pd.DataFrame:
-    """Make a DataFrame's columns lower case.
-
-    :param df: a pandas DataFrame
-    """
-    new = df.copy()
-    if isinstance(new.columns, pd.MultiIndex):
-        new.columns = pd.MultiIndex.from_arrays(
-            [
-                [c.lower() for c in new.columns.get_level_values(i)]
-                for i in range(len(new.columns.levels))
-            ],
-        )
-    else:
-        new.columns = pd.Index([c.lower() for c in new.columns])
-    return new
-
-
-def check_is_df(maybe_df) -> pd.DataFrame:
-    """Shut up the type checker."""
-    assert isinstance(maybe_df, pd.DataFrame)
-    return maybe_df
-
-
-def stanify_dict(d: Dict) -> StanInputDict:
-    """Make sure a dictionary is a valid Stan input.
-
-    :param d: input dictionary, possibly with wrong types
-    """
-    out: StanInputDict = {}
-    for k, v in d.items():
-        if not isinstance(k, str):
-            raise ValueError(f"key {k!s} is not a string!")
-        elif isinstance(v, pd.Series):
-            out[k] = v.to_list()
-        elif isinstance(v, pd.DataFrame):
-            out[k] = v.values.tolist()
-        elif isinstance(v, np.ndarray):
-            out[k] = v.tolist()
-        else:
-            out[k] = v
-    return out
-
-
-def standardise(
-    s: pd.Series,
-    mu: Optional[float] = None,
-    std: Optional[float] = None,
-) -> pd.Series:
-    """Standardise a series by subtracting mu and dividing by sd."""
-    if mu is None:
-        mu = s.mean()
-    if std is None:
-        std = s.std()
-    return s.subtract(mu).divide(std)
-
-
-def center(
[…]

### 2024-02-15 example_projects/baseball/src/data_preparation.py

@@ -1,42 +1,43 @@
-"""Provides function `prepare_data` and runs it.
+"""Provides data preparation definitions and functions.
 
 This function should run some other functions with names `prepare_data_x`, which
 each take in a dataframe of measurements and return a PreparedData object.
 
 """
 
+from __future__ import annotations
+
 import json
-import os
-from io import StringIO
+import logging
 from pathlib import Path
-from typing import Any
+from typing import TYPE_CHECKING
 
 import pandas as pd
 import pandera as pa
 from pandera.typing import DataFrame, Series
-from pandera.typing.common import DataFrameBase
-from pydantic import BaseModel, field_serializer, field_validator
+from pydantic import field_validator
 
-from baseball import util
+from bibat.prepared_data import PreparedData
+from bibat.util import CoordDict, DfInPydanticModel
+
+if TYPE_CHECKING:
+    from pandera.typing.common import DataFrameBase
 
 HERE = Path(__file__).parent
 RAW_DIR = HERE / ".." / "data" / "raw"
 PREPARED_DIR = HERE / ".." / "data" / "prepared"
 RAW_DATA_FILES = {
-    "2006": [os.path.join(RAW_DIR, "2006.csv")],
+    "2006": [RAW_DIR / "2006.csv"],
     "bdb": [
-        os.path.join(RAW_DIR, "bdb-main.csv"),
-        os.path.join(RAW_DIR, "bdb-post.csv"),
-        os.path.join(RAW_DIR, "bdb-apps.csv"),
+        RAW_DIR / "bdb-main.csv",
+        RAW_DIR / "bdb-post.csv",
+        RAW_DIR / "bdb-apps.csv",
     ],
 }
 
 
-class MeasurementsDF(pa.SchemaModel):
-    """A PreparedData should have a measurements dataframe like this.
-
-    Other columns are also allowed!
-    """
+class BaseballMeasurementsDF(pa.SchemaModel):
+    """A BaseballPreparedData should have a measurements table like this."""
 
     player_season: Series[str]
     season: Series[str] = pa.Field(coerce=True)
@@ -44,52 +45,42 @@ class MeasurementsDF(pa.SchemaModel):
     n_success: Series[int] = pa.Field(ge=0)
 
 
-class PreparedData(BaseModel, arbitrary_types_allowed=True):
+class BaseballPreparedData(PreparedData):
     """What prepared data looks like in this analysis."""
 
     name: str
-    coords: util.CoordDict
[…]

### 2024-02-15 example_projects/baseball/src/fetch_data.py

@@ -1,6 +1,7 @@
 """Fetch raw data from the internet."""
 
-import os
+import logging
+from pathlib import Path
 
 import pandas as pd
 
@@ -14,16 +15,18 @@ URLS = {
     "bdb-apps": "https://raw.githubusercontent.com/cbwinslow/"
     "baseballdatabank/master/core/Appearances.csv",
 }
+RAW_DIR = Path("data") / "raw"
 OUT_FILES = {
-    "2006": os.path.join("data", "raw", "2006.csv"),
-    "bdb-main": os.path.join("data", "raw", "bdb-main.csv"),
-    "bdb-post": os.path.join("data", "raw", "bdb-post.csv"),
-    "bdb-apps": os.path.join("data", "raw", "bdb-apps.csv"),
+    "2006": RAW_DIR / "2006.csv",
+    "bdb-main": RAW_DIR / "bdb-main.csv",
+    "bdb-post": RAW_DIR / "bdb-post.csv",
+    "bdb-apps": RAW_DIR / "bdb-apps.csv",
 }
 
 if __name__ == "__main__":
+    logging.basicConfig(level=logging.INFO)
     for name, url in URLS.items():
-        print(f"Fetching {name} data from {url}")
+        logging.info("Fetching %s data from %s", name, url)
         data = pd.read_csv(url, comment="#")
-        print(f"Writing {name} data to {OUT_FILES[name]}")
+        logging.info("Writing %s data to %s", name, OUT_FILES[name])
         data.to_csv(OUT_FILES[name])

### 2024-02-15 example_projects/baseball/src/fitting.py

@@ -3,24 +3,25 @@
 from pathlib import Path
 
 from bibat.fitting import run_all_inferences
-from bibat.fitting_mode import kfold_mode, posterior_mode, prior_mode
+from bibat.fitting_mode import posterior_mode, prior_mode
 from src.data_preparation import load_prepared_data
 from src.stan_input_functions import (
-    get_stan_input_interaction,
-    get_stan_input_no_interaction,
+    get_stan_input_gpareto,
+    get_stan_input_gpareto_fake,
+    get_stan_input_normal,
+    get_stan_input_normal_fake,
 )
 
 HERE = Path(__file__).parent
 INFERENCES_DIR = HERE / ".." / "inferences"
 PREPARED_DATA_DIR = HERE / ".." / "data" / "prepared"
-FITTING_MODE_OPTIONS = {
-    "prior": prior_mode,
-    "posterior": posterior_mode,
-    "kfold": kfold_mode,
-}
+STAN_DIR = HERE / "stan"
+FITTING_MODE_OPTIONS = {"prior": prior_mode, "posterior": posterior_mode}
 LOCAL_FUNCTIONS = {
-    "get_stan_input_interaction": get_stan_input_interaction,
-    "get_stan_input_no_interaction": get_stan_input_no_interaction,
+    "get_stan_input_gpareto": get_stan_input_gpareto,
+    "get_stan_input_gpareto_fake": get_stan_input_gpareto_fake,
+    "get_stan_input_normal": get_stan_input_normal,
+    "get_stan_input_normal_fake": get_stan_input_normal_fake,
 }
 
 if __name__ == "__main__":

### 2024-02-15 example_projects/baseball/src/stan_input_functions.py

@@ -1,0 +1,100 @@
+"""Functions for generating input to Stan from prepared data."""
+
+import numpy as np
+from scipy.special import expit, logit
+
+from bibat.util import returns_stan_input
+from src.data_preparation import BaseballPreparedData
+
+
+@returns_stan_input
+def get_stan_input_normal(ppd: BaseballPreparedData) -> dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"],
+        "y": ppd.measurements["n_success"],
+        "prior_mu": [logit(0.25), 0.2],
+        "prior_tau": [0.2, 0.1],
+        "prior_b_K": [0, 0.03],
+    }
+
+
+@returns_stan_input
+def get_stan_input_gpareto(ppd: BaseballPreparedData) -> dict:
+    """General function for creating a Stan input."""
+    return {
+        "N": len(ppd.measurements),
+        "K": ppd.measurements["n_attempt"],
+        "y": ppd.measurements["n_success"],
+        "min_alpha": logit(0.07),
+        "max_alpha": logit(0.5),
+        "prior_sigma": [1.5, 0.4],
+        "prior_k": [-0.5, 1],
+    }
+
+
+@returns_stan_input
+def get_stan_input_normal_fake(ppd: BaseballPreparedData) -> dict:
+    """Generate fake Stan input consistent with the normal model."""
+    n = len(ppd.measurements)
+    rng = np.random.default_rng(seed=1234)
+    true_param_values = {
+        "mu": logit(0.25),
+        "tau": 0.18,  # 2sds is 0.19 to 0.32 batting average
+        "b_K": 0.04,  # slight effect of more attempts
+        "alpha_std": rng.random.normal(loc=0, scale=1, size=n),
+    }
+    n_attempt = ppd.measurements["n_attempt"]
+    log_k_std = (np.log(n_attempt) - np.log(n_attempt).mean()) / np.log(
+        n_attempt,
+    ).std()
+    alpha = (
+        true_param_values["mu"]
+        + true_param_values["b_K"] * log_k_std
+        + true_param_values["tau"] * true_param_values["alpha_std"]
+    )
+    y = rng.random.binomial(n_attempt, expit(alpha))
+    return {"N": n, "K": n_attempt, "y": y} | true_param_values
+
+
+@returns_stan_input
[…]

### 2024-02-15 example_projects/baseball/tests/test_integration/test_data_preparation.py

@@ -1,0 +1,82 @@
+"""Integration tests for functions in src/data_preparation.py."""
+
+from collections.abc import Callable
+
+import pandas as pd
+import pytest
+from pandas.testing import assert_frame_equal
+
+from bibat.util import CoordDict
+from src.data_preparation import (
+    prepare_data_interaction,
+    prepare_data_no_interaction,
+)
+
+EXAMPLE_RAW_MEASUREMENTS = pd.DataFrame(
+    {
+        "X1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+        "X2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+        "yButIThoughtIdAddSomeLetters": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+    },
+)
+
+
+@pytest.mark.parametrize(
+    (
+        "prepare_data_function",
+        "name",
+        "raw_measurements",
+        "expected_measurements",
+        "expected_coords",
+    ),
+    [
+        (
+            prepare_data_interaction,
+            "interaction",
+            EXAMPLE_RAW_MEASUREMENTS,
+            pd.DataFrame(
+                {
+                    "x1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "y": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x1:x2": [1, 4, 9, 1, 4, 9, 1, 4, 9, 1, 4],
+                },
+            ).apply(lambda s: s.astype(float)),
+            {
+                "covariate": ["x1", "x2", "x1:x2"],
+                "observation": [str(i) for i in range(11)],
+            },
+        ),
+        (
+            prepare_data_no_interaction,
+            "no_interaction",
+            EXAMPLE_RAW_MEASUREMENTS,
+            pd.DataFrame(
+                {
+                    "x1": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x2": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "y": [1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2],
+                    "x1:x2": [1, 4, 9, 1, 4, 9, 1, 4, 9, 1, 4],
+                },
[…]

### 2024-02-15 example_projects/baseball/tests/test_unit/test_inference_configuration.py

@@ -1,97 +1,0 @@
-"""Unit tests for the InferenceConfiguration class."""
-
-import os
-
-import pytest
-
-from baseball.inference_configuration import InferenceConfiguration
-
-SAMPLE_KWARGS = {
-    "iter_warmup": 50,
-    "iter_sampling": 50,
-    "chains": 2,
-    "step_size": 0.01,
-    "adapt_delta": 0.7,
-    "show_progress": False,
-}
-MODES_GOOD = ["prior", "posterior", "kfold"]
-MODES_BAD = ["prio", "prserior", "cross-validation"]
-
-
-def test_model_configuration_good_modes():
-    """Check that an inference configuration with good modes initialises."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_gpareto",
-        sample_kwargs=SAMPLE_KWARGS,
-        mode_options={"kfold": {"n_folds": 10}},
-        modes=MODES_GOOD,
-        cpp_options=None,
-        stanc_options=None,
-    )
-
-
-@pytest.mark.xfail()
-def test_model_configuration_bad_modes():
-    """Check that an inference configuration with bad modes fails."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_gpareto",
-        sample_kwargs=SAMPLE_KWARGS,
-        modes=MODES_BAD,
-        mode_options={"kfold": {"n_folds": 10}},
-        cpp_options=None,
-        stanc_options=None,
-    )
-
-
-@pytest.mark.xfail()
-def test_model_configuration_no_k():
-    """Check that an inference configuration with no kfold options fails."""
-    _ = InferenceConfiguration(
-        name="my_mc",
-        stan_file="multilevel-linear-regression.stan",
-        prepared_data_dir=os.path.join("hi", "hello", "hey"),
-        stan_input_function="get_stan_input_gpareto",
-        sample_kwargs=SAMPLE_KWARGS,
-        mode_options={"kfold": None},  # This is the bad mode!
-        modes=MODES_GOOD,  # it would be ok if 'kfold' weren't in here.
-        cpp_options=None,
[…]

### 2024-02-15 example_projects/baseball/tests/test_unit/test_util.py

@@ -1,80 +1,0 @@
-"""Unit tests for functions in src/util.py."""
-
-import numpy as np
-import pandas as pd
-import pytest
-from pandas.testing import assert_frame_equal, assert_series_equal
-
-from baseball.util import (
-    StanInputDict,
-    make_columns_lower_case,
-    one_encode,
-    stanify_dict,
-)
-
-
-@pytest.mark.parametrize(
-    "s_in,expected",
-    [
-        (
-            pd.Series(["8", 1, "????"], index=["a", "b", "c"]),
-            pd.Series([1, 2, 3], index=["a", "b", "c"]),
-        ),
-        (
-            pd.Series([1, "????", "????"], index=["a", "b", "c"]),
-            pd.Series([1, 2, 2], index=["a", "b", "c"]),
-        ),
-    ],
-)
-def test_one_encode(s_in: pd.Series, expected: pd.Series):
-    """Check that the function one_encode works as expected."""
-    assert_series_equal(one_encode(s_in), expected)
-
-
-@pytest.mark.parametrize(
-    "df_in,expected",
-    [
-        (
-            pd.DataFrame({"A": [1, 2, 3], "B": ["a", "b", "c"]}),
-            pd.DataFrame({"a": [1, 2, 3], "b": ["a", "b", "c"]}),
-        ),
-        (
-            pd.DataFrame(
-                [[1, 1]],
-                columns=pd.MultiIndex.from_product([["A"], ["B", "C"]]),
-            ),
-            pd.DataFrame(
-                [[1, 1]],
-                columns=pd.MultiIndex.from_product([["a"], ["b", "c"]]),
-            ),
-        ),
-    ],
-)
-def test_make_columns_lower_case(df_in: pd.DataFrame, expected: pd.DataFrame):
-    """Check that the function make_columns_lower_case works as expected."""
-    assert_frame_equal(make_columns_lower_case(df_in), expected)
-
-
-@pytest.mark.parametrize(
-    "d_in,expected",
-    [
-        ({"a": pd.Series([1, 2, 3])}, {"a": [1, 2, 3]}),
-        (
-            {"a": pd.DataFrame([[1, 2, 3], [4, 5, 6]])},
-            {"a": [[1, 2, 3], [4, 5, 6]]},
-        ),
-        (
-            {
-                "a": np.array([[1, 2, 3], [4, 5, 6]]),
-            },
-            {"a": [[1, 2, 3], [4, 5, 6]]},
[…]

### 2024-02-15 bibat/fitting.py
add more tests

@@ -76,9 +76,6 @@ def run_inference(
             idata_kwargs[f"{mode.idata_target.value}_predictive"] = "yrep"
         elif mode.idata_target == "log_likelihood":
             llik_outputs[f"llik_{mode.name}"] = output
-        else:
-            msg = f"Unsupported idata_target {mode.idata_target}"
-            raise ValueError(msg)
     idata = az.from_cmdstanpy(**idata_kwargs)
     for varname, output in llik_outputs.items():
         idata.log_likelihood[varname] = output

### 2024-02-15 bibat/fitting_mode.py

@@ -108,9 +108,6 @@ def sample_hmc_kfold(
     CmdStanModel.sample
 
     """
-    if ic.mode_options is None:
-        msg = "k-fold mode requires mode_options"
-        raise ValueError(msg)
     k = ic.mode_options["kfold"]["n_folds"]
     kf = KFold(k, shuffle=True, random_state=1234)
     sif = local_functions[ic.stan_input_function]

### 2024-02-15 bibat/inference_configuration.py

@@ -50,7 +50,7 @@ class InferenceConfiguration(BaseModel):
     fitting_modes: list[str] = Field(alias="modes")
     sample_kwargs: dict = Field(default_factory=lambda: DEFAULT_SAMPLE_KWARGS)
     dims: dict[str, list[str]] = Field(default_factory=lambda: DEFAULT_DIMS)
-    mode_options: dict[str, dict] | None = None
+    mode_options: dict[str, dict] = Field(default_factory=dict)
     cpp_options: dict | None = None
     stanc_options: dict | None = None
 
@@ -58,7 +58,7 @@ class InferenceConfiguration(BaseModel):
     def check_folds(self: InferenceConfiguration) -> InferenceConfiguration:
         """Check that there is a number of folds if required."""
         if any(m == "kfold" for m in self.fitting_modes):
-            if self.mode_options is None:
+            if self.mode_options == {}:
                 msg = "Mode 'kfold' requires a mode_options.kfold table."
                 raise ValueError(
                     msg,
@@ -71,7 +71,8 @@ class InferenceConfiguration(BaseModel):
             if "n_folds" not in self.mode_options["kfold"]:
                 msg = "Set 'n_folds' field in kfold mode options."
                 raise ValueError(msg)
-            if not int(self.mode_options["kfold"]["n_folds"]):
+            mo = self.mode_options["kfold"]["n_folds"]
+            if isinstance(mo, str) and not mo.isdigit():
                 msg = (
                     f"Could not coerce n_folds choice "
                     f"{self.mode_options['kfold']['n_folds']} to int."

### 2024-02-15 bibat/template/src/data_preparation.py

@@ -29,7 +29,7 @@ PREPARED_DIR = HERE / ".." / "data" / "prepared"
 RAW_DATA_FILES = {"measurements": RAW_DIR / "raw_measurements.csv"}
 
 
-class ExampleMeasurementsDF(pa.SchemaModel):
+class ExampleMeasurementsDF(pa.DataFrameModel):
     """An ExamplePreparedData should have a measurements dataframe like this.
 
     Other columns are also allowed!

### 2024-02-15 bibat/util.py

@@ -73,11 +73,3 @@ def make_columns_lower_case(df: pd.DataFrame) -> pd.DataFrame:
     else:
         new.columns = pd.Index([c.lower() for c in new.columns])
     return new
-
-
-def check_is_df(maybe_df: Any) -> pd.DataFrame:  # noqa: ANN401
-    """Check that an object is a DataFrame."""
-    if not isinstance(maybe_df, pd.DataFrame):
-        msg = "Want dataframe."
-        raise TypeError(msg)
-    return maybe_df

### 2024-02-15 tests/test_unit/test_fitting.py

@@ -2,23 +2,43 @@
 
 import json
 import os
-from io import StringIO
 from pathlib import Path
-from typing import Any
 
 import pandas as pd
 import pytest
 import toml
-from pydantic import ConfigDict, field_serializer, field_validator
 
-from bibat.fitting import run_inference
-from bibat.fitting_mode import prior_mode
+from bibat.fitting import IdataSaveFormat, run_all_inferences, run_inference
+from bibat.fitting_mode import (
+    kfold_mode,
+    posterior_mode,
+)
 from bibat.inference_configuration import (
     InferenceConfiguration,
     load_inference_configuration,
 )
 from bibat.prepared_data import PreparedData
-from bibat.util import CoordDict, StanInputDict, returns_stan_input
+from bibat.util import (
+    CoordDict,
+    DfInPydanticModel,
+    StanInputDict,
+    returns_stan_input,
+)
+
+TEST_MODEL = """
+    data {
+        int N;
+        vector[N] y;
+    }
+    generated quantities {
+        vector[N] llik;
+        vector[N] yrep;
+        for (n in 1:N){
+          llik[n] = normal_lpdf(y[n] | 0, 1);
+          yrep[n] = normal_rng(0, 1);
+        }
+    }
+"""
 
 
 @pytest.fixture(scope="session")
@@ -27,41 +47,16 @@ def stan_file(tmp_path_factory: pytest.TempPathFactory) -> Path:
     stan_dir = tmp_path_factory.getbasetemp() / "src" / "stan"
     stan_dir.mkdir(parents=True, exist_ok=True)
     file = stan_dir / "multilevel-linear-regression.stan"
-    file.write_text(
-        "data {int N; vector[N] y;}"
-        "generated quantities {vector[N] yrep;"
-        "for (n in 1:N) yrep[n] = normal_rng(0, 1);}",
-    )
+    file.write_text(TEST_MODEL)
     return file
 
 
 class ExamplePreparedData(PreparedData):
     """An example prepared data dataclass."""
 
-    model_config = ConfigDict(arbitrary_types_allowed=True)
-
     name: str
     coords: CoordDict
-    measurements: Any
-
-    @field_validator("measurements")
-    def validate_measurements(
-        cls,  # noqa: N805, ANN101
-        v: Any,  # noqa: ANN401
-    ) -> pd.DataFrame:
-        """Validate the measurements table."""
-        if isinstance(v, str):
-            v = pd.read_json(StringIO(v))
-        return v
-
-    @field_serializer("measurements")
[…]

### 2024-02-15 tests/test_unit/test_fitting_mode.py

@@ -1,16 +1,109 @@
 """Tests for the fitting mode module."""
 
+import os
 from collections.abc import Callable
 from pathlib import Path
 
+import pandas as pd
 import pytest
+import toml
 from cmdstanpy import CmdStanMCMC, CmdStanModel
 
-from bibat.fitting_mode import FittingMode, IdataTarget
-from bibat.inference_configuration import InferenceConfiguration
+from bibat.fitting_mode import (
+    FittingMode,
+    IdataTarget,
+    sample_hmc_kfold,
+    sample_hmc_posterior,
+    sample_hmc_prior,
+)
+from bibat.inference_configuration import (
+    InferenceConfiguration,
+    load_inference_configuration,
+)
 from bibat.prepared_data import PreparedData
+from bibat.util import CoordDict, DfInPydanticModel
+from tests.test_unit.test_fitting import (
+    get_stan_input_interaction,
+    load_prepared_data,
+)
+
+TEST_MODEL = """
+    data {
+        int N;
+        vector[N] y;
+    }
+    generated quantities {
+        vector[N] llik;
+        vector[N] yrep;
+        for (n in 1:N){
+          llik[n] = normal_lpdf(y[n] | 0, 1);
+          yrep[n] = normal_rng(0, 1);
+        }
+    }
+"""
 
 
+class ExamplePreparedData(PreparedData):
+    """An example prepared data dataclass."""
+
+    name: str
+    coords: CoordDict
+    measurements: DfInPydanticModel
+
+
+@pytest.fixture(scope="session")
+def stan_file(tmp_path_factory: pytest.TempPathFactory) -> Path:
+    """Create a Stan file."""
+    stan_dir = tmp_path_factory.getbasetemp() / "src" / "stan"
+    stan_dir.mkdir(parents=True, exist_ok=True)
+    file = stan_dir / "multilevel-linear-regression.stan"
+    file.write_text(TEST_MODEL)
+    return file
+
+
+@pytest.fixture(scope="session")
+def prepared_data_json(tmp_path_factory: pytest.TempPathFactory) -> Path:
+    """Create a prepared data json file."""
+    data_dir = tmp_path_factory.getbasetemp() / "data" / "prepared"
+    data_dir.mkdir(parents=True, exist_ok=True)
+    path = data_dir / "interaction.json"
+    test_prepared_data = ExamplePreparedData(
+        name="example_prepared_data",
+        coords=CoordDict({"observation": ["a", "b"]}),
[…]

### 2024-02-15 tests/test_unit/test_inference_configuration.py

@@ -4,8 +4,12 @@ import os
 from pathlib import Path
 
 import pytest
+import toml
 
-from bibat.inference_configuration import InferenceConfiguration
+from bibat.inference_configuration import (
+    InferenceConfiguration,
+    load_inference_configuration,
+)
 
 SAMPLE_KWARGS = {
     "iter_warmup": 50,
@@ -27,6 +31,31 @@ def stan_file(tmp_path_factory: pytest.TempPathFactory) -> Path:
     file = stan_dir / "multilevel-linear-regression.stan"
     file.write_text("data {int N;}")
     return file
+
+
+@pytest.fixture(scope="session")
+def inference_config(tmp_path_factory: pytest.TempPathFactory) -> Path:
+    """Create an inference directory containing a toml file."""
+    inf_dir = tmp_path_factory.getbasetemp() / "inferences" / "example"
+    inf_dir.mkdir(parents=True, exist_ok=True)
+    os.chdir(inf_dir.parent.parent)
+    ic = InferenceConfiguration(
+        name="example",
+        prepared_data="interaction",
+        stan_file="multilevel-linear-regression.stan",
+        stan_input_function="get_stan_input_interaction",
+        modes=["posterior", "kfold"],
+        sample_kwargs={"chains": 1, "iter_warmup": 2, "iter_sampling": 2},
+        mode_options={
+            "kfold": {"n_folds": 2},
+            "prior": {"max_treedepth": 2},
+            "posterior": {"iter_warmup": 3},
+        },
+    )
+    path = inf_dir / "config.toml"
+    with path.open("w") as f:
+        toml.dump(ic.model_dump(by_alias=True), f)
+    return path
 
 
 def test_model_configuration_good_modes(stan_file: Path) -> None:
@@ -63,6 +92,40 @@ def test_model_configuration_bad_modes(stan_file: Path) -> None:
 
 
 @pytest.mark.xfail()
+def test_model_configuration_no_mode_options_kfold(stan_file: Path) -> None:
+    """Check that an inference configuration with no kfold options fails."""
+    os.chdir(stan_file.parent.parent.parent)
+    _ = InferenceConfiguration(
+        name="my_mc",
+        stan_file="multilevel-linear-regression.stan",
+        prepared_data=str(Path("hi") / "hello" / "hey"),
+        stan_input_function="get_stan_input_interaction",
+        sample_kwargs=SAMPLE_KWARGS,
+        mode_options={},  # This is the bad bit!
[…]

### 2024-02-15 tests/test_unit/test_util.py

@@ -1,5 +1,7 @@
 """Unit tests for functions in src/util.py."""
 
+import json
+
 import pandas as pd
 import pytest
 from pandas.testing import assert_frame_equal, assert_series_equal
@@ -7,6 +9,8 @@ from pandas.testing import assert_frame_equal, assert_series_equal
 from bibat.util import (
     make_columns_lower_case,
     one_encode,
+    returns_stan_input,
+    validate_df_or_string,
 )
 
 
@@ -53,3 +57,21 @@ def test_make_columns_lower_case(
 ) -> None:
     """Check that the function make_columns_lower_case works as expected."""
     assert_frame_equal(make_columns_lower_case(df_in), expected)
+
+
+def test_validate_df_or_string() -> None:
+    """Test the function validate_df_or_string."""
+    example_df = pd.DataFrame({"a": [1, 2]})
+    _ = validate_df_or_string(example_df)
+    _ = validate_df_or_string(str(example_df.to_json()))
+
+
+def test_returns_stan_input() -> None:
+    """Test the function returns_stan_input."""
+
+    @returns_stan_input
+    def does_return_stan_input() -> dict:
+        return {"a": pd.Series([1, 2, 3])}
+
+    stan_input = does_return_stan_input()
+    _ = json.dumps(stan_input)

### 2024-02-15 paper/main.qmd
Update paper

@@
 Bibat integrates many widely-adopted, open-source and powerful tools to
-implement the components of a Bayesian workflow. Bibat projects are is
-written in modern Python and uses pydantic @pydanticdevelopersPydantic2022
+implement the components of a Bayesian workflow. Bibat projects are  written
+in modern Python and uses pydantic [@pydanticdevelopersPydantic2022]
 and pandera [@niels_bantilan-proc-scipy-2020] for data modelling, Stan
@@
 for documentation.  Bibat itself is also written in modern Python and uses the
-popular tools cookiecutter [@greenfeldCookiecutter2021], pydantic and click
-[@clickdevelopersClickPythonComposable2022].
+previously mentioned software as well as the popular  popular template tool
+copier [@copierdevelopersCopier2024].
 
@@
 first class component of the analysis addresses a common problem in academic
-statistics projects where the paper gets out of sync with the code. Bibat itself
-is continuously tested to ensure that it works on the operating systems Linux,
+statistics projects where the paper gets out of sync with the code. Bibat is
+continuously tested to ensure that it works on the operating systems Linux,
 macOS and Windows. Bibat's continuous integration runs a test suite as well as
@@
 Bibat is developed in public and encourages community contribution.
-Please see the contributing page <https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst> and code of conduct <https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md> if you would like to help develop bibat.
+Please see the contributing page <https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.md> and code of conduct <https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md> if you would like to help develop bibat.
 

### 2024-02-15 paper/main.tex
Update paper

@@
 Bibat integrates many widely-adopted, open-source and powerful tools to
-implement the components of a Bayesian workflow. Bibat projects are is
-written in modern Python and uses pydantic Pydantic developers (2022)
+implement the components of a Bayesian workflow. Bibat projects are
+written in modern Python and uses pydantic (Pydantic developers 2022)
 and pandera (Niels Bantilan 2020) for data modelling, Stan (Carpenter et
@@
 2022) and quarto (Allaire et al. 2022) for documentation. Bibat itself
-is also written in modern Python and uses the popular tools cookiecutter
-(Greenfeld et al. 2021), pydantic and click (Click Developers 2022).
+is also written in modern Python and uses the previously mentioned
+software as well as the popular popular template tool copier (copier
+developers 2024).
 
@@
 common problem in academic statistics projects where the paper gets out
-of sync with the code. Bibat itself is continuously tested to ensure
-that it works on the operating systems Linux, macOS and Windows. Bibat's
+of sync with the code. Bibat is continuously tested to ensure that it
+works on the operating systems Linux, macOS and Windows. Bibat's
 continuous integration runs a test suite as well as an end-to-end
@@
 Please see the contributing page
-\url{https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.rst}
-and code of conduct
+\url{https://github.com/teddygroves/bibat/blob/main/CONTRIBUTING.md} and
+code of conduct
 \url{https://github.com/teddygroves/bibat/blob/main/CODE_OF_CONDUCT.md}
@@
 
-\bibitem[\citeproctext]{ref-clickdevelopersClickPythonComposable2022}
-Click Developers. 2022. {``Click: {Python} Composable Command Line
-Interface Toolkit.''} Pallets. \url{https://pypi.org/project/click/}.
+\bibitem[\citeproctext]{ref-copierdevelopersCopier2024}
+copier developers. 2024. {``Copier.''} copier-org.
+\url{https://github.com/copier-org/copier}.
 
@@
 \url{https://www.sphinx-doc.org/}.
-
-\bibitem[\citeproctext]{ref-greenfeldCookiecutter2021}
-Greenfeld, Audrey Roy, Dainiel Roy Greenfeld, Raphael Pierzina, et al.
-2021. {``Cookiecutter.''} \url{https://pypi.org/project/cookiecutter/}.
 

### 2024-02-16 example_projects/baseball/docs/report.qmd

@@
 
-```{.python include="../../../bibat/template/src/data_preparation.py" start-line=44 end-line=57}
+```{.python include="../../../template/src/data_preparation.py" start-line=44 end-line=57}
 ```
@@
 
-```{.python include="../../../bibat/template/src/data_preparation.py" start-line=32 end-line=41}
+```{.python include="../../../template/src/data_preparation.py" start-line=32 end-line=41}
 ```

### 2024-02-26 paper/automl-template.tex
add automl paper stuff

\documentclass[11pt]{article}

\usepackage{microtype} % microtypography
\usepackage{booktabs}  % tables
\usepackage{url}  % urls

% With no package options, the submission will be anonymized, the supplemental
% material will be visible, and line numbers will be added to the manuscript.
%
% To compile a non-anonymized camera-ready version, add the [final] option:
%
% \usepackage[final]{automl}
%
% To compile for the workshop track, add the [workshop] option:
%
% \usepackage[workshop]{automl}
% \usepackage[final,workshop]{automl}
%
% To compile a single-blind submission, e.g. for the ABCD track, add the
% [revealauthors] option:
%
% \usepackage[revealauthors]{automl}
%
% To compile a non-anonymized pre-print version that can be uploaded to arXiv,
% add the [preprint] option:
%
% \usepackage[preprint]{automl}
%
% The [hidesupplement] option may be used to hide the supplementary material from any of the
% above commands, e.g.:
%
% \usepackage[hidesupplement,final,workshop]{automl}

\usepackage[revealauthors]{automl}

$pandoc.tex()$
$title.tex()$

\begin{document}
\maketitle
\begin{abstract}
$abstract$
\end{abstract}
$body$

\newpage
\section*{Submission Checklist}

% All submissions must include a section containing the AutoML submission
% checklist, shown below. The submission checklist does not count towards the
% page limit and must be included at submission and camera-ready time.  The
% submission checklist draws upon related submission checklists:
%
% the NeurIPS '22 checklist: https://neurips.cc/Conferences/2022/PaperInformation/PaperChecklist
%     and the NAS checklist: https://www.automl.org/wp-content/uploads/NAS/NAS_checklist.pdf
%
% For each question, change the default \answerTODO{} to either:
%
%     \answerYes{[justification]},
%     \answerNo{[justification]}, or
%     \answerNA{[justification]}.

% *You must include a brief justification to your answer,} either by referencing
% *the appropriate section of your paper or providing a brief inline
% *description.* For example:
%
% -  Did you include the license of the code and datasets?
%
%    \answerYes{See Section~\ref{sec:code}.
%
% - Did you include all the code for running experiments?
%
%   \answerNo{We include the code we wrote for conducting the experiments, but
%   complete replication depends on proprietary libraries for executing on a
%   private compute cluster. The code therefore is not runnable without
%   modification. To compensate, we provide a runnable but non-parallelized
%   version of the code that could replicate the results at the expense of a
%   greater wall-clock time.}
%
% - Did you include the license of the datasets?
%
%   \answerNA{Our experiments were conducted on publicly available datasets and
%   we have not introduced new datasets.}
%
% Please note that if you answer a question with \verb|\answerNo{}|, we expect
% that you provide an explanation and/or compensation for the omission. For
% example, if you cannot provide complete evaluation code for some reason, you
% might instead provide code for a minimal reproduction of the main insights of
% your paper.
%
% Please do not modify the questions and only use the provided macros for your
% answers. Note that the submission checklist does not count towards the page
% limit.

\begin{enumerate}
\item For all authors\dots
  %
  \begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately
    reflect the paper's contributions and scope?
    %
    \answerYes{}
    %
  \item Did you describe the limitations of your work?
    %
    \answerYes{See section ``Limitations''}
    %
  \item Did you discuss any potential negative societal impacts of your work?
    %
    \answerYes{}
    %
  \item Did you read the ethics review guidelines and ensure that your paper
    conforms to them? \url{https://2022.automl.cc/ethics-accessibility/}
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you ran experiments\dots
  %
  \begin{enumerate}
  \item Did you use the same evaluation protocol for all methods being compared (e.g.,
    same benchmarks, data (sub)sets, available resources)?
    %
    \answerNA{}
    %
  \item Did you specify all the necessary details of your evaluation (e.g., data splits,
    pre-processing, search spaces, hyperparameter tuning)?
    %
    \answerNA{}
    %
  \item Did you repeat your experiments (e.g., across multiple random seeds or splits) to account for the impact of randomness in your methods or data?
    %
    \answerNA{}
    %
  \item Did you report the uncertainty of your results (e.g., the variance across random seeds or splits)?
    %
    \answerNA{}
    %
  \item Did you report the statistical significance of your results?
    %
    \answerNA{}
    %
  \item Did you use tabular or surrogate benchmarks for in-depth evaluations?
    %
    \answerNA{}
    %
  \item Did you compare performance over time and describe how you selected the maximum duration?
    %
    \answerNA{}
    %
  \item Did you include the total amount of compute and the type of resources
    used (e.g., type of \textsc{gpu}s, internal cluster, or cloud provider)?
    %
    \answerNA{}
    %
  \item Did you run ablation studies to assess the impact of different
    components of your approach?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\item With respect to the code used to obtain your results\dots
  %
  \begin{enumerate}
\item Did you include the code, data, and instructions needed to reproduce the
    main experimental results, including all requirements (e.g.,
    \texttt{requirements.txt} with explicit versions), random seeds, an instructive
    \texttt{README} with installation, and execution commands (either in the
    supplemental material or as a \textsc{url})?
    %
    \answerYes{}
    %
  \item Did you include a minimal example to replicate results on a small subset
    of the experiments or on toy data?
    %
    \answerYes{}
    %
  \item Did you ensure sufficient code quality and documentation so that someone else
    can execute and understand your code?
    %
    \answerYes{See pyopensci review \url{https://github.com/pyOpenSci/software-submission/issues/83}}
    %
  \item Did you include the raw results of running your experiments with the given
    code, data, and instructions?
    %
    \answerNA{}
    %
  \item Did you include the code, additional data, and instructions needed to generate
    the figures and tables in your paper based on the raw results?
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you used existing assets (e.g., code, data, models)\dots
  %
  \begin{enumerate}
  \item Did you cite the creators of used assets?
    %
    \answerYes{}
    %
  \item Did you discuss whether and how consent was obtained from people whose
    data you're using/curating if the license requires it?
    %
    \answerNA{}
    %
  \item Did you discuss whether the data you are using/curating contains
    personally identifiable information or offensive content?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\item If you created/released new assets (e.g., code, data, models)\dots
  %
  \begin{enumerate}
    %
    \item Did you mention the license of the new assets (e.g., as part of your code submission)?
    %
    \answerYes{}
    %
    \item Did you include the new assets either in the supplemental material or as
    a \textsc{url} (to, e.g., GitHub or Hugging Face)?
    %
    \answerYes{}
    %
  \end{enumerate}
  %
\item If you used crowdsourcing or conducted research with human subjects\dots
  %
  \begin{enumerate}
  \item Did you include the full text of instructions given to participants and
    screenshots, if applicable?
    %
    \answerNA{}
    %
  \item Did you describe any potential participant risks, with links to
    Institutional Review Board (\textsc{irb}) approvals, if applicable?
    %
    \answerNA{}
    %
  \item Did you include the estimated hourly wage paid to participants and the
    total amount spent on participant compensation?
    %
    \answerNA{}
    %
  \end{enumerate}
\item If you included theoretical results\dots
  %
  \begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    %
    \answerNA{}
    %
  \item Did you include complete proofs of all theoretical results?
    %
    \answerNA{}
    %
  \end{enumerate}
  %
\end{enumerate}

\end{document}

### 2024-02-26 paper/main.qmd
add automl paper stuff

@@
   an interactive template. Bibat is available on the Python Package index,
-  documented at <https://bibat.readthedocs.io/> and developed at <https://github.com/teddygroves/bibat/>. This paper explains the motivation for bibat,
-  describes intended usage, discusses bibat's design, compares bibat with similar software, highlights several examples of bibat's use in science and provides links to community resources associated with bibat.
+  documented at <https://bibat.readthedocs.io/> and developed at <https://github.com/teddygroves/bibat/>.
+Bibat is free to use under the MIT license.
+This paper explains the motivation for bibat, describes intended usage, discusses bibat's design, compares bibat with similar software, highlights several examples of bibat's use in science and provides links to community resources associated with bibat.
 number-sections: true
@@
 format:
-  typst: default
   html:
     embed-resources: true
-  latex: default
+  pdf:
+    template: automl-template.tex
+    template-partials:
+      - title.tex
 ---
@@
 The most novel part of bibat's design is the introduction of an abstraction
-called "fitting mode", which allows bibat project to handle different ways
-of fitting a model to a dataset appropriately and flexibly.  This is often
-necessary as part of a Bayesian workflow: for example, one might perform
-MCMC sampling of both the prior and posterior distributions, perform multiple
-leave-out-one-fold fits for cross-validation or need to compare MCMC sampling
-with an optimisation-based alternative.
+called "fitting mode", which allows bibat projects to handle fitting a model
+to a dataset in different ways. This is often necessary as part of a Bayesian
+workflow: for example, one might perform MCMC sampling of both the prior
+and posterior distributions, or perform multiple leave-out-one-fold fits for
+cross-validation, or compare MCMC-based posterior inference with an alternative
+computation method.
 
@@
 `FittingMode`. Each fitting mode contains a name, a function that fits a
-prepared dataset and instructions for how and where to save the results.
-For example, the provided prior sampling fitting mode creates a Stan input
-dictionary with the `likelihood` data variable set to `0`, performs MCMC
-sampling and writes data to the `InferenceData` group `prior`. This design
-allows multiple fits to be stored alongside each other naturally, facilitating
-many common Bayesian workflow analyses, and can easily be configured to prevent
-unnecessary computations.
+prepared dataset and instructions for how and where to save the results. For
+example, the provided prior sampling fitting mode is called "prior", contains
+a function that runs MCMC sampling with the `likelihood` input variable set
+to `0`, returning a `CmdStanMCMC` instance, and performs MCMC sampling and writes data to the `InferenceData` group
+`prior`. This design allows multiple fits to be stored alongside each other
+naturally, facilitating many common Bayesian workflow analyses, and can easily
+be configured to prevent unnecessary computations.
 
@@
 analyses where complex datasets or custom models might be required.
+
+# Limitations
+
+Using bibat effectively requires familiarity with Pydantic, pandera, arviz, Stan
+and managing a medium-sized Python project. Many statistical analysis projects
+do not require using these tools, for example if data preparation or validation
+is trivial, if custom statistical models are not required, or if the analysis
+can be carried out by a single script. Practitioners who wish to implement such
+Bayesian workflows may prefer to simply write their software from scratch rather
+than use bibat, using tools like bambi or brms to ensure that the software
+challenge remains manageable.
+
+Similarly, bibat includes features that not all analyses will need, such as
+accommodating plural inferences, fitting modes and datasets. As a result, bibat
+is not optimal for projects that are known in advance to be singular. This could
+happen, for example, because there is no uncertainty as to the necessary data
+manipulations, or because the aim of the project is to maintain and streamline
+an existing statistical model rather than to compare it with others.
+
+While inclusion of potentially unneeded features is an important limitation
+of bibat, it is worth noting that it is often difficult to predict in advance
+what features a Bayesian workflow project will require, and that it can
+be time-consuming to re-write a project after belatedly realising that an
+unsupported feature is needed. We therefore think there is value in using bibat
+as a default.
+
+Another limitation of bibat is that it makes many opinionated choices about
+which tools to use. In particular, languages other than Python, inference
+frameworks other than Stan and validation frameworks other than Pydantic are
+not supported. We think this limitation is best addressed by the development of
+additional analysis templates.
 
@@
 
-# Author contributions
+# Broader impact statement
 
-# Acknowledgements
+After careful reflection, the authors have determined that this work presents no
+notable negative impacts to society or the environment.
+
+\begin{acknowledgements}
+
+This research was funded by the Novo Nordisk Foundation (Grant numbers
+NNF20CC0035580 NNF14OC0009473).
+
+\end{acknowledgements}

### 2024-02-26 paper/automl-template.tex
Update paper

@@
     %
-    \answerYes{}
+    \answerNA{I can't think of any particular negative societal impacts of a Bayesian workflow template}
     %
@@
     %
-    \answerYes{}
+    \answerYes{See \url{https://bibat.readthedocs.io/en/latest/_static/report.html} for instructions to reproduce the main example.}
     %
@@
     %
-    \answerYes{}
+    \answerYes{See section "Generating Stan inputs" here \url{https://bibat.readthedocs.io/en/latest/_static/report.html\#preparing-the-data}}
     %
@@
     %
-    \answerNA{}
+    \answerNo{This is unnecessary as the results are easy to reproduce and the results are large files that would be awkward to store online}
     %
@@
     %
-    \answerYes{}
+    \answerNo{To reproduce the figures from raw data run "make analysis" from the folder \texttt{example\_projects/baseball}, as described in the instructions here \url{https://bibat.readthedocs.io/en/latest/\_static/report.html}}
     %
@@
     %
-    \answerYes{}
+    \answerYes{\url{https://github.com/teddygroves/bibat/}}
     %

### 2024-02-26 paper/main.qmd
Update paper

@@
 criticism. This idea can be found in @boxBayesianInferenceStatistical1992
-and has recently received increasing scholarly recognition of
+and has recently received increasing scholarly recognition
 [@gelmanBayesianWorkflow2020;@grinsztajnBayesianWorkflowDisease2021;
-@gabryVisualizationBayesianWorkflow2019]. Software tools now exist
-that address most individual aspects of a Bayesian workflow: see
-@strumbeljPresentFutureSoftware for a review of the state of the art.
+@gabryVisualizationBayesianWorkflow2019]. Software tools now exist that address
+most individual aspects of a Bayesian workflow: see @vstrumbelj2024past for a
+review of the state of the art.
 
@@
 
-Bibat can be installed on the Windows, Linux or macOS command line on by running
-the command `pip install bibat` and then used by running the command `bibat`.
-This command triggers an interactive form which prompts the user to select a
-range of customisation options. Bibat then creates a new directory containing
-code that implements an example analysis, with customisations reflecting the
-user's choices. This analysis works immediately, and can be reproduced with
-the single command `make analysis` without the need for any further action by
-the user: in this sense bibat comes with batteries included.
+In order to use bibat, a user must first install the templating library copier
+[@copierdevelopersCopier2024], then choose a directory name and run this command
+from the command line:
+
+```sh
+$ copier copy gh:teddygroves/bibat my-chosen-directory-name
+```
+
+This command triggers an interactive form which prompts the user to select
+a range of customisation options. The new directory will then be created
+if necessary and filled with code that implements an example analysis, with
+customisations reflecting the user's choices. This analysis works immediately,
+and can be reproduced with the single command `make analysis` without the need
+for any further action by the user: in this sense bibat comes with batteries
+included.
 
@@
 shows how it proceeds: the project team edits the code components, then runs
-`make analysis` triggering creation of the result components. After inspecting
+`make analysis`, triggering creation of the result components. After inspecting
 these they repeat the process, leading to a cycle that ultimately results in a
@@
 restrictions on the components: for example, datasets need not be singular or
-tabular and statistical models need not be representable in formula syntax.
+tabular, and it is possible to use any statistical model that Stan can compile.
 Thanks to these accommodations a project team using bibat should typically not
@@
 
-Bibat encourages reproducibility mainly by providing a preconfigured makefile
+Bibat encourages reproducibility by providing a preconfigured makefile
 with a target `analysis` that triggers creation of an isolated environment,
@@
 to install Stan, especially on Windows. A second way in which bibat encourages
-reproducibility is by providing its Python code in the form of a preconfigured
-package. Thanks to the use of modern conventions for specifying dependencies
-and configuring tooling, we expect that this will make bibat analyses easier
-for other researchers to work with, thereby facilitating replication while also
-making initial development more convenient.
+reproducibility is by providing a preconfigured Python project following modern
+conventions, making bibat analyses straightforward to replicated and extend for
+other researchers who are already familiar with these conventions.
 
-Bibat integrates many widely-adopted, open-source and powerful tools to
-implement the components of a Bayesian workflow. Bibat projects are written
-in modern Python and uses pydantic [@pydanticdevelopersPydantic2022]
-and pandera [@niels_bantilan-proc-scipy-2020] for data modelling, Stan
-[@carpenterStanProbabilisticProgramming2017] for statistical inference,
-cmdstanpy [@standevelopmentteamCmdStanPy2022] for python-Stan interface, arviz
-[@kumarArviZUnifiedLibrary2019] for storing and analyzing inferences and sphinx
-[@georgbrandlandthesphinxteamSphinx2022] and quarto [@Allaire_Quarto_2022]
-for documentation.  Bibat itself is also written in modern Python and uses the
-previously mentioned software as well as the popular  popular template tool
-copier [@copierdevelopersCopier2024].
+Bibat integrates many widely-adopted open-source tools to implement
+the components of a Bayesian workflow. These include pydantic
+[@pydanticdevelopersPydantic2022] and pandera [@niels_bantilan-proc-scipy-2020]
+for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
+statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
+python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
+analysing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
+quarto [@Allaire_Quarto_2022] for documentation.
 
@@
 projects include a preconfigured test environment, continuous integration,
-linting and pre-commit hooks. In addition, including documentation as a
-first class component of the analysis addresses a common problem in academic
-statistics projects where the paper gets out of sync with the code. Bibat is
-continuously tested to ensure that it works on the operating systems Linux,
-macOS and Windows. Bibat's continuous integration runs a test suite as well as
-an end-to-end functional test on all supported Python versions.
+linting and pre-commit hooks. In addition, bibat includes documentation as a
+first class, integrated component of the analysis, helping to keep it in sync
+with the other components.
+
+Bibat is continuously tested to ensure that it works on the operating systems
+Linux, macOS and Windows. Bibat's continuous integration runs a test suite as
+well as an end-to-end functional test on all supported Python versions.
 
@@
 example, the provided prior sampling fitting mode is called "prior", contains
-a function that runs MCMC sampling with the `likelihood` input variable set
-to `0`, returning a `CmdStanMCMC` instance, and performs MCMC sampling and writes data to the `InferenceData` group
-`prior`. This design allows multiple fits to be stored alongside each other
-naturally, facilitating many common Bayesian workflow analyses, and can easily
-be configured to prevent unnecessary computations.
+a function that runs MCMC sampling with the `likelihood` input variable set to
+`0`, returning a `CmdStanMCMC` instance, and specifies that this result should
+be written to the `InferenceData` group `prior`.  Bibat provides fitting modes
+corresponding for prior sampling, posterior sampling and k-fold posterior
+sampling. Users can easily implement additional fitting modes or modify the
+`FittingMode` class to achieve even richer functionality. Fitting modes can be
+referenced by name from the file that configures an inferece: for example, the
+following lines indicate that the inference should be run in prior, posterior
+and kfold modes:
 
-Bibat provides fitting modes corresponding for prior sampling, posterior
-sampling and k-fold posterior sampling. Users can easily add additional fitting
-modes by copying these examples, or modify the `FittingMode` class to achieve
-even richer functionality.
+```toml
+modes = ["prior", "posterior", "kfold"]
+```
+
+Fitting modes allow bibat projects to succinctly but flexibly declare how to
+perform inferences, and allow results corresponding to the same inference to be
+stored alongside each other appropriately.
 
@@
 
-Other than bibat, there is currently no interactive template that specifically
+Other than bibat, we are not aware of any interactive template that specifically
 targets Bayesian workflow projects. There are some templates that arguably
-encompass Bayesian workflow as a special case of data analysis project, such
-as cookiecutter-data-science [@drivendataCookiecutterdatascience2022], but
-these are of limited use compared with a specialised template due to the many
-specificities of Bayesian workflow.
+encompass Bayesian workflow as a special case of data analysis project,
+such as cookiecutter-data-science [@drivendataCookiecutterdatascience2022],
+but these are of limited use compared with a specialised template due to
+the many specificities of Bayesian workflow. cookiecutter-cmdstanpy-wrapper
+[@wardWardBrianCookiecuttercmdstanpywrapper2024] is an interactive template that
+targets a different use case than Bayesian workflow projects, namely setting up
+a Python package that provides pre-compiled Stan models.
 
@@
 
-Similarly, bibat includes features that not all analyses will need, such as
-accommodating plural inferences, fitting modes and datasets. As a result, bibat
-is not optimal for projects that are known in advance to be singular. This could
-happen, for example, because there is no uncertainty as to the necessary data
-manipulations, or because the aim of the project is to maintain and streamline
-an existing statistical model rather than to compare it with others.
-
-While inclusion of potentially unneeded features is an important limitation
-of bibat, it is worth noting that it is often difficult to predict in advance
-what features a Bayesian workflow project will require, and that it can
-be time-consuming to re-write a project after belatedly realising that an
-unsupported feature is needed. We therefore think there is value in using bibat
-as a default.
+Similarly, bibat accommodates plural inferences, fitting modes and datasets,
+but many analyses are singular in at least one of these components and could
+therefore be implemented more simply and concisely than an equivalent bibat
+project. On the other hand, it is typically difficult to predict in advance
+which components of a Bayesian workflow will be plural, and costly to re-write
+a project after mistakenly assuming that a component will be singular. While we
+acknowledge that accommodating potentially unneeded plurality is an important
+limitation of bibat, we nonetheless think that it is the correct choice for a
+general-purpose template.
 
@@
 which tools to use. In particular, languages other than Python, inference
-frameworks other than Stan and validation frameworks other than Pydantic are
-not supported. We think this limitation is best addressed by the development of
-additional analysis templates.
+frameworks other than Stan and validation frameworks other than Pydantic
+are not supported. We think that it is on the whole good for templates to be
+opinionated, as unopinionated templates are necessarily more complicated; this
+limitation of bibat is therefore best addressed by the development of additional
+analysis templates that make different choices.
 
@@
 statistical models, as latent generalised Pareto distributions are not supported
-by generic regression packages like brms and bambi. Further, bibat's modular
+by any available formula-based regression packages. Further, bibat's modular
 design made it easier to implement this medium-sized analysis with two datasets,
@@
 :::
-
-In @grovesteddySphincter2024, bibat was used to implement a large analysis
-of cerebrovascular data from mice, involving two raw datasets, 6 prepared
-datasets, 15 models and 15 inferences. Again it was not possible to implement
-the analysis using formula-based packages due to the need for custom modelling
-approaches to take into account expert knowledge about both the target system
-and the measurement apparatus. Due to the number of datasets and models that
-needed to be considered, implementing this analysis without using bibat or a
-similar structured approach would have required a combination of more time and/
-or compromises in software quality.
 

### 2024-04-05 paper/main.qmd

@@
   Bayesian statistical workflow offers a powerful way to learn from data,
-  but software software projects that implement complex Bayesian workflows
+  but software projects that implement complex Bayesian workflows
   in practice are unusual, partly due to the difficulty of orchestrating

### 2024-04-10 paper/main.qmd
Add hello world example to paper

@@
 
-In order to use bibat, a user must first install the templating library copier
-[@copierdevelopersCopier2024], then choose a directory name and run this command
-from the command line:
-
-```sh
-$ copier copy gh:teddygroves/bibat my-chosen-directory-name
-```
-
-This command triggers an interactive form which prompts the user to select
-a range of customisation options. The new directory will then be created
+To use bibat, a user must first install the templating library copier
+[@copierdevelopersCopier2024] and then use the command `copier copy` to trigger
+bibat. The user is then presented with an interactive form which prompts them
+to select from a range of customisation options. A new directory is then created
 if necessary and filled with code that implements an example analysis, with
@@
 a section discussing accessibility considerations.
+
+# 'Hello world' example
+
+This section illustrates bibat's use by stepping through a typical example.
+A similar description can be found on the 'Getting started' page of bibat's
+documentation.
+
+The first step to use bibat is installing copier. This can be done using pipx
+[@pipxdevelopersPipx2024] as follows:
+
+```sh
+$ pipx install copier
+```
+
+Next, choose a directory name, e.g. `hello-world` and call trigger bibat using
+copier:
+
+```sh
+$ copier copy gh:teddygroves/bibat hello-world
+```
+
+Here is what should appear next. Some options are text fields, and others, such
+as `open_source_license` are enumerated choices.
+
+```sh
+Welcome to bibat, the batteries-included Bayesian analysis template!
+
+You'll be asked a series of questions whose answers will be used to
+generate a tailored project for you.
+
+🎤 Name of your project
+   Hello world
+🎤 Name of your project, with no spaces (used for venv and package names)
+   hello_world
+🎤 A short description of the project.
+   A 'hello world example of using bibat'
+🎤 Your name (or your organization/company/team)
+   Hello world developers
+🎤 Author email (will be included in pyproject.toml)
+   author@email.com
+🎤 Code of conduct contact
+   author@email.com
+🎤 open_source_license
+   MIT
+🎤 docs_format
+   Quarto
+🎤 create_dotgithub_directory (bool)
+   Yes
+```
+
+After all the fields are completed, the directory `hello-world` is created if it
+doesn't exist and populated or edited based on the form. If this is successful,
+some more messages appear indicating which files have been created or edited and
+explaining what to do next.
+
+The final setup step is to change directory into the new folder and run `make analysis`:
+
+```sh
+$ cd hello-world
+$ make analysis
+```
+
+This will set up a suitable Python virtual environment, install non-Python
+dependencies if necessary and then run the analysis, creating the results shown
+on the right-hand side of @fig-workflow.
 

### 2024-04-10 paper/main.qmd
Mention lovelyplots in paper

@@
 [@pydanticdevelopersPydantic2022] and pandera [@niels_bantilan-proc-scipy-2020]
-for data modelling, Stan [@carpenterStanProbabilisticProgramming2017] for
-statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022] for
-python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019] for storing and
-analysing inferences and sphinx [@georgbrandlandthesphinxteamSphinx2022] and
-quarto [@Allaire_Quarto_2022] for documentation.
+for data modelling, Stan [@carpenterStanProbabilisticProgramming2017]
+for statistical inference, cmdstanpy [@standevelopmentteamCmdStanPy2022]
+for python-Stan interface, arviz [@kumarArviZUnifiedLibrary2019]
+for storing and analysing inferences, matplotlib [@Hunter:2007]
+and lovelyplots [@sheriff_lovelyplots_2022] for plotting and sphinx
+[@georgbrandlandthesphinxteamSphinx2022] and quarto [@Allaire_Quarto_2022] for
+documentation.
 
